---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rcola

<!-- badges: start -->
[![R-CMD-check](https://github.com/Marwolaeth/rcola/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Marwolaeth/rcola/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/Marwolaeth/rcola/graph/badge.svg)](https://app.codecov.io/gh/Marwolaeth/rcola)
<!-- badges: end -->

R package for sentiment and stance analysis using the COLA framework

## Overview

**rcola** is an R implementation of the COLA (**C**ollaborative r**O**le-infused **L**LM-based **A**gents – see Lan et al. (2024)) framework for automated stance detection in texts. The package employs a three-stage collaborative approach where specialised LLM-based agents analyse texts from linguistic, domain-specific, and social perspectives to determine whether an author's position is in favour, against, or neutral towards a given target.

## Installation

You can install the development version of rcola from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("Marwolaeth/rcola")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example, eval=FALSE}
library(rcola)
library(ellmer)

chat <- ellmer::chat_openai()

tweet <- "The carbon tax is just another way for the government to control our lives and stifle economic growth. It's a job killer!"

target <- "Climate Change is Real Concern"

llm_stance(
    tweet,
    target,
    type = "statement",
    chat_base = chat,
    domain_role = "social commentator"
)
```

## Requirements

- R >= 4.1.0
- {ellmer} for LLM integration
- API access to supported language models (OpenAI, Anthropic, etc.)

## Citation & Attribution

This implementation is based on the COLA framework from:

> Lan, X., Gao, C., Jin, D., & Li, Y. (2024). Stance Detection with Collaborative Role-Infused LLM-Based Agents. [arXiv preprint](https://arxiv.org/abs/2310.10467)

When using this package, please cite the original COLA paper.
