---
title: 'rcola: Collaborative Role-infused LLM-based Agents for Stance Detection'
output: html_document
editor_options: 
chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
library(cli)
library(ellmer)
library(yaml)
```

```{r description}
fusen::fill_description(
  pkg = here::here(),
  fields = list(
    Title = "LLM-based Sentiment and Stance Analysis with COLA Framework",
    Description = "An R package for sentiment and stance detection using the COLA (Collaborative rOle-infused LLM-based Agents) framework. Provides functions for single and batch text analysis with support for multiple language models via the ellmer package.",
    `Authors@R` = c(
      person("Andrej", "Pawluczenko", email = "a.pawluczenko@gmail.com", role = c("aut", "cre"))
    )
  ),
  overwrite = TRUE
)
# Define License with use_*_license()
usethis::use_ccby_license()
```

# Introduction

## Overview

**rcola** is an R implementation of the COLA (**C**ollaborative r**O**Le-infused **L**LM-based **A**gents) framework for automated stance detection in texts. 

The package employs a three-stage collaborative approach:

1. **Stage 1: Parallel Analysis** — Specialised LLM-based agents (linguist, domain expert, social interpreter) analyse texts from their respective perspectives
2. **Stage 2: Parallel Debates** — Agents debate their analyses to reach consensus
3. **Stage 3: Parallel Judgment** — A judge agent synthesises the debate and produces final stance classification

## Key Features

- 🤖 Multi-perspective analysis using collaborative LLM agents
- 🌍 Support for multiple languages
- 📊 Multiple stance scales (categorical, numeric, Likert)
- 🎯 Domain-specific analysis with custom roles
- 📈 Batch processing for data frames
- 🔍 Detailed inspection tools for results

## Quick Example

```r

```

---

# Utility Functions

## Text Truncation

```{r function-truncate}
#' Truncate numeric values to a range
#'
#' Constrains numeric values to fall within the range [a, b].
#' Values below a are set to a, values above b are set to b.
#'
#' @param x Numeric vector to truncate
#' @param a Numeric. Lower bound (minimum value)
#' @param b Numeric. Upper bound (maximum value)
#'
#' @return Numeric vector with values constrained to [a, b]
#' @keywords internal
#'
#' @examples
#' truncate(c(-1, 0, 5, 10), a = 0, b = 5)
#' # Returns: 0, 0, 5, 5

truncate <- function(x, a, b) {
    if (!is.numeric(x)) {
        cli::cli_abort("{.arg x} must be numeric")
    }
    if (!is.numeric(a)) {
        cli::cli_abort("{.arg a} must be numeric")
    }
    if (!is.numeric(b)) {
        cli::cli_abort("{.arg b} must be numeric")
    }
    
    pmin(pmax(x, a), b)
}
```

```{r tests-truncate}
# Test 1: Basic functionality - values within range
test_that('truncate returns values unchanged when within range', {
    expect_equal(truncate(3, a = 0, b = 5), 3)
    expect_equal(truncate(c(1, 2, 3), a = 0, b = 5), c(1, 2, 3))
})

# Test 2: Values below lower bound
test_that('truncate sets values below a to a', {
    expect_equal(truncate(-1, a = 0, b = 5), 0)
    expect_equal(truncate(c(-5, -1, 0), a = 0, b = 5), c(0, 0, 0))
})

# Test 3: Values above upper bound
test_that('truncate sets values above b to b', {
    expect_equal(truncate(10, a = 0, b = 5), 5)
    expect_equal(truncate(c(5, 6, 10), a = 0, b = 5), c(5, 5, 5))
})

# Test 4: Mixed values (below, within, above)
test_that('truncate handles mixed values correctly', {
    expect_equal(
        truncate(c(-1, 0, 3, 5, 10), a = 0, b = 5),
        c(0, 0, 3, 5, 5)
    )
})

# Test 5: Negative ranges
test_that('truncate works with negative ranges', {
    expect_equal(truncate(c(-10, -5, 0), a = -8, b = -2), c(-8, -5, -2))
})

# Test 6: Single value at boundaries
test_that('truncate handles boundary values', {
    expect_equal(truncate(0, a = 0, b = 5), 0)
    expect_equal(truncate(5, a = 0, b = 5), 5)
})

# Test 7: Vectorized operation
test_that('truncate is vectorized', {
    x <- c(-10, -5, 0, 5, 10, 15)
    result <- truncate(x, a = 0, b = 10)
    expect_equal(result, c(0, 0, 0, 5, 10, 10))
})

# Test 8: Floating point values
test_that('truncate works with floating point numbers', {
    expect_equal(truncate(3.5, a = 0, b = 5), 3.5)
    expect_equal(truncate(-0.5, a = 0, b = 5), 0)
    expect_equal(truncate(5.5, a = 0, b = 5), 5)
})

# Test 9: Very large and very small numbers
test_that('truncate handles extreme values', {
    expect_equal(truncate(1e10, a = 0, b = 100), 100)
    expect_equal(truncate(-1e10, a = 0, b = 100), 0)
})

# Test 10: Zero as bound
test_that('truncate works with zero as bound', {
    expect_equal(truncate(c(-5, 0, 5), a = 0, b = 10), c(0, 0, 5))
    expect_equal(truncate(c(-5, 0, 5), a = -10, b = 0), c(-5, 0, 0))
})

# Test 11: Equal bounds (a == b)
test_that('truncate works when a equals b', {
    expect_equal(truncate(c(-5, 0, 5), a = 3, b = 3), c(3, 3, 3))
})

# Test 12: Reversed bounds (a > b) - should still work with pmin/pmax
test_that('truncate handles reversed bounds', {
    # pmin and pmax will handle this gracefully
    result <- truncate(5, a = 10, b = 0)
    expect_equal(result, 0)
})

# Test 13: Empty vector
test_that('truncate handles empty vectors', {
    expect_equal(truncate(numeric(0), a = 0, b = 5), numeric(0))
})

# Test 14: NA values are preserved
test_that('truncate preserves NA values', {
    result <- truncate(c(1, NA, 3), a = 0, b = 5)
    expect_equal(result[1], 1)
    expect_true(is.na(result[2]))
    expect_equal(result[3], 3)
})

# Test 15: Inf values
test_that('truncate handles infinite values', {
    expect_equal(truncate(Inf, a = 0, b = 5), 5)
    expect_equal(truncate(-Inf, a = 0, b = 5), 0)
})

# Error handling: x is not numeric
test_that('truncate rejects non-numeric x', {
    expect_error(truncate("text", a = 0, b = 5), 'must be numeric')
    expect_error(truncate(TRUE, a = 0, b = 5), 'must be numeric')
    expect_error(truncate(list(1, 2), a = 0, b = 5), 'must be numeric')
})

# Error handling: a is not numeric
test_that('truncate rejects non-numeric a', {
    expect_error(truncate(5, a = "zero", b = 5), 'must be numeric')
    expect_error(truncate(5, a = list(0), b = 5), 'must be numeric')
})

# Error handling: b is not numeric
test_that('truncate rejects non-numeric b', {
    expect_error(truncate(5, a = 0, b = "five"), 'must be numeric')
    expect_error(truncate(5, a = 0, b = list(5)), 'must be numeric')
})

# Error handling: all three are non-numeric
test_that('truncate rejects multiple non-numeric arguments', {
    expect_error(truncate("x", a = "a", b = "b"), 'must be numeric')
})

# Error handling: NULL values
test_that('truncate rejects NULL values', {
    expect_error(truncate(NULL, a = 0, b = 5), 'must be numeric')
    expect_error(truncate(5, a = NULL, b = 5), 'must be numeric')
    expect_error(truncate(5, a = 0, b = NULL), 'must be numeric')
})
```

## Error Handling

```{r function-abort}
#' Abort with formatted error message
#'
#' @param ... Arguments passed to cli::cli_abort()
#'
#' @return Throws error
#' @keywords internal

abort <- function(msg) {
    cli::cli_abort(
        msg,
        call = rlang::expr(llm_stance()),
        .envir = rlang::caller_env()
    )
}
```

```{r tests-abort}
# Test 1: abort throws an error
test_that('abort throws an error', {
    expect_error(abort("Test error message"))
})

# Test 2: abort with simple string message
test_that('abort with simple string message', {
    expect_error(
        abort("This is an error"),
        "This is an error"
    )
})

# Test 3: abort with cli formatting - .arg
test_that('abort with .arg formatting', {
    expect_error(
        abort("{.arg x} must be numeric"),
        "must be numeric"
    )
})

# Test 4: abort with cli formatting - .val
test_that('abort with .val formatting', {
    expect_error(
        abort("Invalid value: {.val {5}}"),
        "Invalid value"
    )
})

# Test 5: abort with cli formatting - .cls
test_that('abort with .cls formatting', {
    expect_error(
        abort("Expected {.cls character}, got {.cls numeric}"),
        "Expected"
    )
})

# Test 6: abort with multiple cli formatting elements
test_that('abort with multiple formatting elements', {
    expect_error(
        abort(
            "Argument {.arg x} must be {.cls numeric}, not {.cls list}"
        ),
        "must be"
    )
})

# Test 7: abort with interpolation - variable substitution
test_that('abort with variable interpolation', {
    var_name <- "test_var"
    expect_error(
        abort("Variable '{var_name}' not found"),
        "test_var"
    )
})

# Test 8: abort with complex interpolation
test_that('abort with complex interpolation', {
    lang <- "en"
    expect_error(
        abort("Language '{lang}' not available"),
        "Language 'en' not available"
    )
})

# Test 9: abort preserves error class
test_that('abort creates rlang_error', {
    expect_error(
        abort("Test error"),
        class = "rlang_error"
    )
})

# Test 10: abort with empty message
test_that('abort with empty message', {
    expect_error(abort(""))
})

# Test 11: abort with very long message
test_that('abort with long message', {
    long_msg <- paste(rep("a", 1000), collapse = "")
    expect_error(abort(long_msg))
})

# Test 12: abort with special characters
test_that('abort with special characters', {
    expect_error(
        abort("Error with special chars: @#$%^&*()"),
        "special chars"
    )
})

# Test 13: abort with quotes in message
test_that('abort with quotes in message', {
    expect_error(
        abort('Error with "double" and \'single\' quotes'),
        "quotes"
    )
})

# Test 14: abort with newlines in message
test_that('abort with newlines', {
    expect_error(
        abort("First line\nSecond line"),
        "First line"
    )
})

# Test 15: abort with list formatting
test_that('abort with list formatting', {
    expect_error(
        abort("Available options: {.val {c('a', 'b', 'c')}}"),
        "Available"
    )
})

# Test 16: abort with vector in message
test_that('abort with vector interpolation', {
    vals <- c("x", "y", "z")
    expect_error(
        abort("Invalid values: {.val {vals}}"),
        "Invalid"
    )
})

# Test 17: abort with numeric values in message
test_that('abort with numeric interpolation', {
    n <- 111
    expect_error(
        abort("Expected {n} items, got 0"),
        "Expected 111"
    )
})

# Test 18: abort with logical values in message
test_that('abort with logical values', {
    flag <- TRUE
    expect_error(
        abort("Flag is {flag}"),
        "Flag is TRUE"
    )
})

# Test 19: abort with NA values
test_that('abort with NA values', {
    na_val <- NA
    expect_error(
        abort("Value is {na_val}"),
        "Value"
    )
})

# Test 20: abort with function name in message
test_that('abort with function reference', {
    expect_error(
        abort("Function {.fn truncate} failed"),
        "failed"
    )
})

# Test 21: abort uses caller environment
test_that('abort uses caller environment', {
    test_var <- "from_caller"
    expect_error(
        abort("Variable: {test_var}"),
        "from_caller"
    )
})

# Test 22: abort in nested function call
test_that('abort in nested function', {
    inner_func <- function() {
        abort("Error from inner function")
    }
    
    outer_func <- function() {
        inner_func()
    }
    
    expect_error(
        outer_func(),
        "Error from inner function"
    )
})

# Test 23: abort does not return value (throws immediately)
test_that('abort does not return', {
    result <- tryCatch(
        {
            abort("Test")
            "should not reach here"
        },
        error = function(e) {
            "caught error"
        }
    )
    expect_equal(result, "caught error")
})

# Test 24: abort with .arg and variable
test_that('abort with .arg and variable name', {
    param_name <- "data"
    expect_error(
        abort("Argument {.arg {param_name}} is invalid"),
        "invalid"
    )
})

# Test 25: abort with .cls and actual class
test_that('abort with .cls and object class', {
    obj <- list(a = 1)
    expect_error(
        abort("Expected data.frame, got {.cls {class(obj)}}"),
        "Expected"
    )
})

# Test 26: abort with multiple .val elements
test_that('abort with multiple .val elements', {
    expect_error(
        abort(
            "Got {.val {1}}, expected one of {.val {c(2, 3, 4)}}"
        ),
        "Got"
    )
})

# Test 27: abort preserves message formatting
test_that('abort message is not modified', {
    msg <- "Exact error message"
    tryCatch(
        abort(msg),
        error = function(e) {
            expect_true(grepl(msg, conditionMessage(e)))
        }
    )
})

# Test 28: abort called multiple times creates different errors
test_that('multiple abort calls create separate errors', {
    error1 <- tryCatch(
        abort("Error 1"),
        error = function(e) e
    )
    
    error2 <- tryCatch(
        abort("Error 2"),
        error = function(e) e
    )
    
    expect_true(grepl("Error 1", conditionMessage(error1)))
    expect_true(grepl("Error 2", conditionMessage(error2)))
    expect_false(identical(error1, error2))
})

# Test 29: abort with complex nested formatting
test_that('abort with complex nested formatting', {
    lang <- "en"
    available <- c("en", "ru", "fr")
    expect_error(
        abort(
            "Language '{lang}' not available. Use: {.val {available}}"
        ),
        "not available"
    )
})

# Test 30: abort requires exactly one argument
test_that('abort requires msg argument', {
    expect_error(abort(), "missing")
})

# Test 31: abort with NULL message
test_that('abort with NULL message', {
    expect_error(abort(NULL), 'must be supplied')
})

# Test 32: abort with numeric message (coerced to string)
test_that('abort with numeric message', {
    expect_error(
        abort(123),
        "123"
    )
})

# Test 33: abort with expression in message
test_that('abort with expression evaluation', {
    x <- 5
    y <- 3
    expect_error(
        abort("Sum is {x + y}"),
        "Sum is 8"
    )
})

# Test 34: abort message with unicode characters
test_that('abort with unicode characters', {
    expect_error(
        abort("Hello \U1F680"),
        "Hello"
    )
})
```

## Expression Catching

```{r function-catch}
#' Safely evaluate expression and catch errors
#'
#' @param expr Expression to evaluate
#' @param expr_name Name of expression for error messages
#'
#' @return Result of expression or throws an error
#' @keywords internal

catch <- function(
        expr,
        expr_name = paste(deparse(substitute(expr)), collapse = '')
) {
    if (!rlang::is_scalar_character(expr_name)) {
        cli::cli_abort(
            "{.arg expr_name} must be a character scalar.
            Got {.cls {class(expr_name)}}"
        )
    }
    
    rlang::try_fetch(
        expr,
        error = function(e) {
            abort("Error in {expr_name}: {e$message}")
        }
    )
}
```

```{r tests-catch}
# Test 1: catch executes expression successfully
test_that('catch executes expression without error', {
    expect_no_error(catch(1 + 1))
})

# Test 2: catch with variable assignment
test_that('catch evaluates expression with variables', {
    x <- 5
    result <- catch(x * 2)
    expect_equal(result, 10)
})

# Test 3: catch with function call
test_that('catch executes function call', {
    result <- catch(sqrt(16))
    expect_equal(result, 4)
})

# Test 4: catch catches errors and wraps them
test_that('catch catches error and wraps message', {
    expect_error(
        catch(stop("Original error")),
        "Error in"
    )
})

# Test 5: catch preserves original error message
test_that('catch preserves original error message', {
    expect_error(
        catch(stop("Original error")),
        "Original error"
    )
})

# Test 6: catch with custom expr_name
test_that('catch with custom expr_name parameter', {
    expect_error(
        catch(stop("Test"), expr_name = "custom_operation"),
        "Error in custom_operation"
    )
})

# Test 7: catch with default expr_name from deparse
test_that('catch uses deparse(substitute(expr)) for default name', {
    expect_error(
        catch(1 / "1"),
        "Error in 1"
    )
})

# Test 8: catch rejects non-character expr_name
test_that('catch rejects non-character expr_name', {
    expect_error(
        catch(1 + 1, expr_name = 123),
        "must be a character scalar"
    )
})

# Test 9: catch rejects non-scalar character expr_name
test_that('catch rejects non-scalar character expr_name', {
    expect_error(
        catch(1 + 1, expr_name = c("a", "b")),
        "must be a character scalar"
    )
})

# Test 10: catch with complex expression
test_that('catch evaluates complex expression', {
    result <- catch({
        x <- 10
        y <- 20
        x + y
    })
    expect_equal(result, 30)
})

# Test 11: catch with error in nested function call
test_that('catch catches error in nested function', {
    expect_error(
        catch(sqrt(as.numeric(sqrt("NA")))),
        "Error in"
    )
})

# Test 12: catch with NULL expr_name throws error
test_that('catch rejects NULL expr_name', {
    expect_error(
        catch(1 + 1, expr_name = NULL),
        "must be a character scalar"
    )
})

# Test 13: catch with empty string expr_name
test_that('catch accepts empty string expr_name', {
    result <- catch(1 + 1, expr_name = "")
    expect_equal(result, 2)
})
```

## Input Validation

```{r function-validate_inputs}
#' Validate input length
#'
#' @param inputs Input vector
#' @param expected_length Expected length
#' @param input_name Informative name for error messages
#'
#' @return Invisibly TRUE or throws error
#' @keywords internal

validate_inputs <- function(inputs, expected_length, input_name = 'Inputs') {
    if (!rlang::is_scalar_character(input_name)) {
        cli::cli_abort("{.arg input_name} must be a character scalar.")
    }
    
    if (is.null(inputs) || length(inputs) != expected_length) {
        abort("{input_name} returned unexpected results")
    }
    
    invisible(TRUE)
}
```

```{r tests-validate_inputs}
# Test 1: validate_inputs returns TRUE invisibly for valid inputs
test_that('validate_inputs returns TRUE invisibly for valid inputs', {
    result <- validate_inputs(c(1, 2, 3), expected_length = 3)
    expect_true(result)
    expect_invisible(validate_inputs(c(1, 2, 3), expected_length = 3))
})

# Test 2: validate_inputs with matching length
test_that('validate_inputs accepts inputs with matching length', {
    expect_invisible(validate_inputs(list(a = 1, b = 2), expected_length = 2))
})

# Test 3: validate_inputs rejects NULL inputs
test_that('validate_inputs rejects NULL inputs', {
    expect_error(
        validate_inputs(NULL, expected_length = 1),
        "returned unexpected results"
    )
})

# Test 4: validate_inputs rejects mismatched length
test_that('validate_inputs rejects mismatched length', {
    expect_error(
        validate_inputs(c(1, 2, 3), expected_length = 2),
        "returned unexpected results"
    )
})

# Test 5: validate_inputs with custom input_name in error message
test_that('validate_inputs includes custom input_name in error', {
    expect_error(
        validate_inputs(c(1, 2), expected_length = 3, input_name = "Results"),
        "Results returned unexpected results"
    )
})

# Test 6: validate_inputs rejects non-character input_name
test_that('validate_inputs rejects non-character input_name', {
    expect_error(
        validate_inputs(c(1, 2, 3), expected_length = 3, input_name = 123),
        "must be a character scalar"
    )
})

# Test 7: validate_inputs rejects non-scalar character input_name
test_that('validate_inputs rejects non-scalar character input_name', {
    expect_error(
        validate_inputs(c(1, 2, 3), expected_length = 3, input_name = c("a", "b")),
        "must be a character scalar"
    )
})

# Test 8: validate_inputs with empty vector
test_that('validate_inputs rejects empty vector when expecting non-zero length', {
    expect_error(
        validate_inputs(c(), expected_length = 1),
        "returned unexpected results"
    )
})
```

```{r function-validate_character}
#' Validate a character argument
#'
#' @param x Input to validate
#'
#' @return Invisibly TRUE or throws error
#' @keywords internal

validate_character <- function(x) {
    arg_name <- deparse(substitute(x))
    
    if (!rlang::is_character(x)) {
        abort(
            "{.arg {arg_name}} must be a character vector, got {.cls {class(x)}}"
        )
    }
    if (length(x) == 0) {
        abort("{.arg {arg_name}} cannot be empty")
    }
    if (any(is.na(x)) || any(ellmer:::trim(x) == '')) {
        abort(
            c(
                "{.arg {arg_name}} contains empty strings.",
                "i" = "consider removing them before analysis."
            )
        )
    }
    
    invisible(TRUE)
}
```

```{r tests-validate_character}
# Test 1: validate_character accepts valid character vector
test_that('validate_character accepts valid character vector', {
    result <- validate_character(c("text1", "text2"))
    expect_true(result)
    expect_invisible(validate_character(c("text1", "text2")))
})

# Test 2: validate_character accepts single character string
test_that('validate_character accepts single character string', {
    expect_invisible(validate_character("single text"))
})

# Test 3: validate_character rejects non-character input
test_that('validate_character rejects numeric input', {
    expect_error(
        validate_character(123),
        "must be a character vector"
    )
})

# Test 4: validate_character rejects logical input
test_that('validate_character rejects logical input', {
    expect_error(
        validate_character(TRUE),
        "must be a character vector"
    )
})

# Test 5: validate_character rejects list input
test_that('validate_character rejects list input', {
    expect_error(
        validate_character(list("a", "b")),
        "must be a character vector"
    )
})

# Test 6: validate_character rejects empty character vector
test_that('validate_character rejects empty vector', {
    expect_error(
        validate_character(character()),
        "cannot be empty"
    )
})

# Test 7: validate_character rejects vector with NA values
test_that('validate_character rejects vector with NA', {
    expect_error(
        validate_character(c("text", NA, "more")),
        "contains empty strings"
    )
})

# Test 8: validate_character rejects vector with empty strings
test_that('validate_character rejects vector with empty strings', {
    expect_error(
        validate_character(c("text", "", "more")),
        "contains empty strings"
    )
})

# Test 9: validate_character rejects vector with whitespace-only strings
test_that('validate_character rejects whitespace-only strings', {
    expect_error(
        validate_character(c("text", "   ", "more")),
        "contains empty strings"
    )
})

# Test 10: validate_character includes argument name in error message
test_that('validate_character includes arg name in error', {
    my_text <- 123
    expect_error(
        validate_character(my_text),
        "my_text"
    )
})

# Test 11: validate_character with mixed whitespace and empty
test_that('validate_character rejects mixed empty and whitespace', {
    expect_error(
        validate_character(c("valid", "\t", "\n")),
        "contains empty strings"
    )
})

# Test 12: validate_character accepts vector with special characters
test_that('validate_character accepts special characters', {
    expect_invisible(
        validate_character(c("text@123", "привіт", "\U1F680 emoji"))
    )
})

# Test 13: validate_character error message includes info about removal
test_that('validate_character error includes removal suggestion', {
    expect_error(
        validate_character(c("text", NA)),
        "consider removing them"
    )
})
```

```{r function-validate_stage}
#' Validate stage configuration
#'
#' @param x Stage configuration
#' @param stage_name Name of stage for error messages
#'
#' @return Unmodified `x` or throws an error
#' @keywords internal

validate_stage <- function(x, stage_name) {
    if (is.null(x)) {
        abort("{stage_name} returned NULL")
    }
    x
}
```

```{r tests-validate_stage}
# Test 1: validate_stage returns input unchanged when not NULL
test_that('validate_stage returns input unchanged', {
    input <- list(a = 1, b = 2)
    result <- validate_stage(input, stage_name = "Processing")
    expect_identical(result, input)
})

# Test 2: validate_stage rejects NULL input
test_that('validate_stage rejects NULL input', {
    expect_error(
        validate_stage(NULL, stage_name = "MyStage"),
        "MyStage returned NULL"
    )
})
```

## Argument Recycling

```{r function-recycle_arg}
#' Recycle argument to specified length
#'
#' @param arg Values to recycle
#' @param n Target length
#'
#' @return Recycled argument values
#' @keywords internal

recycle_arg <- function(arg, n) {
    arg_name <- deparse(substitute(arg))
    
    if (length(arg) == 1) {
        return(rep(arg, n))
    } else if (length(arg) == n) {
        return(arg)
    } else {
        abort(
            "{.arg {arg_name}} must have length 1 or {n} (same as {.arg text}). \\
            Got {length(arg)}."
        )
    }
}
```

```{r tests-recycle_arg}
# Test 1: recycle_arg replicates single value to length n
test_that('recycle_arg replicates single value', {
    result <- recycle_arg("x", n = 3)
    expect_equal(result, c("x", "x", "x"))
})

# Test 2: recycle_arg returns unchanged vector of length n
test_that('recycle_arg returns vector of matching length', {
    input <- c("a", "b", "c")
    result <- recycle_arg(input, n = 3)
    expect_identical(result, input)
})

# Test 3: recycle_arg rejects vector with mismatched length
test_that('recycle_arg rejects mismatched length', {
    expect_error(
        recycle_arg(c("a", "b"), n = 3),
        "must have length 1 or 3"
    )
})

# Test 4: recycle_arg includes argument name in error message
test_that('recycle_arg includes arg name in error', {
    my_vector <- c(1, 2, 3, 4)
    expect_error(
        recycle_arg(my_vector, n = 2),
        "my_vector"
    )
})
```

## Timing

```{r function-stage_complete}
#' Format timing message
#'
#' @param tic Start time
#' @param toc End time
#' @param msg Message prefix
#'
#' @return Formatted timing message
#' @keywords internal

stage_complete <- function(tic, toc, msg) {
    tocmsg <- paste0(round(toc - tic, 3), " seconds elapsed")
    
    if (is.null(msg) || is.na(msg) || length(msg) == 0) {
        outmsg <- tocmsg
    } else {
        outmsg <- glue::glue('\u2705 {msg} complete – {tocmsg}\n\n')
    }
    outmsg
}
```

```{r tests-stage_complete}
# Test 1: stage_complete returns elapsed time when msg is NULL
test_that('stage_complete returns elapsed time for NULL msg', {
    result <- stage_complete(tic = 0, toc = 1.5, msg = NULL)
    expect_equal(result, "1.5 seconds elapsed")
})

# Test 2: stage_complete returns elapsed time when msg is NA
test_that('stage_complete returns elapsed time for NA msg', {
    result <- stage_complete(tic = 0, toc = 2.345, msg = NA)
    expect_equal(result, "2.345 seconds elapsed")
})

# Test 3: stage_complete formats message with checkmark and time
test_that('stage_complete formats message with checkmark', {
    result <- stage_complete(tic = 0, toc = 1, msg = "Processing")
    expect_match(result, "\u2705 Processing complete – 1 seconds elapsed")
})

# Test 4: stage_complete rounds elapsed time to 3 decimal places
test_that('stage_complete rounds time to 3 decimals', {
    result <- stage_complete(tic = 0, toc = 1.23456, msg = "Test")
    expect_match(result, "1.235 seconds elapsed")
})
```

---

# Translations and Localization

## Load Translations

```{r function-load_translations}
#' Load translation dictionary
#'
#' @return List containing translations for all supported languages
#' @keywords internal

load_translations <- function() {
    # yaml_file <- system.file("i18n/translations.yaml", package = "rcola")
    yaml_file <- here::here("inst/i18n/translations.yaml")
    rlang::try_fetch(
        yaml::read_yaml(yaml_file),
        error = function(e) cli::cli_abort("Tranlsations not found.")
    )
}
```

```{r tests-load_translations}
test_that('load_translations is a function', {
    expect_true(inherits(load_translations, 'function'))
})

test_that('load_translations works', {
    expect_no_error(dict <- load_translations())
    expect_is(dict, 'list')
    expect_named(dict, c('en', 'uk', 'ru'))
})
```

## Available Languages

```{r function-rcola_available_languages}
#' Get available languages for rcola
#'
#' @return Character vector of supported language codes
#' @export

rcola_available_languages <- function() {
    names(load_translations())
}
```

```{r examples-rcola_available_languages}
rcola_available_languages()
```

```{r tests-rcola_available_languages}
test_that('rcola_available_languages is a function', {
    expect_true(inherits(rcola_available_languages, 'function'))
})

test_that('load_translations works', {
    expect_no_error(rcola_available_languages())
    expect_is(rcola_available_languages(), 'character')
    expect_equal(rcola_available_languages(), c('en', 'uk', 'ru'))
})
```

## Localization Helper

```{r function-l}
#' Get localised string
#'
#' @param language Target language code
#' @param caption_id Translation key
#' @param case Optional second level key (e.g., 'gen', 'dat', 'likert')
#'
#' @return Character string with translation
#' @keywords internal

l <- function(language, caption_id, case = NULL) {
    # Check lanuage
    if (!language %in% rcola_available_languages()) {
        abort(
            "Language '{language}' not available. \\
            Use: {.val {rcola_available_languages()}}"
        )
    }
    
    dict <- load_translations()
    
    # Проверка ключа
    if (!caption_id %in% names(dict[[language]])) {
        cli::cli_abort(
            "Caption ID '{caption_id}' not found ",
            "in language '{language}'"
        )
    }
    
    translation <- dict[[language]][[caption_id]]
    
    if (!is.null(case)) {
        case <- rlang::arg_match(
            case,
            c('gen', 'dat', 'categorical', 'numeric', 'likert'),
            multiple = FALSE
        )
        
        if (!is.list(translation)) {
            cli::cli_abort(
                "Case '{case}' not supported for '{caption_id}' ",
                "(expected list with cases)"
            )
        }
        
        if (!case %in% names(translation)) {
            cli::cli_abort(
                "Case '{case}' not found for '{caption_id}'"
            )
        }
        
        translation[[case]]
    } else {
        if (is.list(translation)) {
            cli::cli_abort(
                "Case required for '{caption_id}'. ",
                "Available: {.val {names(translation)}}"
            )
        }
        
        translation
    }
}
```

```{r tests-l}
# Test 1: l() returns simple string translation without case
test_that('l() returns simple string translation', {
    result <- l(language = "en", caption_id = "analysis")
    expect_equal(result, "Text analysis for")
    expect_type(result, "character")
})

# Test 2: l() returns translation with case parameter
test_that('l() returns translation with case', {
    result <- l(language = "en", caption_id = "object", case = "gen")
    expect_equal(result, "the object")
})

# Test 3: l() rejects invalid language
test_that('l() rejects invalid language', {
    expect_error(
        l(language = "fr", caption_id = "analysis"),
        "Language 'fr' not available"
    )
})

# Test 4: l() rejects invalid caption_id
test_that('l() rejects invalid caption_id', {
    expect_error(
        l(language = "en", caption_id = "nonexistent"),
        "Caption ID 'nonexistent' not found"
    )
})

# Test 5: l() rejects invalid case value
test_that('l() rejects invalid case value', {
    expect_error(
        l(language = "en", caption_id = "object", case = "invalid"),
        "must be one of"
    )
})

# Test 6: l() requires case for dictionary entries with cases
test_that('l() requires case for multi-case entries', {
    expect_error(
        l(language = "en", caption_id = "object"),
        "Case required for 'object'"
    )
})

# Test 7: l() rejects case for simple string entries
test_that('l() rejects case for simple string entries', {
    expect_error(
        l(language = "en", caption_id = "analysis", case = "gen"),
        "Case 'gen' not supported for 'analysis'"
    )
})

# Test 8: l() rejects missing case in multi-case entry
test_that('l() rejects missing case in entry', {
    expect_error(
        l(language = "en", caption_id = "stance_positive", case = "ablative"),
        "must be one of"
    )
})

# Test 9: l() works with different languages
test_that('l() works with Ukrainian translation', {
    result <- l(language = "uk", caption_id = "analysis")
    expect_equal(result, "Аналіз тексту для")
})

# Test 10: l() works with Russian translation
test_that('l() works with Russian translation', {
    result <- l(language = "ru", caption_id = "stage_1")
    expect_equal(result, "Проведение многомерного анализа")
})

# Test 11: l() returns correct case variant
test_that('l() returns correct case variant for different cases', {
    gen <- l(language = "en", caption_id = "statement", case = "gen")
    dat <- l(language = "en", caption_id = "statement", case = "dat")
    expect_equal(gen, "the statement")
    expect_equal(dat, "the following statement:")
    expect_true(gen != dat)
})

# Test 12: l() works with numeric case type
test_that('l() works with numeric case type', {
    result <- l(language = "en", caption_id = "stance_positive", case = "numeric")
    expect_equal(result, "positive, near 1")
})

# Test 13: l() works with likert case type
test_that('l() works with likert case type', {
    result <- l(language = "ru", caption_id = "stance_neutral", case = "likert")
    expect_equal(result, "нейтральная")
})
```

## Check Translations

```{r function-rcola_check_translations}
#' Check translation dictionary structure
#'
#' @description
#' Validates and displays information about the translation dictionary,
#' including available languages, number of keys, and keys with case variants.
#' Useful for verifying translation completeness and structure.
#'
#' @details
#' This function performs the following checks:
#' - Lists all available languages in the dictionary
#' - Counts total translation keys per language
#' - Identifies keys with case variants (e.g., 'gen', 'dat', 'likert')
#'
#' The function prints a formatted report to the console and returns
#' invisibly TRUE.
#'
#' @return
#' Invisibly `TRUE`. Called for side effects (printing to console).
#'
#' @examples
#' \dontrun{
#'   rcola_check_translations()
#'   # Available languages: en, uk, ru
#'   # 
#'   # Language: en
#'   #   Keys: 13
#'   #   Keys with cases: 5
#'   # 
#'   # Language: uk
#'   #   Keys: 13
#'   #   Keys with cases: 5
#'   # 
#'   # Language: ru
#'   #   Keys: 13
#'   #   Keys with cases: 5
#' }
#'
#' @keywords internal
#' @export

rcola_check_translations <- function() {
    dict <- load_translations()
    
    cat("Available languages:", paste(names(dict), collapse = ", "), "\n\n")
    
    for (language in names(dict)) {
        cat("Language:", language, "\n")
        cat("  Keys:", length(dict[[language]]), "\n")
        
        # Проверка структуры
        cases_keys <- sapply(dict[[language]], function(x) is.list(x))
        if (any(cases_keys)) {
            cat("  Keys with cases:", sum(cases_keys), "\n")
        }
        cat("\n")
    }
    
    invisible(TRUE)
}
```

```{r examples-rcola_check_translations}
rcola_check_translations()
```

```{r tests-rcola_check_translations}
test_that('rcola_check_translations returns invisibly TRUE', {
    result <- rcola_check_translations()
    expect_true(result)
    expect_invisible(rcola_check_translations())
})

test_that('rcola_check_translations prints available languages', {
    expect_output(
        rcola_check_translations(),
        "Available languages:"
    )
})

test_that('rcola_check_translations lists all languages', {
    expect_output(
        rcola_check_translations(),
        "en"
    )
    expect_output(
        rcola_check_translations(),
        "uk"
    )
    expect_output(
        rcola_check_translations(),
        "ru"
    )
})

test_that('rcola_check_translations shows key count per language', {
    expect_output(
        rcola_check_translations(),
        "Language: en"
    )
    expect_output(
        rcola_check_translations(),
        "Keys: \\d+"
    )
})

test_that('rcola_check_translations identifies keys with cases', {
    expect_output(
        rcola_check_translations(),
        "Keys with cases:"
    )
})

test_that('rcola_check_translations output is consistent', {
    output1 <- capture.output(rcola_check_translations())
    output2 <- capture.output(rcola_check_translations())
    expect_equal(output1, output2)
})

test_that('rcola_check_translations handles all languages equally', {
    output <- capture.output(rcola_check_translations())
    
    # Проверяем, что каждый язык имеет свою секцию
    en_count <- sum(grepl("Language: en", output))
    uk_count <- sum(grepl("Language: uk", output))
    ru_count <- sum(grepl("Language: ru", output))
    
    expect_equal(en_count, 1)
    expect_equal(uk_count, 1)
    expect_equal(ru_count, 1)
})
```

## Type to Term Conversion

```{r function-type_to_term}
#' Convert analysis type to human-readable term
#'
#' @param type Analysis type: "object" or "statement"
#' @param language Language code
#'
#' @return Character string with localised term
#' @keywords internal

type_to_term <- function(
        type = c('object', 'statement'),
        language = rcola_available_languages()
) {
    type <- rlang::arg_match(type, c('object', 'statement'), multiple = FALSE)
    language <- rlang::arg_match(language, rcola_available_languages(), multiple = FALSE)
    
    switch(
        type,
        object = l(language, 'object', 'dat'),
        statement = l(language, 'statement', 'dat')
    )
}
```

```{r tests-type_to_term}
# Test 1: type_to_term returns correct term for object type
test_that('type_to_term returns object term in English', {
    result <- type_to_term(type = "object", language = "en")
    expect_equal(result, "the object")
})

# Test 2: type_to_term returns correct term for statement type
test_that('type_to_term returns statement term in English', {
    result <- type_to_term(type = "statement", language = "en")
    expect_equal(result, "the following statement:")
})

# Test 3: type_to_term works with Ukrainian language
test_that('type_to_term works with Ukrainian', {
    result <- type_to_term(type = "object", language = "uk")
    expect_equal(result, "об'єкту")
})

# Test 4: type_to_term works with Russian language
test_that('type_to_term works with Russian', {
    result <- type_to_term(type = "statement", language = "ru")
    expect_equal(result, "суждению")
})

# Test 5: type_to_term rejects invalid type
test_that('type_to_term rejects invalid type', {
    expect_error(
        type_to_term(type = "invalid", language = "en"),
        "must be one of"
    )
})

# Test 6: type_to_term rejects invalid language
test_that('type_to_term rejects invalid language', {
    expect_error(
        type_to_term(type = "object", language = "fr"),
        "must be one of"
    )
})

# Test 7: type_to_term uses dative case (dat) for all terms
test_that('type_to_term uses dative case', {
    # Проверяем, что используется именно дательный падеж (dat), а не именительный (gen)
    object_dat <- type_to_term(type = "object", language = "en")
    # Для английского "the object" одинаков, но проверяем логику
    expect_type(object_dat, "character")
    expect_length(object_dat, 1)
})

# Test 8: type_to_term returns character string
test_that('type_to_term returns character string', {
    result <- type_to_term(type = "object", language = "en")
    expect_type(result, "character")
    expect_length(result, 1)
})

# Test 9: type_to_term works with all available languages
test_that('type_to_term works with all available languages', {
    available_langs <- rcola_available_languages()
    
    for (lang in available_langs) {
        result_obj <- type_to_term(type = "object", language = lang)
        result_stmt <- type_to_term(type = "statement", language = lang)
        
        expect_type(result_obj, "character")
        expect_type(result_stmt, "character")
        expect_length(result_obj, 1)
        expect_length(result_stmt, 1)
    }
})

# Test 10: type_to_term returns different terms for different types
test_that('type_to_term returns different terms for different types', {
    object_term <- type_to_term(type = "object", language = "en")
    statement_term <- type_to_term(type = "statement", language = "en")
    
    expect_false(object_term == statement_term)
})
```

---

# Scale System

## Scale Definitions

```{r function-type_stance}
#' Define stance scales for structured data extraction
#'
#' @description
#' These functions define scale types for extracting stance information
#' using `ellmer::parallel_chat_structured()`. They serve as scale definitions
#' for the ellmer structured data extraction tool.
#'
#' @details
#' These functions create type definitions compatible with ellmer's
#' structured extraction framework:
#'
#' - `type_stance_categorical()`: Defines a three-point categorical scale
#'   (Positive, Negative, Neutral)
#' - `type_stance_numeric()`: Defines a continuous numeric scale
#'   (typically -1 to 1)
#' - `type_stance_likert()`: Defines a five-point Likert scale
#'   (Strongly Disagree to Strongly Agree)
#'
#' Each function reads a description from a file using
#' `ellmer::interpolate_file()`, allowing for dynamic content insertion
#' and template rendering.
#'
#' @param description_file Path to a file containing the scale description.
#'   The file may include template variables that will be interpolated
#'   using `ellmer::interpolate_file()`.
#'
#' @return
#' A type definition object compatible with `ellmer::parallel_chat_structured()`.
#' - `type_stance_categorical()` returns an enum type with three values
#' - `type_stance_numeric()` returns a numeric type
#' - `type_stance_likert()` returns an enum type with five values
#'
#' @seealso
#' [ellmer::parallel_chat_structured()] for using these type definitions
#' in structured data extraction tasks.
#'
#' @examples
#' \dontrun{
#'   # Define scale types for stance extraction
#'   categorical_scale <- type_stance_categorical("scales/categorical.txt")
#'   numeric_scale <- type_stance_numeric("scales/numeric.txt")
#'   likert_scale <- type_stance_likert("scales/likert.txt")
#'
#'   # Use in structured extraction
#'   result <- ellmer::parallel_chat_structured(
#'     chat = my_chat,
#'     schema = list(
#'       stance_cat = categorical_scale,
#'       stance_num = numeric_scale,
#'       stance_likert = likert_scale
#'     )
#'   )
#' }
#'
#' @keywords internal
#' @name type_stance

#' @rdname type_stance
type_stance_categorical <- function(description_file) {
  ellmer::type_enum(
      values = c('Positive', 'Negative', 'Neutral'),
      description = ellmer::interpolate_file(description_file)
  )
}

#' @rdname type_stance
type_stance_numeric <- function(description_file) {
  ellmer::type_number(
      description = ellmer::interpolate_file(description_file)
  )
}

#' @rdname type_stance
type_stance_likert <- function(description_file) {
  ellmer::type_enum(
      values = c(
          'Strongly Disagree',
          'Disagree',
          'Neutral',
          'Agree',
          'Strongly Agree'
      ),
      description = ellmer::interpolate_file(description_file)
  )
}
```

```{r tests-type_stance}
prompts_dir <- here::here('inst', 'prompts', 'en')
# prompts_dir <- system.file('prompts', 'en', package = 'rcola')

test_that('type_stance_categorical returns enum type', {
    template_file <- file.path(prompts_dir, 'description-categorical.md')
    skip_if_not(file.exists(template_file))
    result <- type_stance_categorical(template_file)
    expect_is(result, "ellmer::TypeEnum")
})

test_that('type_stance_categorical includes three values', {
    template_file <- file.path(prompts_dir, 'description-categorical.md')
    skip_if_not(file.exists(template_file))
    result <- type_stance_categorical(template_file)
    expect_equal(result@values, c('Positive', 'Negative', 'Neutral'))
})

test_that('type_stance_categorical loads description from file', {
    template_file <- file.path(prompts_dir, 'description-categorical.md')
    skip_if_not(file.exists(template_file))
    result <- type_stance_categorical(template_file)
    expect_type(result@description, "character")
    expect_length(result@description, 1)
    expect_true(nchar(result@description) > 0)
})

test_that('type_stance_numeric returns basic type', {
    template_file <- file.path(prompts_dir, 'description-numeric.md')
    skip_if_not(file.exists(template_file))
    result <- type_stance_numeric(template_file)
    expect_is(result, "ellmer::TypeBasic")
})

test_that('type_stance_numeric loads description from file', {
    template_file <- file.path(prompts_dir, 'description-numeric.md')
    skip_if_not(file.exists(template_file))
    result <- type_stance_numeric(template_file)
    expect_type(result@description, "character")
    expect_length(result@description, 1)
    expect_true(nchar(result@description) > 0)
})

test_that('type_stance_likert returns enum type', {
    template_file <- file.path(prompts_dir, 'description-likert.md')
    skip_if_not(file.exists(template_file))
    result <- type_stance_likert(template_file)
    expect_is(result, "ellmer::TypeEnum")
})

test_that('type_stance_likert includes five values', {
    template_file <- file.path(prompts_dir, 'description-likert.md')
    skip_if_not(file.exists(template_file))
    result <- type_stance_likert(template_file)
    expect_equal(
        result@values,
        c('Strongly Disagree', 'Disagree', 'Neutral', 'Agree', 'Strongly Agree')
    )
})

test_that('type_stance_likert loads description from file', {
    template_file <- file.path(prompts_dir, 'description-likert.md')
    skip_if_not(file.exists(template_file))
    result <- type_stance_likert(template_file)
    expect_type(result@description, "character")
    expect_length(result@description, 1)
    expect_true(nchar(result@description) > 0)
})

test_that('type_stance_categorical and likert have different values', {
    cat_file <- file.path(prompts_dir, 'description-categorical.md')
    likert_file <- file.path(prompts_dir, 'description-likert.md')
    skip_if_not(file.exists(cat_file) && file.exists(likert_file))
    
    cat_result <- type_stance_categorical(cat_file)
    likert_result <- type_stance_likert(likert_file)
    
    expect_length(cat_result@values, 3)
    expect_length(likert_result@values, 5)
    expect_false(cat_result@description == likert_result@description)
})

test_that('type_stance_categorical handles missing file', {
    template_file <- file.path(prompts_dir, 'nonexistent.md')
    skip_if(file.exists(template_file))
    expect_error(
        type_stance_categorical(template_file),
        class = "error"
    )
})

test_that('type_stance_numeric handles missing file', {
    template_file <- file.path(prompts_dir, 'nonexistent.md')
    skip_if(file.exists(template_file))
    expect_error(
        type_stance_numeric(template_file),
        class = "error"
    )
})

test_that('type_stance_likert handles missing file', {
    template_file <- file.path(prompts_dir, 'nonexistent.md')
    skip_if(file.exists(template_file))
    expect_error(
        type_stance_likert(template_file),
        class = "error"
    )
})
```

```{r function-type_stance_analysis}
#' Define analysis stance type
#'
#' @param language Language code
#' @param scale Stance scale type
#'
#' @return List defining analysis stance scale
#' @keywords internal

type_stance_analysis <- function(language, scale) {
# Implementation
}
```

```{r tests-type_stance_analysis}
test_that('type_stance_analysis is a function', {
expect_true(inherits(type_stance_analysis, 'function'))
})
```

---

# Prompt and Task Preparation

## Prompts

```{r function-prompts_prepare}
#' Prepare prompt templates
#'
#' @param role Agent role (linguist, domain, interpreter)
#' @param templates Prompt templates
#' @param language Language code
#' @param add_reference_resolution Add reference resolution
#' @param add_statement_resolution Add statement resolution
#' @param ... Additional arguments
#'
#' @return Prepared prompts
#' @keywords internal

prompts_prepare <- function(
  role,
  templates,
  language = rcola_available_languages(),
  add_reference_resolution = FALSE,
  add_statement_resolution = FALSE,
  ...
) {
# Implementation
}
```

```{r tests-prompts_prepare}
test_that('prompts_prepare is a function', {
expect_true(inherits(prompts_prepare, 'function'))
})
```

## Tasks

```{r function-tasks_validate}
#' Validate task configuration
#'
#' @param tasks Task list
#' @param expected_length Expected number of tasks
#'
#' @return Invisibly TRUE or throws error
#' @keywords internal

tasks_validate <- function(tasks, expected_length) {
# Implementation
}
```

```{r tests-tasks_validate}
test_that('tasks_validate is a function', {
expect_true(inherits(tasks_validate, 'function'))
})
```

```{r function-tasks_prepare}
#' Prepare tasks for execution
#'
#' @param chat_base Base chat configuration
#' @param prompts Prepared prompts
#' @param n_texts Number of texts
#'
#' @return Prepared tasks
#' @keywords internal

tasks_prepare <- function(chat_base, prompts, n_texts) {
# Implementation
}
```

```{r tests-tasks_prepare}
test_that('tasks_prepare is a function', {
expect_true(inherits(tasks_prepare, 'function'))
})
```

---

# Expert Analysis

## Prepare Expert Chats

```{r function-prepare_expert_chats}
#' Prepare chat configurations for expert agents
#'
#' @param inputs Input texts
#' @param chat_base Base chat configuration
#' @param expert_role Expert role (linguist, domain, interpreter)
#'
#' @return List of chat configurations
#' @keywords internal

prepare_expert_chats <- function(
  inputs,
  chat_base,
  expert_role = c("linguist", "domain", "interpreter")
) {
# Implementation
}
```

```{r tests-prepare_expert_chats}
test_that('prepare_expert_chats is a function', {
expect_true(inherits(prepare_expert_chats, 'function'))
})
```

## Execute Role

```{r function-execute_role}
#' Execute expert role analysis
#'
#' @param expert_role Expert role
#' @param inputs Input texts
#' @param chat_base Chat configuration
#' @param info Additional information
#' @param rpm Requests per minute limit
#' @param verbose Show progress messages
#'
#' @return Analysis results
#' @keywords internal

execute_role <- function(
  expert_role,
  inputs,
  chat_base,
  info,
  rpm = 500,
  verbose = TRUE
) {
# Implementation
}
```

```{r tests-execute_role}
test_that('execute_role is a function', {
expect_true(inherits(execute_role, 'function'))
})
```

---

# Debate Stage

## Prepare Debater Chats

```{r function-prepare_debater_chats}
#' Prepare chat configurations for debate agents
#'
#' @param stance Initial stance analysis
#' @param inputs Input texts
#' @param chat_base Base chat configuration
#'
#' @return List of chat configurations
#' @keywords internal

prepare_debater_chats <- function(stance, inputs, chat_base) {
# Implementation
}
```

```{r tests-prepare_debater_chats}
test_that('prepare_debater_chats is a function', {
expect_true(inherits(prepare_debater_chats, 'function'))
})
```

---

# Judgment Stage

## Prepare Judger Chats

```{r function-prepare_judger_chats}
#' Prepare chat configurations for judge agent
#'
#' @param inputs Input texts
#' @param chat_base Base chat configuration
#'
#' @return List of chat configurations
#' @keywords internal

prepare_judger_chats <- function(inputs, chat_base) {
# Implementation
}
```

```{r tests-prepare_judger_chats}
test_that('prepare_judger_chats is a function', {
expect_true(inherits(prepare_judger_chats, 'function'))
})
```

---

# Pipeline Stages

## Stage 1: Parallel Analysis

```{r function-stage_1_parallel_analysis}
#' Execute Stage 1: Parallel expert analysis
#'
#' @param inputs Input texts
#' @param chat_base Chat configuration
#' @param verbose Show progress
#' @param rpm Requests per minute limit
#'
#' @return Analysis results
#' @keywords internal

stage_1_parallel_analysis <- function(
  inputs,
  chat_base,
  verbose = TRUE,
  rpm = 500
) {
# Implementation
}
```

```{r tests-stage_1_parallel_analysis}
test_that('stage_1_parallel_analysis is a function', {
expect_true(inherits(stage_1_parallel_analysis, 'function'))
})
```

## Stage 2: Parallel Debates

```{r function-stage_2_parallel_debates}
#' Execute Stage 2: Parallel agent debates
#'
#' @param inputs Input texts
#' @param chat_base Chat configuration
#' @param verbose Show progress
#' @param rpm Requests per minute limit
#'
#' @return Debate results
#' @keywords internal

stage_2_parallel_debates <- function(
  inputs,
  chat_base,
  verbose = TRUE,
  rpm = 500
) {
# Implementation
}
```

```{r tests-stage_2_parallel_debates}
test_that('stage_2_parallel_debates is a function', {
expect_true(inherits(stage_2_parallel_debates, 'function'))
})
```

## Stage 3: Parallel Judgment

```{r function-stage_3_parallel_judgment}
#' Execute Stage 3: Judge synthesis
#'
#' @param inputs Input texts
#' @param chat_base Chat configuration
#' @param verbose Show progress
#' @param rpm Requests per minute limit
#' @param ... Additional arguments
#'
#' @return Final stance judgments
#' @keywords internal

stage_3_parallel_judgment <- function(
  inputs,
  chat_base,
  verbose = TRUE,
  rpm = 500,
  ...
) {
# Implementation
}
```

```{r tests-stage_3_parallel_judgment}
test_that('stage_3_parallel_judgment is a function', {
expect_true(inherits(stage_3_parallel_judgment, 'function'))
})
```

---

# Main Analysis Function

## LLM Stance Detection

```{r function-llm_stance}
#' Detect stance using COLA framework
#'
#' @param text Input text(s)
#' @param ... Additional arguments passed to methods
#'
#' @return stance_result object
#' @export

llm_stance <- function(text, ...) {
UseMethod("llm_stance")
}
```

```{r function-llm_stance.character}
#' Stance detection for character vectors
#'
#' @param text Character vector of texts
#' @param target Target entity or statement
#' @param chat_base Chat model configuration
#' @param type Analysis type: "object" or "statement"
#' @param language Language code
#' @param scale Stance scale: "categorical", "numeric", or "likert"
#' @param domain_role Optional domain-specific role
#' @param prompts_dir Optional custom prompts directory
#' @param verbose Show progress messages
#' @param rpm Requests per minute limit
#' @param ... Additional arguments
#'
#' @return stance_result object
#' @export

llm_stance.character <- function(
  text,
  target,
  chat_base,
  type = c("object", "statement"),
  language = rcola_available_languages(),
  scale = c("categorical", "numeric", "likert"),
  domain_role = NULL,
  prompts_dir = NULL,
  verbose = TRUE,
  rpm = 20,
  ...
) {
# Implementation
}
```

```{r examples-llm_stance.character}
# Example usage would go here
# result <- llm_stance(
#   text = "I support renewable energy",
#   target = "renewable energy",
#   chat_base = chat_model,
#   type = "object",
#   scale = "categorical"
# )
```

```{r tests-llm_stance.character}
test_that('llm_stance.character is a function', {
expect_true(inherits(llm_stance.character, 'function'))
})
```

```{r function-llm_stance.data.frame}
#' Stance detection for data frames
#'
#' @param data Input data frame
#' @param text Column name containing texts
#' @param target Column name or vector of targets
#' @param type Analysis type
#' @param domain_role Optional domain role
#' @param language Language code
#' @param chat_base Chat model configuration
#' @param scale Stance scale
#' @param prompts_dir Optional custom prompts directory
#' @param verbose Show progress
#' @param rpm Requests per minute limit
#' @param .output_col Name of output column
#' @param ... Additional arguments
#'
#' @return Data frame with stance results
#' @export

llm_stance.data.frame <- function(
  data,
  text,
  target,
  type = NULL,
  domain_role = NULL,
  language = NULL,
  chat_base,
  scale = "categorical",
  prompts_dir = NULL,
  verbose = TRUE,
  rpm = 20,
  .output_col = ".stance",
  ...
) {
# Implementation
}
```

```{r examples-llm_stance.data.frame}
# Example usage would go here
# results <- llm_stance(
#   data = tweets_df,
#   text = "text",
#   target = "climate_change",
#   chat_base = chat_model
# )
```

```{r tests-llm_stance.data.frame}
test_that('llm_stance.data.frame is a function', {
expect_true(inherits(llm_stance.data.frame, 'function'))
})
```

---

# Result Inspection and Display

## Print Method

```{r function-print.stance_result}
#' Print stance_result object
#'
#' @param x stance_result object
#' @param ... Additional arguments
#'
#' @return Invisibly x
#' @export

print.stance_result <- function(x, ...) {
# Implementation
}
```

```{r examples-print.stance_result}
# result <- llm_stance(...)
# print(result)
```

```{r tests-print.stance_result}
test_that('print.stance_result is a function', {
expect_true(inherits(print.stance_result, 'function'))
})
```

## Summary Method

```{r function-summary.stance_result}
#' Summarize stance_result object
#'
#' @param object stance_result object
#' @param ... Additional arguments
#'
#' @return Invisibly object
#' @export

summary.stance_result <- function(object, ...) {
# Implementation
}
```

```{r examples-summary.stance_result}
# result <- llm_stance(...)
# summary(result)
```

```{r tests-summary.stance_result}
test_that('summary.stance_result is a function', {
expect_true(inherits(summary.stance_result, 'function'))
})
```

## Data Frame Conversion

```{r function-as.data.frame.stance_result}
#' Convert stance_result to data frame
#'
#' @param x stance_result object
#' @param row.names Optional row names
#' @param optional Logical
#' @param ... Additional arguments
#'
#' @return Data frame
#' @export

as.data.frame.stance_result <- function(
  x,
  row.names = NULL,
  optional = FALSE,
  ...
) {
# Implementation
}
```

```{r examples-as.data.frame.stance_result}
# result <- llm_stance(...)
# df <- as.data.frame(result)
```

```{r tests-as.data.frame.stance_result}
test_that('as.data.frame.stance_result is a function', {
expect_true(inherits(as.data.frame.stance_result, 'function'))
})
```

## Inspect Method

```{r function-inspect}
#' Inspect stance_result contents
#'
#' @param x stance_result object
#' @param ... Additional arguments passed to methods
#'
#' @return Invisibly x
#' @export

inspect <- function(x, ...) {
UseMethod("inspect")
}
```

```{r function-inspect.stance_result}
#' Inspect stance_result object details
#'
#' @param x stance_result object
#' @param what Primary inspection level: 'metadata', 'analysis', or 'debates'
#' @param within Secondary inspection level
#' @param index Row index to inspect
#'
#' @return Invisibly x
#' @export

inspect.stance_result <- function(
  x,
  what = c('metadata', 'analysis', 'debates'),
  within = NULL,
  index = 1
) {
# Implementation
}
```

```{r examples-inspect.stance_result}
# result <- llm_stance(...)
# inspect(result, what = 'metadata')
# inspect(result, what = 'analysis', within = 'linguistic_analysis')
# inspect(result, what = 'debates', within = 'counterarguments', index = 1)
```

```{r tests-inspect.stance_result}
test_that('inspect.stance_result is a function', {
expect_true(inherits(inspect.stance_result, 'function'))
})
```

---

# Inflate

```{r development-inflate, eval=FALSE}
fusen::inflate(
flat_file = 'dev/flat_rcola.Rmd',
codecov = TRUE,
clean = TRUE
)
```
