---
title: 'stancer: Stance Analysis using Ensemble of LLM Agents through `ellmer`'
output: html_document
editor_options:
chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
library(withr)
library(mockery)
library(rlang)
library(cli)
library(ellmer)
library(yaml)
```

```{r description}
fusen::fill_description(
  pkg = here::here(),
  fields = list(
    package = "stancer",
    Title = "Stance Analysis using Ensemble of LLM Agents through `ellmer`",
    Description = "Tools for stance analysis using an ensemble of large language models. The package follows a multi-stage process in which different agents perform linguistic analysis, role-play as domain experts, and engage in structured debates to reach a final judgment. It works with 'ellmer' chat objects, so a user can choose which models and providers to use. Each result includes the full reasoning and intermediate steps from the agents.",
    `Authors@R` = c(
      person("Andrej", "Pawluczenko", email = "a.pawluczenko@gmail.com", role = c("aut", "cre"))
    )
  ),
  overwrite = TRUE
)
# Define License with use_*_license()
usethis::use_ccby_license()

usethis::use_package('ellmer')
usethis::use_package('yaml')
usethis::use_package('glue')
usethis::use_package('cli')
usethis::use_package('rlang')
usethis::use_package('tibble')
usethis::use_package('cld2', type = "Suggests")

usethis::reso
```

# Introduction

## Overview

---

# Utility Functions

## Text Truncation

```{r function-truncate}
#' Truncate numeric values to a range
#'
#' Constrains numeric values to fall within the range [a, b].
#' Values below a are set to a, values above b are set to b.
#'
#' @param x Numeric vector to truncate
#' @param a Numeric. Lower bound (minimum value)
#' @param b Numeric. Upper bound (maximum value)
#'
#' @return Numeric vector with values constrained to [a, b]
#' @keywords internal
#'
#' @examples
#' truncate(c(-1, 0, 5, 10), a = 0, b = 5)
#' # Returns: 0, 0, 5, 5
#' @filename utils

truncate <- function(x, a, b) {
  if (!is.numeric(x)) {
    cli::cli_abort("{.arg x} must be numeric")
  }
  if (!is.numeric(a)) {
    cli::cli_abort("{.arg a} must be numeric")
  }
  if (!is.numeric(b)) {
    cli::cli_abort("{.arg b} must be numeric")
  }

  pmin(pmax(x, a), b)
}
```

## Expression Catching

```{r function-catch}
#' Safely evaluate expression and catch errors
#'
#' @param expr Expression to evaluate
#' @param expr_name Name of expression for error messages
#'
#' @return Result of expression or throws an error
#' @keywords internal
#' @importFrom rlang is_scalar_character try_fetch
#' @importFrom cli cli_abort
#' @filename utils

catch <- function(
    expr,
    expr_name = paste(deparse(substitute(expr)), collapse = '')
) {
  if (!rlang::is_scalar_character(expr_name)) {
    cli::cli_abort(
      "{.arg expr_name} must be a character scalar.
            Got {.cls {class(expr_name)}}"
    )
  }

  rlang::try_fetch(
    expr,
    error = function(e) {
      abort("Error in {expr_name}: {e$message}")
    }
  )
}
```

```{r function-catch_eval}
#' Safely evaluate quosures (throws an error)
#'
#' @param quo A quosure to evaluate
#' @param data A data frame, or named list or vector.
#'
#' @return Result of evaluation or throws an error
#' @keywords internal
#' @importFrom rlang is_quosure try_fetch as_name
#' @importFrom cli cli_abort

catch_eval <- function(
    quo,
    data
) {
  if (!rlang::is_quosure(quo)) {
    cli::cli_abort(
      "{.arg quo} must be a quosure.
            Got {.cls {class(quo)}}"
    )
  }

  rlang::try_fetch(
    rlang::eval_tidy(quo, data),
    error = function(e) {
      abort(
        c(
          "Cannot find {.val {rlang::as_name(quo)}} in data or environment",
          "i" = "Use a column name or provide a vector"
        )
      )
    }
  )
}
```

```{r function-safe_eval}
#' Safely evaluate quosures (returns NULL)
#'
#' @param quo A quosure to evaluate
#' @param data A data frame, or named list or vector.
#'
#' @return Result of evaluation or `NULL`
#' @keywords internal
#' @importFrom rlang try_fetch
#' @filename utils

safe_eval <- function(
    quo,
    data
) {
  rlang::try_fetch(
    catch_eval(quo, data),
    error = function(e) NULL
  )
}
```

## Input Validation

```{r function-validate_fields}
#' Validate input structure and required fields
#'
#' @description
#' An internal helper to ensure that a list (typically an input or result
#' object) contains all the necessary fields required for a specific
#' analysis stage.
#'
#' @param inputs The list or named vector to validate.
#' @param required_fields Character vector of the field names that must
#'   be present in `inputs`.
#' @param input_name Character scalar. A descriptive name for the
#'   object being validated, used in error messages.
#'
#' @return Invisibly TRUE if all required fields are present or throws an error if `inputs` is NULL, if any required fields are missing, or if `input_name` is not a character scalar.
#'
#' @keywords internal
#' @importFrom rlang is_scalar_character
#' @importFrom cli cli_abort
#' @filename validate

validate_fields <- function(inputs, required_fields, input_name = 'analysis') {
  if (!rlang::is_scalar_character(input_name)) {
    cli::cli_abort("{.arg input_name} must be a character scalar.")
  }

  if (is.null(inputs)) {
    abort("Input for {input_name} is empty")
  }

  missing_fields <- setdiff(required_fields, names(inputs))
  if (length(missing_fields) > 0) {
    abort(
      "Input for {input_name} misses required fields: {.var {missing_fields}}"
    )
  }

  invisible(TRUE)
}
```

```{r function-validate_inputs}
#' Validate input length
#'
#' @param inputs Input vector
#' @param expected_length Expected length
#' @param input_name Informative name for error messages
#'
#' @return Invisibly TRUE or throws error
#' @keywords internal
#' @importFrom rlang is_scalar_character
#' @importFrom cli cli_abort
#' @filename validate

validate_inputs <- function(inputs, expected_length, input_name = 'Inputs') {
  if (!rlang::is_scalar_character(input_name)) {
    cli::cli_abort("{.arg input_name} must be a character scalar.")
  }

  if (is.null(inputs) || length(inputs) != expected_length) {
    abort("{input_name} returned unexpected results")
  }

  invisible(TRUE)
}
```

```{r function-validate_character}
#' Validate a character argument
#'
#' @param x Input to validate
#'
#' @return Invisibly TRUE or throws error
#' @keywords internal
#' @importFrom rlang is_character
#' @filename validate

validate_character <- function(x) {
  arg_name <- deparse(substitute(x))

  if (!rlang::is_character(x)) {
    abort(
      "{.arg {arg_name}} must be a character vector, got {.cls {class(x)}}"
    )
  }
  if (length(x) == 0) {
    abort("{.arg {arg_name}} cannot be empty")
  }
  if (any(is.na(x)) || any(ellmer:::trim(x) == '')) {
    abort(
      c(
        "{.arg {arg_name}} contains empty strings.",
        "i" = "consider removing them before analysis."
      )
    )
  }

  invisible(TRUE)
}
```

```{r function-validate_stage}
#' Validate stage configuration
#'
#' @param x Stage configuration
#' @param stage_name Name of stage for error messages
#'
#' @return Unmodified `x` or throws an error
#' @keywords internal
#' @filename validate

validate_stage <- function(x, stage_name) {
  if (is.null(x)) {
    abort("{stage_name} returned NULL")
  }
  x
}
```

## Argument Recycling

```{r function-recycle_arg}
#' Recycle argument to specified length
#'
#' @param arg Values to recycle
#' @param n Target length
#'
#' @return Recycled argument values
#' @keywords internal
#' @filename utils

recycle_arg <- function(arg, n) {
  arg_name <- deparse(substitute(arg))

  if (length(arg) == 1) {
    return(rep(arg, n))
  } else if (length(arg) == n) {
    return(arg)
  } else {
    abort(
      "{.arg {arg_name}} must have length 1 or {n} (same as {.arg text}). \\
            Got {length(arg)}."
    )
  }
}
```

## Timing

```{r function-stage_complete}
#' Format timing message
#'
#' @param tic Start time
#' @param toc End time
#' @param msg Message prefix
#'
#' @return Formatted timing message
#' @keywords internal
#' @importFrom glue glue
#' @filename verbose

stage_complete <- function(tic, toc, msg) {
  tocmsg <- paste0(round(toc - tic, 3), " seconds elapsed")

  if (is.null(msg) || is.na(msg) || length(msg) == 0) {
    outmsg <- tocmsg
  } else {
    outmsg <- glue::glue('\u2705 {msg} complete \U2014 {tocmsg}\n\n')
  }
  outmsg
}
```

---

# Translations and Localization

## Load Translations

```{r function-load_translations}
#' Load translation dictionary
#'
#' @return List containing translations for all supported languages
#' @keywords internal
#' @filename localisation

load_translations <- function() {
  yaml_file <- system.file("i18n/translations.yaml", package = "stancer")
  rlang::try_fetch(
    yaml::read_yaml(yaml_file),
    error = function(e) cli::cli_abort("Tranlsations not found.")
  )
}
```

## Available Languages

```{r function-stancer_available_languages}
#' Get available languages for stancer
#'
#' @return Character vector of supported language codes
#' @filename localisation
#' @export

stancer_available_languages <- function() {
  names(load_translations())
}
```

```{r examples-stancer_available_languages}
stancer_available_languages()
```

## Localisation Helper

```{r function-l}
#' Get localised string
#'
#' @param language Target language code
#' @param caption_id Translation key
#' @param case Optional second level key (e.g., 'gen', 'dat', 'likert')
#'
#' @return Character string with translation
#' @keywords internal
#' @filename localisation

l <- function(language, caption_id, case = NULL) {
  # Check lanuage
  if (!language %in% stancer_available_languages()) {
    abort(
      "Language '{language}' not available. \\
            Use: {.val {stancer_available_languages()}}"
    )
  }

  dict <- load_translations()

  # Проверка ключа
  if (!caption_id %in% names(dict[[language]])) {
    cli::cli_abort(
      "Caption ID '{caption_id}' not found ",
      "in language '{language}'"
    )
  }

  translation <- dict[[language]][[caption_id]]

  if (!is.null(case)) {
    case <- rlang::arg_match(
      case,
      c('gen', 'dat', 'categorical', 'numeric', 'likert'),
      multiple = FALSE
    )

    if (!is.list(translation)) {
      cli::cli_abort(
        "Case '{case}' not supported for '{caption_id}' ",
        "(expected list with cases)"
      )
    }

    if (!case %in% names(translation)) {
      cli::cli_abort(
        "Case '{case}' not found for '{caption_id}'"
      )
    }

    translation[[case]]
  } else {
    if (is.list(translation)) {
      cli::cli_abort(
        "Case required for '{caption_id}'. ",
        "Available: {.val {names(translation)}}"
      )
    }

    translation
  }
}
```

## Check Translations

```{r function-stancer_check_translations}
#' Check translation dictionary structure
#'
#' @description
#' Validates and displays information about the translation dictionary,
#' including available languages, number of keys, and keys with case variants.
#' Useful for verifying translation completeness and structure.
#'
#' @details
#' This function performs the following checks:
#' - Lists all available languages in the dictionary
#' - Counts total translation keys per language
#' - Identifies keys with case variants (e.g., 'gen', 'dat', 'likert')
#'
#' The function prints a formatted report to the console and returns
#' invisibly TRUE.
#'
#' @return
#' Invisibly `TRUE`. Called for side effects (printing to console).
#'
#' @keywords internal
#' @filename localisation
#' @export

stancer_check_translations <- function() {
  dict <- load_translations()

  cat("Available languages:", paste(names(dict), collapse = ", "), "\n\n")

  for (language in names(dict)) {
    cat("Language:", language, "\n")
    cat("  Keys:", length(dict[[language]]), "\n")

    # Проверка структуры
    cases_keys <- sapply(dict[[language]], function(x) is.list(x))
    if (any(cases_keys)) {
      cat("  Keys with cases:", sum(cases_keys), "\n")
    }
    cat("\n")
  }

  invisible(TRUE)
}
```

```{r examples-stancer_check_translations}
stancer_check_translations()
```

## Type to Term Conversion

```{r function-type_to_term}
#' Convert analysis type to human-readable term
#'
#' @param type Analysis type: "object" or "claim"
#' @param language Language code
#'
#' @return Character string with localised term
#' @keywords internal
#' @filename localisation

type_to_term <- function(
    type = c('object', 'claim'),
    language = stancer_available_languages()
) {
  type <- rlang::arg_match(type, c('object', 'claim'), multiple = FALSE)
  language <- rlang::arg_match(language, stancer_available_languages(), multiple = FALSE)

  switch(
    type,
    object = l(language, 'object', 'dat'),
    claim = l(language, 'claim', 'dat')
  )
}
```

---

# Scale System

## Scale Definitions

```{r function-type_stance}
#' Define stance scales for structured data extraction
#'
#' @description
#' These functions define scale types for extracting stance information
#' using `ellmer::parallel_chat_structured()`. They serve as scale definitions
#' for the ellmer structured data extraction tool.
#'
#' @details
#' These functions create type definitions compatible with ellmer's
#' structured extraction framework:
#'
#' - `type_stance_categorical()`: Defines a three-point categorical scale
#'   (Positive, Negative, Neutral)
#' - `type_stance_numeric()`: Defines a continuous numeric scale
#'   (typically -1 to 1)
#' - `type_stance_likert()`: Defines a five-point Likert scale
#'   (Strongly Disagree to Strongly Agree)
#'
#' Each function reads a description from a file using
#' `ellmer::interpolate_file()`, allowing for dynamic content insertion
#' and template rendering.
#'
#' @param description_file Path to a file containing the scale description.
#'   The file may include template variables that will be interpolated
#'   using `ellmer::interpolate_file()`.
#'
#' @return
#' A type definition object compatible with `ellmer::parallel_chat_structured()`.
#' - `type_stance_categorical()` returns an enum type with three values
#' - `type_stance_numeric()` returns a numeric type
#' - `type_stance_likert()` returns an enum type with five values
#'
#' @seealso
#' [ellmer::parallel_chat_structured()] for using these type definitions
#' in structured data extraction tasks.
#'
#'
#' @keywords internal
#' @filename scales

#' @rdname type_stance
type_stance_categorical <- function(description_file) {
  ellmer::type_enum(
    values = c('Negative', 'Neutral', 'Positive'),
    description = ellmer::interpolate_file(description_file)
  )
}

#' @rdname type_stance
type_stance_numeric <- function(description_file) {
  ellmer::type_number(
    description = ellmer::interpolate_file(description_file)
  )
}

#' @rdname type_stance
type_stance_likert <- function(description_file) {
  ellmer::type_enum(
    values = c(
      'Strongly Disagree',
      'Disagree',
      'Neutral',
      'Agree',
      'Strongly Agree'
    ),
    description = ellmer::interpolate_file(description_file)
  )
}
```

## Data Schema

```{r function-type_stance_analysis}
#' Define stance analysis schema for structured extraction
#'
#' @description
#' Creates a complete schema for stance analysis that combines a stance scale
#' with an explanation field. This schema is designed for use with
#' `ellmer::parallel_chat_structured()` to extract structured stance analysis results.
#'
#' @details
#' This function creates a composite type object containing:
#' - **stance:** A scale type (categorical, numeric, or Likert) for the stance value
#' - **explanation:** A string field for the stance classification explanation
#'
#' The function supports three scale types:
#' - `"numeric"`: Continuous scale (typically -1 to 1)
#' - `"likert"`: Five-point Likert scale (Strongly Disagree to Strongly Agree)
#' - `"categorical"`: Three-point categorical scale (Positive, Negative, Neutral)
#'
#' The explanation field description is automatically localised based on the
#' provided language code using the `'explanation_description'` translation key.
#'
#' @param language Language code for localisation (e.g., "en", "uk", "ru").
#'   Must be one of the available languages in the stancer translation dictionary.
#'
#' @param description_file Path to a file containing the stance scale description.
#'   The file may include template variables that will be interpolated
#'   using `ellmer::interpolate_file()`.
#'
#' @param scale Type of stance scale to use. One of:
#'   - `"numeric"`: Continuous numeric scale
#'   - `"likert"`: Five-point Likert scale
#'   - `"categorical"`: Three-point categorical scale (default)
#'
#' @return
#' A type object compatible with `ellmer::parallel_chat_structured()`.
#' The returned object is a composite type containing:
#' - `stance`: A scale type definition
#' - `explanation`: A string type definition with localised description
#'
#' @seealso
#' [type_stance_categorical()], [type_stance_numeric()], [type_stance_likert()]
#' for individual scale definitions.
#'
#' [ellmer::parallel_chat_structured()] for using this schema in
#' structured data extraction tasks.
#'
#' [l()] for translation lookup.
#'
#'
#' @keywords internal
#' @filename scales
type_stance_analysis <- function(language, description_file, scale = "categorical") {
  # Validate scale parameter
  scale <- rlang::arg_match(scale, c("numeric", "likert", "categorical"))

  # A schema for the structured output
  type_stance <- switch(
    scale,
    numeric = type_stance_numeric(description_file),
    likert = type_stance_likert(description_file),
    type_stance_categorical(description_file)
  )

  ellmer::type_object(
    stance = type_stance,
    explanation = ellmer::type_string(
      description = l(language, 'explanation_description')
    )
  )
}
```

---

# Prompt and Task Preparation

## Prompts

```{r function-templates_collect}
#' Collect prompt template file paths
#'
#' @description
#' Scans specified and default directories to locate Markdown template files
#' for a given language and scale type. It follows a priority order: user-defined
#' directory first, followed by the package's internal prompt directories.
#'
#' @param prompts_dir Character string. Path to a custom directory containing
#'   prompt templates. Can be `NULL`.
#' @param language Character string. Language code (e.g., "en", "uk").
#' @param scale Character string. The rating scale type: "categorical",
#'   "numeric", or "likert".
#'
#' @return A named list where keys represent template roles (e.g.,
#'   `system-linguist`, `scale_description`) and values are absolute file paths.
#'
#' @keywords internal
#' @filename prompting
templates_collect <- function(prompts_dir, language, scale) {
  # 1. Determine default package directories
  DEFAULT_PROMPTS_DIR <- system.file(
    file.path('prompts', language),
    package = 'stancer'
  )
  # Fall-back for development environment
  # if (!dir.exists(DEFAULT_PROMPTS_DIR)) {
  #   DEFAULT_PROMPTS_DIR <- file.path('inst', 'prompts', language)
  # }

  # 2. Filter existing directories
  search_dirs <- c(prompts_dir, DEFAULT_PROMPTS_DIR) |>
    Filter(f = \(x) !is.null(x) && dir.exists(x))

  if (length(search_dirs) == 0) {
    cli::cli_abort(
      "No prompt directories found for language {.val {language}}"
    )
  }

  # 3. Define the set of required role templates
  role_templates <- expand.grid(
    call = c('system', 'user'),
    role = c('linguist', 'domain', 'interpreter', 'debater', 'judger'),
    stringsAsFactors = FALSE
  ) |>
    glue::glue_data('{call}-{role}.md') |>
    as.character() |>
    as.list()

  # Clean names (e.g., "system-linguist.md" -> "system-linguist")
  names(role_templates) <- gsub('\\.md', '', role_templates)

  # 4. Combine with scale-specific and helper files
  # Note: scale is now 'categorical', 'numeric', or 'likert'
  prompt_files <- list(
    scale_description = glue::glue('description-{scale}.md'),
    scale_guide = glue::glue('guide-{scale}.md'),
    reference_resolution = 'reference-resolution.md',
    claim_resolution = 'claim-resolution.md'
  ) |>
    c(role_templates)

  prompt_templates <- list()

  # 5. Resolve paths with priority
  for (name in names(prompt_files)) {
    filename <- prompt_files[[name]]

    found <- FALSE
    for (dir in search_dirs) {
      filepath <- file.path(dir, filename)
      if (file.exists(filepath)) {
        prompt_templates[[name]] <- filepath
        found <- TRUE
        break
      }
    }

    if (!found) {
      cli::cli_abort(
        "Prompt {.file {filename}} not found for language {.val {language}}"
      )
    }
  }

  prompt_templates
}
```

```{r function-prompts_prepare}
#' Prepare system and user prompts for a given role
#'
#' @description
#' Loads and interpolates system and user prompt templates for a specific role
#' in the stance analysis workflow. This function prepares the prompts that
#' guide the language model's behaviour for different analysis stages.
#'
#' @details
#' The function supports five distinct roles in the stance analysis pipeline:
#'
#' - **linguist**: Performs linguistic analysis of statements
#' - **domain**: Provides domain-specific expertise and context
#' - **interpreter**: Interprets social media content
#' - **debater**: Presents arguments and counter-arguments
#' - **judger**: Makes final stance judgements (supports additional options)
#'
#' For the `judger` role, additional instructions can be included:
#' - `add_reference_resolution`: Include instructions for resolving references
#' - `add_claim_resolution`: Include instructions for resolving propositions
#'
#' The function performs the following operations:
#' 1. Validates the role and language parameters
#' 2. Loads template files for the specified role
#' 3. Interpolates templates with provided variables (via `...`)
#' 4. Validates that prompts are non-empty after interpolation
#'
#' Templates are loaded using `ellmer::interpolate_file()`, which allows
#' for dynamic content insertion and template variable rendering.
#'
#' @param role Role for which to prepare prompts. One of:
#'   - `"linguist"`: Linguistic analysis
#'   - `"domain"`: Domain expertise
#'   - `"interpreter"`: Text interpretation
#'   - `"debater"`: Argument presentation
#'   - `"judger"`: Final judgement
#'
#' @param templates A list containing template file paths. Expected keys:
#'   - `system-{role}`: Path to system prompt template
#'   - `user-{role}`: Path to user prompt template
#'   - `reference_resolution`: Path to reference resolution instructions (for judger)
#'   - `claim_resolution`: Path to claim resolution instructions (for judger)
#'   - `scale_guide`: Path to scale guide (for judger)
#'
#' @param language Language code for localisation (e.g., "en", "uk", "ru").
#'   Defaults to the first available language. Must be one of the available
#'   languages in the stancer translation dictionary.
#'
#' @param add_reference_resolution Logical. For the `judger` role only.
#'   If `TRUE`, includes instructions for resolving references in texts.
#'   Defaults to `FALSE`.
#'
#' @param add_claim_resolution Logical. For the `judger` role only.
#'   If `TRUE`, includes instructions for resolving propositions.
#'   Defaults to `FALSE`.
#'
#' @param ... Additional named arguments passed to `ellmer::interpolate_file()`.
#'   These allow for template variable substitution (e.g., `target = "climate change"`).
#'
#' @return
#' A list with two elements:
#' - `system`: Character string containing the system prompt
#' - `task`: Character string containing the user prompt
#'
#' Both strings are non-empty after interpolation.
#'
#' @seealso
#' [ellmer::interpolate_file()] for template interpolation details.
#'
#' @keywords internal
#' @filename prompting
prompts_prepare <- function(
    role,
    templates,
    language = stancer_available_languages(),
    add_reference_resolution = FALSE,
    add_claim_resolution = FALSE,
    ...
) {
  role <- rlang::arg_match(
    role,
    values = c(
      'linguist', 'domain', 'interpreter', 'debater', 'judger'
    ),
    multiple = FALSE
  )
  language <- rlang::arg_match(
    language,
    stancer_available_languages(),
    multiple = FALSE
  )

  # A lot of work for a judger
  if (role == 'judger') {
    if (add_reference_resolution) {
      reference_instruction <- ellmer::interpolate_file(
        templates$reference_resolution
      )
    } else {
      reference_instruction <- ''
    }

    if (add_claim_resolution) {
      claim_instruction <- ellmer::interpolate_file(
        templates$claim_resolution
      )
    } else {
      # Default settings
      claim_instruction <- ''
    }

    scale_guide <- ellmer::interpolate_file(templates$scale_guide)
  }

  template_system <- templates[[glue::glue('system-{role}')]]
  template_user <- templates[[glue::glue('user-{role}')]]

  # Check for file existence
  if (!file.exists(template_system)) {
    abort(
      c(
        "System template not found: {.file {template_system}}",
        "i" = "Available roles: {.val linguist}, {.val domain}, {.val interpreter}, {.val debater}, {.val judger}"
      )
    )
  }
  if (!file.exists(template_user)) {
    abort(
      c(
        "User template not found: {.file {template_user}}",
        "i" = "Available roles: {.val linguist}, {.val domain}, {.val interpreter}, {.val debater}, {.val judger}"
      )
    )
  }

  prompts <- list(
    system = ellmer::interpolate_file(template_system, ...),
    task = ellmer::interpolate_file(template_user, ...)
  )

  # Check for user prompt existence
  if (is.null(prompts$task)) {
    abort("User prompt is NULL for role {.val {role}}")
  }

  # Check for empty strings
  if (any(nchar(prompts$system) == 0)) {
    abort("System prompt is empty after interpolation for role {.val {role}}")
  }
  if (any(nchar(prompts$task) == 0)) {
    abort("User prompt is empty after interpolation for role {.val {role}}")
  }

  prompts
}
```

```{r examples-prompts-prepare}
# Prepare prompts for linguistic analysis
# templates <- list(
#   `system-linguist` = "prompts/en/system-linguist.md",
#   `user-linguist` = "prompts/en/user-linguist.md"
# )
# 
# prompts <- stancer:::prompts_prepare(
#   role = "linguist",
#   templates = templates,
#   language = "en",
#   target = "climate change",
#   text = c("some text", "and more"),
#   target_type = "the object"
# )
# 
# # Prepare prompts for domain expert
# templates <- list(
#   `system-domain` = "prompts/en/system-domain.md",
#   `user-domain` = "prompts/en/user-domain.md"
# )
# 
# prompts <- stancer:::prompts_prepare(
#   role = "domain",
#   templates = templates,
#   language = "en",
#   target = "artificial intelligence",
#   text = "AI will replace all human jobs",
#   target_type = "the object",
#   domain = "technology and employment expert"
# )
# 
# # Prepare prompts for judger with additional options
# templates <- list(
#   `system-judger` = "prompts/en/system-judger.md",
#   `user-judger` = "prompts/en/user-judger.md",
#   reference_resolution = "prompts/en/reference-resolution.md",
#   claim_resolution = "prompts/en/claim-resolution.md",
#   scale_guide = "prompts/en/guide-likert.md"
# )
# 
# prompts <- stancer:::prompts_prepare(
#   role = "judger",
#   templates = templates,
#   language = "en",
#   add_reference_resolution = TRUE,
#   add_claim_resolution = TRUE,
#   target = "Climate change is a real concern",
#   text = "Global warming is a serious threat",
#   target_type = "the claim",
#   FavourResponse = "This text is almost a paraphrase of the target claim",
#   AgainstResponse = "This is most probably a parody of climate speeches.",
#   NeutralResponse = "Context is missing and we should be sceptical"
# )
```

## Tasks

```{r function-tasks_validate}
#' Validate task configuration
#'
#' @description
#' Internal helper to ensure that the generated tasks (user prompts) meet
#' the expected criteria after interpolation. It checks for correct length,
#' data type, and ensures no empty strings were produced.
#'
#' @param tasks A character vector of interpolated tasks.
#' @param expected_length The number of tasks expected (usually matching the number of input texts).
#'
#' @return Invisibly TRUE if validation passes or throws an error if tasks are not a character vector, have incorrect length, or contain empty strings.
#'
#' @keywords internal
#' @filename prompting

tasks_validate <- function(tasks, expected_length) {
  rlang::try_fetch(
    ellmer:::check_number_whole(expected_length, min = 1),
    error = function(e) abort("Wrong {.arg expected_length} argument.")
  )

  if (any(is.na(tasks))) {
    abort("User prompt list contains missing values")
  }

  if (length(tasks) != expected_length || !is.character(tasks)) {
    abort(
      c(
        "User prompt interpolation returned unexpected results",
        "i" = "Consider checking variable placeholders in user templates"
      )
    )
  }

  if (any(nchar(tasks) == 0)) {
    abort(
      c(
        "User prompt is empty after interpolation",
        "i" = "Check that all required variables are provided"
      )
    )
  }

  invisible(TRUE)
}
```

```{r function-tasks_prepare}
#' Prepare chat objects and tasks for execution
#'
#' @description
#' Organises the final structure for parallel execution by cloning the base
#' chat object and pairing it with the corresponding system and user prompts.
#'
#' @details
#' This function handles two main scenarios for system prompts:
#' 1. A single system prompt applied to all tasks.
#' 2. A vector of system prompts matching the number of texts (e.g., when
#'    different domain experts are assigned to different texts).
#'
#' The function clones the `chat_base` for each system prompt to ensure
#' environment isolation during parallel processing. It also performs
#' validation on the user tasks using [tasks_validate()].
#'
#' @param chat_base An `ellmer::Chat` object used as a template.
#' @param prompts A list containing `system` (character vector) and
#'   `task` (character vector) prompts, typically from [prompts_prepare()].
#' @param n_texts Integer. The expected number of texts/tasks to process.
#'
#' @return A list with two elements:
#' - `chats`: A list of cloned and configured `ellmer::Chat` objects.
#' - `tasks`: A character vector of user prompts (tasks).
#'
#' @keywords internal
#' @filename prompting

tasks_prepare <- function(chat_base, prompts, n_texts) {
  # General case when there may be > 1 system prompts
  ## (e.g. multiple domain roles)
  chats <- lapply(
    prompts$system,
    function(system_prompt) {
      # Each time clone the base chat
      chat <- chat_base$clone(deep = TRUE)
      # Set a system prompt
      chat$set_system_prompt(system_prompt)
      return(chat)
    }
  )

  if (length(chats) != 1 & length(chats) != n_texts) {
    abort(
      "Length of system and user prompts must match {.val {n_texts}}, \\
            got {.val {length(chats)}}"
    )
  }

  tasks_validate(prompts$task, n_texts)

  list(
    chats = chats,
    tasks = prompts$task
  )
}
```

---

# Expert Analysis

## Prepare Expert Chats

```{r function-prepare_expert_chats}
#' Prepare chat configurations for expert agents
#'
#' @description
#' A high-level internal helper that orchestrates the preparation of expert
#' sessions for specific analysis roles (linguist, domain expert, or interpreter).
#'
#' @details
#' This function acts as a bridge between the input data and the execution
#' engine. It extracts necessary variables from the `inputs` list,
#' interpolates the corresponding templates using [prompts_prepare()],
#' and then initialises the chat objects via [tasks_prepare()].
#'
#' @param inputs A list containing all necessary data for prompt interpolation:
#'   - `texts`: Corpus of documents to analyse.
#'   - `targets`: Character vector of analysis targets.
#'   - `target_types`: Character vector of localised target types.
#'   - `prompt_templates`: List of paths to markdown template files.
#'   - `language`: Character string (e.g., "en").
#'   - `domain_roles`: (Optional) Character vector for the domain expert role.
#' @param chat_base An `ellmer::Chat` object used as a template for cloning.
#' @param expert_role Character. One of `"linguist"`, `"domain"`, or `"interpreter"`.
#'
#' @return A list containing:
#' - `chats`: A list of configured `ellmer::Chat` objects.
#' - `tasks`: A character vector of interpolated user prompts.
#'
#' @keywords internal
#' @importFrom rlang arg_match
#' @filename agents

prepare_expert_chats <- function(
    inputs,
    chat_base,
    expert_role = c('linguist', 'domain', 'interpreter')
) {
  expert_role <- rlang::arg_match(
    expert_role,
    c('linguist', 'domain', 'interpreter'),
    multiple = FALSE
  )

  # Load instructions for the expert
  prompts <- with(
    inputs,
    prompts_prepare(
      expert_role,
      templates = prompt_templates,
      language = language,
      text = texts,
      target = targets,
      target_type = target_types,
      domain = domain_roles
    )
  )

  tasks_prepare(chat_base, prompts, length(inputs$texts))
}
```

## Execute Role

```{r function-execute_role}
#' Execute expert role analysis
#'
#' @description
#' An internal worker that handles the execution of LLM requests for a
#' specific expert role. It switches between parallel execution
#' (when a single system prompt is shared) and sequential execution (when
#' each task requires a unique system prompt).
#'
#' @param expert_role Character. The role to execute (e.g., "linguist").
#' @param inputs A list containing texts, targets, and prompt templates.
#' @param chat_base An `ellmer::Chat` object used as a template.
#' @param info Character. A descriptive label for logging and error messages.
#' @param rpm Integer. Requests per minute limit for parallel execution.
#' @param verbose Logical. Whether to print progress messages to the console.
#'
#' @return A character vector of LLM responses, one for each input text.
#'
#' @keywords internal
#' @importFrom ellmer parallel_chat_text
#' @filename agents

execute_role <- function(
    expert_role,
    inputs,
    chat_base,
    info,
    rpm = 500,
    verbose = TRUE
) {
  tasks <- prepare_expert_chats(
    inputs,
    chat_base = chat_base,
    expert_role = expert_role
  )

  if (verbose) cat('    \U1F4CA', glue::glue('{info}...'), '\n')

  # If multiple system_prompts, e.g
  if (length(tasks$chats) > 1) {
    # Use sequential analysis
    results <- vapply(
      seq_along(tasks$tasks),
      function(task_i) {
        chat <- tasks$chats[[task_i]]
        result <- chat$chat(tasks$tasks[[task_i]], echo = 'none')
        if (!is.character(result)) {
          abort("Unexpected results in {info} {task_i}")
        }
        result
      },
      character(1)
    ) |> catch(tolower(info))
  } else {
    # Use parallel analysis
    results <- ellmer::parallel_chat_text(
      # Unpack the Chat object
      chat = tasks$chats[[1]],
      prompts = tasks$tasks,
      rpm = rpm
    ) |> catch(tolower(info))
  }

  if (!is.character(results) || (length(results) != length(inputs$texts))) {
    abort(
      c(
        "Unexpected results in {info}",
        "x" = "Input and output lengths differ: {length(inputs$texts)} vs {length(results)}"
      )
    )
  }

  results
}
```

---

# Debate Stage

## Prepare Debater Chats

```{r function-prepare_debater_chats}
#' Prepare chat configurations for debate agents
#'
#' @description
#' Prepares the prompts and chat objects for Stage 2 (Debate). It takes the
#' linguistic, domain, and social media analysis results from Stage 1 and
#' injects them into the 'debater' templates.
#'
#' @param stance Character. The stance label to be debated for (used in prompts).
#' @param inputs A list containing:
#'   - `texts`, `targets`, `target_types`, `language`, `prompt_templates`
#'   - `analysis_results`: A list with `linguistic`, `domain`, and `social_media`
#'     vectors from Stage 1.
#' @param chat_base An `ellmer::Chat` object.
#'
#' @return A list with two elements:
#'   - `chats`: A list containing a single `ellmer::Chat` object.
#'   - `tasks`: A character vector of user prompts for each input text.
#'
#' @keywords internal
#' @filename agents

prepare_debater_chats <- function(
    stance,
    inputs,
    chat_base
) {
  # Prepare prompts using all context from Stage 1
  prompts <- with(
    inputs,
    prompts_prepare(
      'debater',
      language = language,
      templates = prompt_templates,
      text = texts,
      target_type = target_types,
      target = targets,
      stance = stance,
      LingResponse = analysis_results$linguistic,
      ExpertResponse = analysis_results$domain,
      UserResponse = analysis_results$social_media
    )
  )

  # Stage 2 requires a consistent system prompt across all items
  if (length(prompts$system) > 1) {
    abort(
      c(
        "Multiple system prompts are not allowed at stage 2",
        "i" = "Consider checking variable placeholders in system templates"
      )
    )
  }

  tasks_prepare(chat_base, prompts, length(inputs$texts))
}
```

---

# Judgement Stage

## Prepare Judger Chats

```{r function-prepare_judger_chats}
#' Prepare chat configurations for judge agent
#'
#' @param inputs Input texts
#' @param chat_base Base chat configuration
#'
#' @return List of chat configurations
#' @keywords internal
#' @filename agents

prepare_judger_chats <- function(
    inputs,
    chat_base
) {
  prompts <- with(
    inputs,
    prompts_prepare(
      'judger',
      language = language,
      templates = prompt_templates,
      text = texts,
      target_type = target_types,
      target = targets,
      FavourResponse = debate_results$positive,
      AgainstResponse = debate_results$negative,
      NeutralResponse = debate_results$neutral,
      # Additional instructions based on object types
      add_reference_resolution = 'object' %in% types,
      add_claim_resolution = 'claim' %in% types
    )
  )

  if (length(prompts$system) > 1) {
    abort(
      c(
        "Multiple system prompts are not allowed at stage 3",
        "i" = "Consider checking variable placeholders in system templates"
      )
    )
  }

  tasks_prepare(chat_base, prompts, length(inputs$texts))
}
```

---

# Pipeline Stages

## Stage 1: Parallel Analysis

```{r function-stage_1_parallel_analysis}
#' Execute Stage 1: Parallel expert analysis
#'
#' @description
#' Orchestrates the first stage of the analysis pipeline. It validates the
#' input data, then sequentially triggers three distinct expert analyses:
#' linguistic, domain-specific, and social media interpretation.
#'
#' @details
#' Although the function name contains "parallel", it refers to the
#' independent nature of the expert roles (they don't depend on each
#' other's output). Each role is executed via [execute_role()], which
#' may itself use parallel API requests.
#'
#' @param inputs A list containing all necessary data:
#'   - `texts`: Character vector to analyse.
#'   - `targets`: Character vector of targets.
#'   - `target_types`: Character vector of target types.
#'   - `prompt_templates`: List of template paths.
#'   - `language`: Character string.
#'   - `domain_roles`: Character vector for the domain expert.
#' @param chat_base An `ellmer::Chat` object used as a template.
#' @param verbose Logical. Whether to print stage progress and timing.
#' @param rpm Integer. Requests per minute limit for API calls.
#'
#' @return The `inputs` list with an additional element `analysis_results`,
#'   which is a named list containing character vectors:
#'   - `linguistic`: Results from the linguist role.
#'   - `domain`: Results from the domain expert role.
#'   - `social_media`: Results from the interpreter role.
#'
#' @keywords internal
#' @importFrom tictoc tic toc
#' @importFrom glue glue
#' @filename stages

stage_1_parallel_analysis <- function(
    inputs,
    chat_base,
    verbose = TRUE,
    rpm = 500
) {
  # Validate input structure
  required_fields <- c(
    'prompt_templates',
    'language',
    'texts',
    'targets',
    'target_types',
    'domain_roles'
  )
  validate_fields(inputs, required_fields, 'expert analysis')

  n <- length(inputs$texts)

  tictoc::tic('Stage 1')
  if (verbose) {
    cat(
      glue::glue("\U23F3 Stage 1: Expert analysis ({n} items)..."),
      "\n"
    )
  }

  expert_roles <- c('linguist', 'domain', 'interpreter')
  expert_info <- c(
    linguist = 'Linguistic analysis',
    domain = 'Domain expert analysis',
    interpreter = 'Social media interpretation'
  )
  analysis_results <- lapply(
    expert_roles,
    function(role) {
      execute_role(
        role,
        inputs = inputs,
        chat_base = chat_base,
        info = expert_info[[role]],
        rpm = rpm,
        verbose = verbose
      )
    }
  )
  names(analysis_results) <- c('linguistic', 'domain', 'social_media')

  # The result is a list of character vectors of length length(texts)
  ## One vector for each role
  inputs[['analysis_results']] <- analysis_results

  tictoc::toc(func.toc = stage_complete, quiet = !verbose)

  inputs
}
```

## Stage 2: Parallel Debates

```{r function-stage_2_parallel_debates}
#' Execute Stage 2: Parallel agent debates
#'
#' @param inputs Input texts
#' @param chat_base Chat configuration
#' @param verbose Show progress
#' @param rpm Requests per minute limit
#'
#' @return Debate results
#' @keywords internal
#' @filename stages

stage_2_parallel_debates <- function(
    inputs,
    chat_base,
    verbose = TRUE,
    rpm = 500
) {
  # Validate input structure
  required_fields <- c(
    'prompt_templates',
    'language',
    'texts',
    'targets',
    'target_types',
    'analysis_results',
    'scale'
  )
  validate_fields(inputs, required_fields, 'stance debates')

  n <- length(inputs$texts)

  # Ensure Stage 1 results are present and valid
  validate_inputs(inputs$analysis_results$linguistic, n, 'Linguistic analysis')
  validate_inputs(inputs$analysis_results$domain, n, 'Domain expert analysis')
  validate_inputs(inputs$analysis_results$social_media, n, 'Social media analysis')

  tictoc::tic('Stage 2')
  if (verbose) {
    cat(
      glue::glue(
        "\U23F3 Stage 2: Parallel debates ({n} items \U00D7 3 stances)..."),
      "\n"
    )
  }

  stance_labels <- c('positive', 'negative', 'neutral')

  # 1. Prepare tasks for each stance
  debater_tasks <- lapply(
    stance_labels,
    function(stance_label) {
      prepare_debater_chats(
        stance = l(
          inputs$language,
          glue::glue('stance_{stance_label}'),
          inputs$scale
        ),
        inputs = inputs,
        chat_base = chat_base
      )
    }
  )
  names(debater_tasks) <- stance_labels

  # 2. Execute parallel debates for each stance group
  inputs[['debate_results']] <- lapply(
    debater_tasks,
    function(debater_task) {
      ellmer::parallel_chat_text(
        chat = debater_task$chats[[1]],
        prompts = debater_task$tasks,
        rpm = rpm
      ) |> catch('stance debates')
    }
  )

  tictoc::toc(func.toc = stage_complete, quiet = !verbose)

  inputs
}
```

## Stage 3: Parallel Judgement

```{r function-stage_3_parallel_judgement}
#' Execute Stage 3: Judge synthesis
#'
#' @description
#' Orchestrates the final stage of the analysis. It takes the results from all
#' three debate branches and uses a 'judge' agent to produce a structured
#' final stance judgement. This stage uses structured output to ensure the
#' results follow a predefined schema.
#'
#' @param inputs A list containing:
#'   - `texts`, `targets`, `types`, `target_types`, `language`, `prompt_templates`, `scale`.
#'   - `debate_results`: Results from Stage 2.
#' @param chat_base An `ellmer::Chat` object.
#' @param verbose Logical. Whether to print progress and timing.
#' @param rpm Integer. Requests per minute limit for API calls.
#' @param ... Additional arguments passed to `ellmer::parallel_chat_structured`.
#'
#' @return The `inputs` list with an additional element `judgement_results`,
#'   containing the structured stance analysis.
#'
#' @keywords internal
#' @importFrom tictoc tic toc
#' @importFrom glue glue
#' @importFrom ellmer parallel_chat_structured
#' @importFrom cli cli_warn
#' @filename stages

stage_3_parallel_judgement <- function(
    inputs,
    chat_base,
    verbose = TRUE,
    rpm = 500,
    ...
) {
  n <- length(inputs$texts)

  # Validate input structure
  required_fields <- c(
    'prompt_templates',
    'language',
    'texts',
    'targets',
    'types',
    'target_types',
    'debate_results',
    'scale'
  )
  validate_fields(inputs, required_fields, 'stance decision')

  # Ensure Stage 2 results are present and valid
  validate_inputs(inputs$debate_results$positive, n, 'Positive stance debates')
  validate_inputs(inputs$debate_results$negative, n, 'Negative stance debates')
  validate_inputs(inputs$debate_results$neutral, n, 'Neutral stance debates')

  tictoc::tic('Stage 3')
  if (verbose) {
    cat(glue::glue("\U23F3 Stage 3: Parallel judgement ({n} items)..."), "\n")
  }

  # 1. Prepare tasks for the judge
  judger_tasks <- prepare_judger_chats(
    inputs,
    chat_base = chat_base
  )

  if (verbose) cat("    \u2696\UFE0F Running parallel judgements...\n")

  # 2. Execute structured parallel judgements
  inputs[['judgement_results']] <- ellmer::parallel_chat_structured(
    chat = judger_tasks$chats[[1]],
    prompts = judger_tasks$tasks,
    type = type_stance_analysis(
      inputs$language,
      inputs$prompt_templates$scale_description,
      inputs$scale
    ),
    rpm = rpm,
    convert = TRUE,
    ...
  ) |> catch('making final judgement')

  # 3. Check for failed judgements (NA values)
  if (any(is.na(inputs$judgement_results$stance))) {
    n_na <- sum(is.na(inputs$judgement_results$stance))
    cli::cli_warn(
      "{n_na} item{?s} returned NA stance",
      call = rlang::expr(llm_stance())
    )
  }

  tictoc::toc(func.toc = stage_complete, quiet = !verbose)

  inputs
}
```

---

# Main Analysis Function

## LLM Stance Detection

### GENERIC FUNCTION

```{r function-llm_stance}
#' Detect stance using COLA framework
#'
#' @description
#' `llm_stance()` implements a multi-stage "Collaborative rOle-infused LLM-based
#' Agents" (COLA) framework to detect stance in text towards a specific target
#' (entity or proposition).
#' It supports character vectors and data frames (compatible with `mall` style).
#'
#' @param x Input text(s) as a character vector or a data.frame containing (any of) `text`, `target`, `type`, `domain_role` or `language` arguments.
#' @param ... Additional arguments passed to `ellmer::parallel_chat_structured`.
#'
#' @return A `stance_result` object (for character method) or a modified
#'   data frame (for data.frame method).
#'
#' @export
llm_stance <- function(x, ...) {
  UseMethod("llm_stance")
}
```

### CHARACTER METHOD

```{r function-llm_stance.character}
#' @rdname llm_stance
#' @description
#' The `character` method performs the full three-stage analysis:
#' 1. **Expert Analysis**: Linguistic, domain, and social media perspectives.
#' 2. **Stance Debates**: Parallel generation of arguments for, against, and neutral.
#' 3. **Judgement**: Final synthesis into a structured stance score.
#'
#' @details
#' ## Chat Configuration Recommendations
#' For optimal results, consider the following requirements for each stage:
#' \itemize{
#'   \item \strong{Expert Analysis (Stage 1)}: Requires excellent real-world
#'   awareness and reasoning. Structured output is NOT required.
#'   \item \strong{Stance Debates (Stage 2)}: Requires creativity and strong
#'   writing skills. Structured output is NOT required.
#'   \item \strong{Stance Judgement (Stage 3)}: Requires large context width,
#'   superior reasoning, and \strong{MUST} support structured output (JSON mode).
#' }
#' 
#' @param target Character. The entity or claim the stance is directed at.
#' @param chat_base An `ellmer::Chat` object or a list of up to 3 such objects.
#' @param type Analysis type: "object" (entity-based) or "claim" (proposition-based).
#' @param language Language code (e.g., "en", "ru"). If `NULL`, auto-detected.
#' @param scale Stance scale: "categorical", "numeric", or "likert".
#' @param domain_role Optional specific role for the domain expert (e.g., "historian").
#' @param prompts_dir Optional path to custom prompt templates.
#' @param verbose Logical. Show progress and summary table.
#' @param rpm Integer. Global requests per minute limit.
#'
#' @return S3 object of class "stance_result" with:
#'   - summary: data.frame with results
#'   - analysis: expert analysis results
#'   - debates: debate results
#'   - judgements: final judgements
#'   - metadata: processing metadata
#' ⁠@importFrom ellmer is_chat
#' @importFrom cli cli_abort cli_warn
#' @importFrom rlang is_character is_scalar_character is_list arg_match is_installed
#' @importFrom glue glue
#' @importFrom tibble as_tibble
#' @importFrom tictoc tic toc
#' @export
llm_stance.character <- function(
    x,
    target,
    chat_base,
    type = c('object'),
    language = stancer_available_languages(),
    scale = c('categorical', 'numeric', 'likert'),
    domain_role = NULL,
    prompts_dir = NULL,
    verbose = TRUE,
    rpm = 20,
    ...
) {
  ## Validation ----
  tictoc::tic('Analysis')

  # Validate text and target
  text <- x
  validate_character(text)
  validate_character(target)

  n <- length(text)

  # Validate type
  if (rlang::is_character(type)) {
    if (length(type == 2) != n) {
      type <- rlang::arg_match(type, c('object', 'claim'), multiple = FALSE)
    } else {
      type <- rlang::arg_match(type, c('object', 'claim'), multiple = TRUE)
    }
  } else {
    cli::cli_abort(
      "{.arg type} must be a character vector, got {.cls {class(type)}}"
    )
  }

  # Validate language
  if (is.null(language) || length(language) > 1) {
    if (rlang::is_installed("cld2")) {
      language <- cld2::detect_language(text[[1]], lang_code = TRUE)
      cli::cli_warn(
        "The analysis language was automatically detected: {.val {language}}"
      )
    }
  }
  if (rlang::is_scalar_character(language)) {
    if (!language %in% stancer_available_languages()) {
      cli::cli_abort(
        c(
          "Language {.val {language}} not available",
          "i" = "Use one of: {.val {stancer_available_languages()}}"
        )
      )
    }
  } else {
    cli::cli_abort(
      "{.arg language} must be a character scalar, got \
            {.cls {class(language)}} of length {.val {length(language)}}"
    )
  }

  # Validate scale
  scale <- tolower(scale) # Can write ‘Likert’
  scale <- rlang::arg_match(scale, c('categorical', 'numeric', 'likert'))

  # Validate domain_role
  if (is.null(domain_role)) {
    domain_role <- switch(
      language,
      uk = 'sociologist', # 'соціолог',
      ru = 'sociologist', # 'социолог',
      'social commentator'
    )
  } else {
    if (!rlang::is_character(domain_role) || length(domain_role) < 1) {
      cli::cli_abort(
        "{.arg domain_role} must be a character vector, got {.cls {class(domain_role)}}"
      )
    }
    if (!(length(domain_role) == 1 || length(domain_role) == n)) {
      cli::cli_abort(
        c(
          "{.arg domain_role} must have length 1 or {n} (same as {.arg text})",
          "x" = "Got {length(domain_role)}"
        )
      )
    }
  }

  if (length(domain_role) > 1) {
    cli::cli_warn(
      "Multiple domain roles detected. Parallel execution is unsupported."
    )
  }

  ### Chat Validation ----
  # Validate and arrange the `chat_base` argument across stages
  #
  # Recommendations (see vignette and docstrings):
  # 1. Expert analysis (Stage 1):
  #     - Excellent real-world awareness
  #     - Decent reasoning skills
  #     - Sufficient writing skills
  #     - Context width depends on task
  #     - Structured output: NOT required
  # 2. Stance debates (Stage 2):
  #     - Sufficient context width
  #     - Creative and writing skills
  #     - Structured output: NOT required
  # 3. Stance decision (Stage 3):
  #     - Large context width
  #     - Excellent reasoning skills
  #     - Excellent real-world awareness
  #     - Structured output: REQUIRED
  chats <- if (rlang::is_list(chat_base)) {
    # Validate each element
    for (i in seq_along(chat_base)) {
      if (!ellmer:::is_chat(chat_base[[i]])) {
        cli::cli_abort(
          "Element {i} of {.arg chat_base} is not an {.cls ellmer::Chat} object"
        )
      }
    }

    # Arrange based on list length
    indices <- switch(
      length(chat_base),
      `1` = c(1, 1, 1),
      `2` = c(1, 1, 2),
      `3` = c(1, 2, 3),
      cli::cli_abort(
        c(
          "{.arg chat_base} must have 1, 2, or 3 elements",
          "x" = "You provided {length(chat_base)}"
        )
      )
    )
    lapply(indices, \(i) chat_base[[i]]$clone(deep = TRUE))

  } else if (ellmer:::is_chat(chat_base)) {
    rep(list(chat_base$clone(deep = TRUE)), times = 3)

  } else {
    cli::cli_abort(
      c(
        "{.arg chat_base} must be an {.cls ellmer::Chat} object or \
                a list of 1-3 {.cls ellmer::Chat} objects",
        "x" = "Got {.cls {class(chat_base)}}"
      )
    )
  }

  ## Prompts ----
  prompt_templates <- templates_collect(prompts_dir, language, scale)

  ## Preparation ----
  target <- recycle_arg(target, n)
  type <- recycle_arg(type, n)

  target_types <- sapply(type, type_to_term, language = language)

  inputs <- list(
    texts = text,
    targets = target,
    types = type,
    target_types = target_types,
    language = language,
    scale = scale,
    domain_roles = domain_role,
    prompt_templates = prompt_templates
  )

  if (verbose) {
    cat("\n")
    cat(strrep("=", 70), "\n")
    cat(glue::glue("\U1F50D STANCE ANALYSIS - Processing {n} item(s)"), "\n")
    cat(strrep("=", 70), "\n\n")
    cat(glue::glue("Types: {paste(unique(type), collapse = ', ')}"), "\n")
    cat(glue::glue("Language: {language}"), "\n")
    cat(
      glue::glue("Domain roles: {paste(unique(domain_role), collapse = ', ')}"),
      "\n"
    )
    cat("\n")
  }

  ## Analysis ----
  output <- inputs |>
    stage_1_parallel_analysis(chats[[1]], verbose, rpm) |>
    validate_stage('Stage 1 (Expert analysis)') |>
    stage_2_parallel_debates(chats[[2]], verbose, rpm) |>
    validate_stage('Stage 2 (Stance debates)') |>
    stage_3_parallel_judgement(chats[[3]], verbose, rpm, ...)

  if (is.null(output$judgement_results) || nrow(output$judgement_results) != n) {
    cli::cli_abort("Final stance judgement returned unexpected results")
  }

  # Additional postprocessing of quantitative and Likert stance labels
  if (scale == 'numeric') {
    # output$judgement_results$stance <- truncate(
    #     (output$judgement_results$stance - 50) / 50,
    #     -1,
    #     1
    # )
  } else if (scale == 'likert') {
    # Rename as proper noun
    scale <- 'Likert'
    output$judgement_results$stance <- factor(
      output$judgement_results$stance,
      levels = c(
        'Strongly Disagree',
        'Disagree',
        'Neutral',
        'Agree',
        'Strongly Agree'
      ),
      ordered = TRUE
    )
  } else if (scale == "categorical") {
    output$judgement_results$stance <- factor(
      as.character(output$judgement_results$stance),
      levels = c('Negative', 'Neutral', 'Positive'),
      ordered = FALSE
    )
  }

  ## Postprocessing ----
  summary_df <- data.frame(
    text = text,
    target = target,
    target_type = type,
    language = language
  ) |>
    cbind(output$judgement_results) |>
    tibble::as_tibble()

  if (verbose) {
    cat("\U1F4CA Summary Table:\n")
    print(summary_df)
    cat("\n")
    cat(strrep("=", 70), "\n")
    toc <- tictoc::toc(func.toc = stage_complete, quiet = FALSE)
    cat(strrep("=", 70), "\n\n")
  } else {
    toc <- tictoc::toc(func.toc = stage_complete, quiet = TRUE)
  }

  ## Return ----
  structure(
    list(
      summary = summary_df,
      analysis = output$analysis_results,
      debates = output$debate_results,
      judgements = output$judgement_results,
      metadata = list(
        n_total = n,
        n_processed = nrow(output$judgement_results),
        n_failed = sum(is.na(output$judgement_results$stance)),
        scale = scale,
        language = language,
        types = unique(type),
        domain_role = domain_role,
        model = unique(sapply(chats, \(chat) chat$get_model())),
        elapsed = toc$toc - toc$tic,
        timestamp = Sys.time()
      )
    ),
    class = c("stance_result", "list")
  )
}
```

```{r examples-llm_stance.character}
# result <- llm_stance(
#   x = "I support renewable energy",
#   target = "renewable energy",
#   chat_base = chat_model,
#   type = "object",
#   scale = "categorical"
# )
```

### DATA.FRAME METHOD

```{r function-llm_stance.data.frame}
#' @rdname llm_stance
#' @param text The text column (if `x` is a data.frame) or character vector to analyse.
#' @param .output_col Name of the column where results will be stored.
#'
#' @importFrom cli cli_abort cli_warn cli_inform
#' @importFrom rlang is_null as_name
#' @filename methods
#' @export

llm_stance.data.frame <- function(
    x,
    text = NULL,
    target = NULL,
    type = NULL,
    domain_role = NULL,
    language = NULL,
    chat_base,
    scale = 'categorical',
    prompts_dir = NULL,
    verbose = TRUE,
    rpm = 20,
    .output_col = '.stance',
    ...
) {
  data <- x

  # Tidy evaluation
  text_quo <- rlang::enquo(text)
  target_quo <- rlang::enquo(target)
  type_quo <- rlang::enquo(type)
  language_quo <- rlang::enquo(language)
  domain_role_quo <- rlang::enquo(domain_role)

  # Validate and get argument vectors
  text_vec <- catch_eval(text_quo, data)
  target_vec <-  catch_eval(target_quo, data)
  type_vec <- safe_eval(type_quo, data)
  language_vec <- safe_eval(language_quo, data)
  domain_role_vec <- safe_eval(domain_role_quo, data)

  # Fall-back to the defaults
  if (rlang::is_null(type_vec)) {
    type_vec <- c('object', 'claim')
  }

  if (rlang::is_null(language_vec)) {
    cli::cli_warn(
      c("Cannot find {.arg language} in data or environment")
    )
  }

  # Names for messages
  text_name <- rlang::as_name(text_quo)
  target_name <- rlang::as_name(target_quo)

  if (verbose) {
    cli::cli_inform(
      c(
        "i" = "Processing {.val {text_name}} and {.val {target_name}} columns"
      )
    )
  }

  result <- llm_stance(
    x = text_vec,
    target = target_vec,
    type = type_vec,
    language = language_vec,
    domain_role = domain_role_vec,
    chat_base = chat_base,
    scale = scale,
    prompts_dir = prompts_dir,
    verbose = verbose,
    rpm = rpm,
    ...
  )

  # Augment `data`
  n_rows <- nrow(data)
  n_results <- nrow(result$summary)

  if (n_results != n_rows) {
    cli::cli_warn(
      c(
        "Result has {n_results} rows, but data has {n_rows} rows",
        "i" = "Returning only matching rows"
      )
    )
    result$summary <- result$summary[seq_len(min(n_rows, n_results)), ]
  }

  data[[.output_col]] <- c(result$summary$stance, rep(NA, n_rows - n_results))

  attr(data, 'llm_stance_metadata') <- result$metadata

  if (verbose) {
    cli::cli_inform(
      c(
        "v" = "Analysis complete!",
        "i" = "Results added to {.val {(.output_col)}} column"
      )
    )
  }

  return(data)
}
```

---

# Result Inspection and Display

## Print Method

```{r function-print.stance_result}
#' Print stance_result object
#'
#' @description
#' A custom print method for `stance_result` objects. It displays a formatted
#' header with processing metadata (items processed, language, models used, etc.)
#' followed by the summary table of stance results.
#' @param x A `stance_result` object.
#' @param ... Additional arguments to [base::print.data.frame()] or other `print` methods for the summary table.
#'
#' @return The input object (invisibly).
#' @importFrom glue glue
#' @filename methods
#' @export

print.stance_result <- function(x, ...) {
    cat("Stance Analysis Result\n")
    cat(strrep("=", 60), "\n")
    cat(
        glue::glue("Processed: {x$metadata$n_processed}/{x$metadata$n_total} items"),
        "\n"
    )
    cat(glue::glue("Failed: {x$metadata$n_failed} items"), "\n")
    cat(glue::glue("Scale: {x$metadata$scale}"), "\n")
    cat(glue::glue("Language: {x$metadata$language}"), "\n")
    cat(glue::glue("Types: {paste(x$metadata$types, collapse = ', ')}"), "\n")
    cat(
        glue::glue(
            "Domain role(s): {paste(x$metadata$domain_role, collapse = ', ')}"
        ),
        "\n"
    )
    cat(
        glue::glue("Model(s) used: {paste(x$metadata$model, collapse = ', ')}"),
        "\n"
    )
    cat(glue::glue("Time elapsed: {round(x$metadata$elapsed, 2)} sec"), "\n")
    cat(glue::glue("Timestamp: {x$metadata$timestamp}"), "\n")
    cat(strrep("=", 60), "\n\n")

    cat("Summary Table:\n")
    print(x$summary, ...)

    invisible(x)
}
```

```{r examples-print.stance_result}
# result <- llm_stance(
#   x = "R is smart and tidy",
#   target = "R programming language",
#   chat_base = ellmer::chat_anthropic()
# )
# print(result)
```

## Summary Method

```{r function-summary.stance_result}
#' Summarise stance_result object
#'
#' @description
#' Provides a statistical summary of the stance analysis results. It displays
#' the frequency distribution of detected stances and a cross-tabulation
#' of stances by target entity or proposition.
#' @param object A `stance_result` object.
#' @param ... Additional arguments passed to [base::print.table()].
#'
#' @return The input object (invisibly).
#' @filename methods
#' @export

summary.stance_result <- function(object, ...) {
    cat("Stance Distribution:\n")
    print(table(object$summary$stance))
    cat("\n")

    cat("By Target:\n")
    print(table(object$summary$target, object$summary$stance), ...)

    invisible(object)
}
```

## Data Frame Conversion

```{r function-as.data.frame.stance_result}
#' Convert stance_result to data frame
#'
#' @description
#' Extracts the summary table from a `stance_result` object and returns it as
#' a data frame. This allows for easy further analysis using standard R tools.
#' @param x A `stance_result` object
#' @param row.names Optional NULL or a character vector giving the row names for the data frame. Missing values are not allowed.
#' @param optional Logical. If TRUE, setting row names and converting column names to syntactic names is optional.
#' @param ... Additional arguments passed to [base::as.data.frame()].
#'
#' @return A Data frame
#' @filename methods
#' @export

as.data.frame.stance_result <- function(
    x,
    row.names = NULL,
    optional = FALSE,
    ...
) {
  x$summary
}
```

## Inspect Method

```{r function-inspect}
#' Inspect method for `stance_result` objects
#'
#' @param x A `stance_result` object.
#' @param ... Additional arguments passed to methods.
#'
#' @return The input object (invisibly).
#' @filename methods
#' @export
#' @rdname inspect.stance_result

inspect <- function(x, ...) {
  UseMethod("inspect")
}
```

```{r function-inspect.stance_result}
#' Inspect stance_result contents
#'
#' @description
#' A diagnostic tool to peer into the internal steps of the stance analysis.
#' It allows to view metadata, raw linguistic/expert/veteran analysis,
#' internal debates between agents, or the final explanation for a specific row.
#'
#' @param what Character. Primary level: 'metadata', 'analysis', 'debates', or 'explanation'.
#' @param within Optional character. Secondary level (e.g., 'linguistic' for analysis, 'positive' for debates).
#' @param index Integer. The row index of the text item to inspect.
#'
#' @filename methods
#' @export

inspect.stance_result <- function(
    x,
    what = c('metadata', 'analysis', 'debates', 'explanation'),
    within = NULL,
    index = 1,
    ...
) {

  what <- match.arg(what)
  ellmer:::check_number_whole(index, min = 1)

  switch(
    what,

    #### Inspect Metadata ----
    metadata = {
      cat("\n", cli::rule("METADATA"), "\n\n")
      yaml::write_yaml(x$metadata, file = stdout())
    },

    #### Inspect Analysis Results ----
    analysis = {
      if (is.null(within)) {
        cat("\n", cli::rule("ANALYSIS - Available sections"), "\n\n")
        available <- names(x$analysis)
        cat("Available sections:\n")
        for (section in available) {
          cat(
            cli::col_cyan(
              sprintf("  \U2022 %s\n", section)
            )
          )
        }
        cat(
          "\nUsage: ",
          "inspect(result, what = 'analysis', ",
          "within = 'section_name')\n\n",
          sep = ""
        )
      } else {
        if (!within %in% names(x$analysis)) {
          cli::cli_abort(
            "Section '{within}' not found in analysis. ",
            "Available: {.val {names(x$analysis)}}"
          )
        }

        if (index > length(x$analysis)) {
          cli::cli_abort(
            "Index {index} out of bounds. ",
            "Analysis has {length(x$analysis)} rows."
          )
        }

        content <- x$analysis[[within]][index]

        cat(
          "\n",
          cli::rule(
            sprintf(
              "ANALYSIS: %s [row %d]",
              within,
              index
            )
          ),
          "\n\n"
        )

        if (is.na(content) || is.null(content)) {
          cat(cli::col_yellow("(empty)\n\n"))
        } else {
          cat(ellmer::interpolate(content), fill = TRUE)
          cat("\n")
        }
      }
    },

    #### Inspect Stance Debates ----
    debates = {
      if (is.null(within)) {
        cat("\n", cli::rule("DEBATES - Available sections"), "\n\n")
        available <- names(x$debates)
        cat("Available sections:\n")
        for (section in available) {
          cat(
            cli::col_cyan(
              sprintf("  \U2022 %s\n", section)
            )
          )
        }
        cat(
          "\nUsage: ",
          "inspect(result, what = 'debates', ",
          "within = 'section_name')\n\n",
          sep = ""
        )
      } else {
        if (!within %in% names(x$debates)) {
          cli::cli_abort(
            "Section '{within}' not found in debates. ",
            "Available: {.val {names(x$debates)}}"
          )
        }

        if (index > length(x$debates)) {
          cli::cli_abort(
            "Index {index} out of bounds. ",
            "Debates has {length(x$debates)} rows."
          )
        }

        content <- x$debates[[within]][index]

        cat(
          "\n",
          cli::rule(
            sprintf(
              "DEBATES: %s [row %d]",
              within,
              index
            )
          ),
          "\n\n"
        )

        if (is.na(content) || is.null(content)) {
          cat(cli::col_yellow("(empty)\n\n"))
        } else {
          cat(ellmer::interpolate(content), fill = TRUE)
          cat("\n")
        }
      }
    },

    #### Inspect Explanation ----
    explanation = {
      if (!is.null(within)) {
        cli::cli_warn(
          c(
            'Stance result explanation has no sections.',
            'Unused argument: {.arg within}'
          )
        )
      }

      if (index > length(x$summary$explanation)) {
        cli::cli_abort(
          c(
            "Index {index} out of bounds. ",
            "The data have {nrow(x$summary)} rows."
          )
        )
      }

      content <- x[['summary']][['explanation']][index]

      cat(
        "\n",
        cli::rule(
          sprintf("EXPLANATION: [row %d]", index)
        ),
        "\n\n"
      )

      if (is.na(content) || is.null(content)) {
        cat(cli::col_yellow("(empty)\n\n"))
      } else {
        cat(ellmer::interpolate(content), fill = TRUE)
        cat("\n")
      }
    }
  )

  invisible(x)
}
```

---

# Inflate

```{r development-inflate, eval=FALSE}
fusen::inflate(
  pkg = here::here(),
  flat_file = here::here('dev', 'flat_stancer.Rmd'),
  codecov = TRUE,
  clean = TRUE,
  vignette_name = "Stance Analysis with ellmer",
  open_vignette = FALSE
)
```

```{r development-document, eval = FALSE}
# usethis::use_logo('../hex-stancer.png')
# usethis::use_lifecycle_badge('Experimental')
devtools::build_readme()
```
