---
title: 'rcola: Collaborative Role-infused LLM-based Agents for Stance Detection'
output: html_document
editor_options: 
chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
library(rlang)
library(mockery)
library(cli)
library(ellmer)
library(yaml)
```

```{r description}
fusen::fill_description(
  pkg = here::here(),
  fields = list(
    Title = "LLM-based Sentiment and Stance Analysis with COLA Framework",
    Description = "An R package for sentiment and stance detection using the COLA (Collaborative rOle-infused LLM-based Agents) framework. Provides functions for single and batch text analysis with support for multiple language models via the ellmer package.",
    `Authors@R` = c(
      person("Andrej", "Pawluczenko", email = "a.pawluczenko@gmail.com", role = c("aut", "cre"))
    )
  ),
  overwrite = TRUE
)
# Define License with use_*_license()
usethis::use_ccby_license()
```

# Introduction

## Overview

**rcola** is an R implementation of the COLA (**C**ollaborative r**O**Le-infused **L**LM-based **A**gents) framework for automated stance detection in texts. 

The package employs a three-stage collaborative approach:

1. **Stage 1: Parallel Analysis** — Specialised LLM-based agents (linguist, domain expert, social interpreter) analyse texts from their respective perspectives
2. **Stage 2: Parallel Debates** — Agents debate their analyses to reach consensus
3. **Stage 3: Parallel Judgment** — A judge agent synthesises the debate and produces final stance classification

## Key Features

- 🤖 Multi-perspective analysis using collaborative LLM agents
- 🌍 Support for multiple languages
- 📊 Multiple stance scales (categorical, numeric, Likert)
- 🎯 Domain-specific analysis with custom roles
- 📈 Batch processing for data frames
- 🔍 Detailed inspection tools for results

## Quick Example

```r

```

---

# Utility Functions

## Text Truncation

```{r function-truncate}
#' Truncate numeric values to a range
#'
#' Constrains numeric values to fall within the range [a, b].
#' Values below a are set to a, values above b are set to b.
#'
#' @param x Numeric vector to truncate
#' @param a Numeric. Lower bound (minimum value)
#' @param b Numeric. Upper bound (maximum value)
#'
#' @return Numeric vector with values constrained to [a, b]
#' @keywords internal
#'
#' @examples
#' truncate(c(-1, 0, 5, 10), a = 0, b = 5)
#' # Returns: 0, 0, 5, 5

truncate <- function(x, a, b) {
  if (!is.numeric(x)) {
    cli::cli_abort("{.arg x} must be numeric")
  }
  if (!is.numeric(a)) {
    cli::cli_abort("{.arg a} must be numeric")
  }
  if (!is.numeric(b)) {
    cli::cli_abort("{.arg b} must be numeric")
  }
  
  pmin(pmax(x, a), b)
}
```

```{r tests-truncate}
# Test 1: Basic functionality - values within range
test_that('truncate returns values unchanged when within range', {
  expect_equal(truncate(3, a = 0, b = 5), 3)
  expect_equal(truncate(c(1, 2, 3), a = 0, b = 5), c(1, 2, 3))
})

# Test 2: Values below lower bound
test_that('truncate sets values below a to a', {
  expect_equal(truncate(-1, a = 0, b = 5), 0)
  expect_equal(truncate(c(-5, -1, 0), a = 0, b = 5), c(0, 0, 0))
})

# Test 3: Values above upper bound
test_that('truncate sets values above b to b', {
  expect_equal(truncate(10, a = 0, b = 5), 5)
  expect_equal(truncate(c(5, 6, 10), a = 0, b = 5), c(5, 5, 5))
})

# Test 4: Mixed values (below, within, above)
test_that('truncate handles mixed values correctly', {
  expect_equal(
    truncate(c(-1, 0, 3, 5, 10), a = 0, b = 5),
    c(0, 0, 3, 5, 5)
  )
})

# Test 5: Negative ranges
test_that('truncate works with negative ranges', {
  expect_equal(truncate(c(-10, -5, 0), a = -8, b = -2), c(-8, -5, -2))
})

# Test 6: Single value at boundaries
test_that('truncate handles boundary values', {
  expect_equal(truncate(0, a = 0, b = 5), 0)
  expect_equal(truncate(5, a = 0, b = 5), 5)
})

# Test 7: Vectorized operation
test_that('truncate is vectorized', {
  x <- c(-10, -5, 0, 5, 10, 15)
  result <- truncate(x, a = 0, b = 10)
  expect_equal(result, c(0, 0, 0, 5, 10, 10))
})

# Test 8: Floating point values
test_that('truncate works with floating point numbers', {
  expect_equal(truncate(3.5, a = 0, b = 5), 3.5)
  expect_equal(truncate(-0.5, a = 0, b = 5), 0)
  expect_equal(truncate(5.5, a = 0, b = 5), 5)
})

# Test 9: Very large and very small numbers
test_that('truncate handles extreme values', {
  expect_equal(truncate(1e10, a = 0, b = 100), 100)
  expect_equal(truncate(-1e10, a = 0, b = 100), 0)
})

# Test 10: Zero as bound
test_that('truncate works with zero as bound', {
  expect_equal(truncate(c(-5, 0, 5), a = 0, b = 10), c(0, 0, 5))
  expect_equal(truncate(c(-5, 0, 5), a = -10, b = 0), c(-5, 0, 0))
})

# Test 11: Equal bounds (a == b)
test_that('truncate works when a equals b', {
  expect_equal(truncate(c(-5, 0, 5), a = 3, b = 3), c(3, 3, 3))
})

# Test 12: Reversed bounds (a > b) - should still work with pmin/pmax
test_that('truncate handles reversed bounds', {
  # pmin and pmax will handle this gracefully
  result <- truncate(5, a = 10, b = 0)
  expect_equal(result, 0)
})

# Test 13: Empty vector
test_that('truncate handles empty vectors', {
  expect_equal(truncate(numeric(0), a = 0, b = 5), numeric(0))
})

# Test 14: NA values are preserved
test_that('truncate preserves NA values', {
  result <- truncate(c(1, NA, 3), a = 0, b = 5)
  expect_equal(result[1], 1)
  expect_true(is.na(result[2]))
  expect_equal(result[3], 3)
})

# Test 15: Inf values
test_that('truncate handles infinite values', {
  expect_equal(truncate(Inf, a = 0, b = 5), 5)
  expect_equal(truncate(-Inf, a = 0, b = 5), 0)
})

# Error handling: x is not numeric
test_that('truncate rejects non-numeric x', {
  expect_error(truncate("text", a = 0, b = 5), 'must be numeric', fixed = TRUE)
  expect_error(truncate(TRUE, a = 0, b = 5), 'must be numeric', fixed = TRUE)
  expect_error(truncate(list(1, 2), a = 0, b = 5), 'must be numeric', fixed = TRUE)
})

# Error handling: a is not numeric
test_that('truncate rejects non-numeric a', {
  expect_error(truncate(5, a = "zero", b = 5), 'must be numeric', fixed = TRUE)
  expect_error(truncate(5, a = list(0), b = 5), 'must be numeric', fixed = TRUE)
})

# Error handling: b is not numeric
test_that('truncate rejects non-numeric b', {
  expect_error(truncate(5, a = 0, b = "five"), 'must be numeric', fixed = TRUE)
  expect_error(truncate(5, a = 0, b = list(5)), 'must be numeric', fixed = TRUE)
})

# Error handling: all three are non-numeric
test_that('truncate rejects multiple non-numeric arguments', {
  expect_error(truncate("x", a = "a", b = "b"), 'must be numeric', fixed = TRUE)
})

# Error handling: NULL values
test_that('truncate rejects NULL values', {
  expect_error(truncate(NULL, a = 0, b = 5), 'must be numeric', fixed = TRUE)
  expect_error(truncate(5, a = NULL, b = 5), 'must be numeric', fixed = TRUE)
  expect_error(truncate(5, a = 0, b = NULL), 'must be numeric', fixed = TRUE)
})
```

## Error Handling

```{r function-abort}
#' Abort with formatted error message
#'
#' @param ... Arguments passed to cli::cli_abort()
#'
#' @return Throws error
#' @keywords internal
#' @importFrom cli cli_abort

abort <- function(msg) {
  cli::cli_abort(
    msg,
    call = rlang::expr(llm_stance()),
    .envir = rlang::caller_env()
  )
}
```

```{r tests-abort}
# Test 1: abort throws an error
test_that('abort throws an error', {
  expect_error(abort("Test error message"))
})

# Test 2: abort with simple string message
test_that('abort with simple string message', {
  expect_error(
    abort("This is an error"),
    "This is an error",
    fixed = TRUE
  )
})

# Test 3: abort with cli formatting - .arg
test_that('abort with .arg formatting', {
  expect_error(
    abort("{.arg x} must be numeric"),
    "must be numeric",
    fixed = TRUE
  )
})

# Test 4: abort with cli formatting - .val
test_that('abort with .val formatting', {
  expect_error(
    abort("Invalid value: {.val {5}}"),
    "Invalid value",
    fixed = TRUE
  )
})

# Test 5: abort with cli formatting - .cls
test_that('abort with .cls formatting', {
  expect_error(
    abort("Expected {.cls character}, got {.cls numeric}"),
    "Expected",
    fixed = TRUE
  )
})

# Test 6: abort with multiple cli formatting elements
test_that('abort with multiple formatting elements', {
  expect_error(
    abort(
      "Argument {.arg x} must be {.cls numeric}, not {.cls list}"
    ),
    "must be",
    fixed = TRUE
  )
})

# Test 7: abort with interpolation - variable substitution
test_that('abort with variable interpolation', {
  var_name <- "test_var"
  expect_error(
    abort("Variable '{var_name}' not found"),
    "test_var",
    fixed = TRUE
  )
})

# Test 8: abort with complex interpolation
test_that('abort with complex interpolation', {
  lang <- "en"
  expect_error(
    abort("Language '{lang}' not available"),
    "Language 'en' not available",
    fixed = TRUE
  )
})

# Test 9: abort preserves error class
test_that('abort creates rlang_error', {
  expect_error(
    abort("Test error"),
    class = "rlang_error"
  )
})

# Test 10: abort with empty message
test_that('abort with empty message', {
  expect_error(abort(""))
})

# Test 11: abort with very long message
test_that('abort with long message', {
  long_msg <- paste(rep("a", 1000), collapse = "")
  expect_error(abort(long_msg))
})

# Test 12: abort with special characters
test_that('abort with special characters', {
  expect_error(
    abort("Error with special chars: @#$%^&*()"),
    "special chars",
    fixed = TRUE
  )
})

# Test 13: abort with quotes in message
test_that('abort with quotes in message', {
  expect_error(
    abort('Error with "double" and \'single\' quotes'),
    "quotes",
    fixed = TRUE
  )
})

# Test 14: abort with list formatting
test_that('abort with list formatting', {
  expect_error(
    abort("Available options: {.val {c('a', 'b', 'c')}}"),
    "Available",
    fixed = TRUE
  )
})

# Test 15: abort with vector in message
test_that('abort with vector interpolation', {
  vals <- c("x", "y", "z")
  expect_error(
    abort("Invalid values: {.val {vals}}"),
    "Invalid",
    fixed = TRUE
  )
})

# Test 16: abort with numeric values in message
test_that('abort with numeric interpolation', {
  n <- 111
  expect_error(
    abort("Expected {n} items, got 0"),
    "Expected 111",
    fixed = TRUE
  )
})

# Test 17: abort with logical values in message
test_that('abort with logical values', {
  flag <- TRUE
  expect_error(
    abort("Flag is {flag}"),
    "Flag is TRUE",
    fixed = TRUE
  )
})

# Test 18: abort with NA values
test_that('abort with NA values', {
  na_val <- NA
  expect_error(
    abort("Value is {na_val}"),
    "Value",
    fixed = TRUE
  )
})

# Test 19: abort with function name in message
test_that('abort with function reference', {
  expect_error(
    abort("Function {.fn truncate} failed"),
    "failed",
    fixed = TRUE
  )
})

# Test 20: abort uses caller environment
test_that('abort uses caller environment', {
  test_var <- "from_caller"
  expect_error(
    abort("Variable: {test_var}"),
    "from_caller",
    fixed = TRUE
  )
})

# Test 21: abort in nested function call
test_that('abort in nested function', {
  inner_func <- function() {
    abort("Error from inner function")
  }
  
  outer_func <- function() {
    inner_func()
  }
  
  expect_error(
    outer_func(),
    "Error from inner function",
    fixed = TRUE
  )
})

# Test 22: abort does not return value (throws immediately)
test_that('abort does not return', {
  result <- tryCatch(
    {
      abort("Test")
      "should not reach here"
    },
    error = function(e) {
      "caught error"
    }
  )
  expect_equal(result, "caught error")
})

# Test 23: abort with .arg and variable
test_that('abort with .arg and variable name', {
  param_name <- "data"
  expect_error(
    abort("Argument {.arg {param_name}} is invalid"),
    "invalid",
    fixed = TRUE
  )
})

# Test 24: abort with .cls and actual class
test_that('abort with .cls and object class', {
  obj <- list(a = 1)
  expect_error(
    abort("Expected data.frame, got {.cls {class(obj)}}"),
    "Expected",
    fixed = TRUE
  )
})

# Test 25: abort with multiple .val elements
test_that('abort with multiple .val elements', {
  expect_error(
    abort(
      "Got {.val {1}}, expected one of {.val {c(2, 3, 4)}}"
    ),
    "Got",
    fixed = TRUE
  )
})

# Test 26: abort preserves message formatting
test_that('abort message is not modified', {
  msg <- "Exact error message"
  tryCatch(
    abort(msg),
    error = function(e) {
      expect_true(grepl(msg, conditionMessage(e), fixed = TRUE))
    }
  )
})

# Test 27: abort called multiple times creates different errors
test_that('multiple abort calls create separate errors', {
  error1 <- tryCatch(
    abort("Error 1"),
    error = function(e) e
  )
  
  error2 <- tryCatch(
    abort("Error 2"),
    error = function(e) e
  )
  
  expect_true(grepl("Error 1", conditionMessage(error1)))
  expect_true(grepl("Error 2", conditionMessage(error2)))
  expect_false(identical(error1, error2))
})

# Test 28: abort with complex nested formatting
test_that('abort with complex nested formatting', {
  lang <- "en"
  available <- c("en", "ru", "fr")
  expect_error(
    abort(
      "Language '{lang}' not available. Use: {.val {available}}"
    ),
    "not available",
    fixed = TRUE
  )
})

# Test 29: abort requires exactly one argument
test_that('abort requires msg argument', {
  expect_error(abort(), "missing", fixed = TRUE)
})

# Test 30: abort with NULL message
test_that('abort with NULL message', {
  expect_error(abort(NULL), 'must be supplied', fixed = TRUE)
})

# Test 31: abort with numeric message (coerced to string)
test_that('abort with numeric message', {
  expect_error(
    abort(123),
    "123",
    fixed = TRUE
  )
})

# Test 32: abort with expression in message
test_that('abort with expression evaluation', {
  x <- 5
  y <- 3
  expect_error(
    abort("Sum is {x + y}"),
    "Sum is 8",
    fixed = TRUE
  )
})

# Test 33: abort message with unicode characters
test_that('abort with unicode characters', {
  expect_error(
    abort("Hello \U1F680"),
    "Hello",
    fixed = TRUE
  )
})
```

## Expression Catching

```{r function-catch}
#' Safely evaluate expression and catch errors
#'
#' @param expr Expression to evaluate
#' @param expr_name Name of expression for error messages
#'
#' @return Result of expression or throws an error
#' @keywords internal
#' @importFrom rlang is_scalar_character try_fetch
#' @importFrom cli cli_abort

catch <- function(
    expr,
    expr_name = paste(deparse(substitute(expr)), collapse = '')
) {
  if (!rlang::is_scalar_character(expr_name)) {
    cli::cli_abort(
      "{.arg expr_name} must be a character scalar.
            Got {.cls {class(expr_name)}}"
    )
  }
  
  rlang::try_fetch(
    expr,
    error = function(e) {
      abort("Error in {expr_name}: {e$message}")
    }
  )
}
```

```{r tests-catch}
# Test 1: catch executes expression successfully
test_that('catch executes expression without error', {
  expect_no_error(catch(1 + 1))
})

# Test 2: catch with variable assignment
test_that('catch evaluates expression with variables', {
  x <- 5
  result <- catch(x * 2)
  expect_equal(result, 10)
})

# Test 3: catch with function call
test_that('catch executes function call', {
  result <- catch(sqrt(16))
  expect_equal(result, 4)
})

# Test 4: catch catches errors and wraps them
test_that('catch catches error and wraps message', {
  expect_error(
    catch(stop("Original error")),
    "Error in",
    fixed = TRUE
  )
})

# Test 5: catch preserves original error message
test_that('catch preserves original error message', {
  expect_error(
    catch(stop("Original error")),
    "Original error",
    fixed = TRUE
  )
})

# Test 6: catch with custom expr_name
test_that('catch with custom expr_name parameter', {
  expect_error(
    catch(stop("Test"), expr_name = "custom_operation"),
    "Error in custom_operation",
    fixed = TRUE
  )
})

# Test 7: catch with default expr_name from deparse
test_that('catch uses deparse(substitute(expr)) for default name', {
  expect_error(
    catch(1 / "1"),
    "Error in 1",
    fixed = TRUE
  )
})

# Test 8: catch rejects non-character expr_name
test_that('catch rejects non-character expr_name', {
  expect_error(
    catch(1 + 1, expr_name = 123),
    "must be a character scalar",
    fixed = TRUE
  )
})

# Test 9: catch rejects non-scalar character expr_name
test_that('catch rejects non-scalar character expr_name', {
  expect_error(
    catch(1 + 1, expr_name = c("a", "b")),
    "must be a character scalar",
    fixed = TRUE
  )
})

# Test 10: catch with complex expression
test_that('catch evaluates complex expression', {
  result <- catch({
    x <- 10
    y <- 20
    x + y
  })
  expect_equal(result, 30)
})

# Test 11: catch with error in nested function call
test_that('catch catches error in nested function', {
  expect_error(
    catch(sqrt(as.numeric(sqrt("NA")))),
    "Error in",
    fixed = TRUE
  )
})

# Test 12: catch with NULL expr_name throws error
test_that('catch rejects NULL expr_name', {
  expect_error(
    catch(1 + 1, expr_name = NULL),
    "must be a character scalar",
    fixed = TRUE
  )
})

# Test 13: catch with empty string expr_name
test_that('catch accepts empty string expr_name', {
  result <- catch(1 + 1, expr_name = "")
  expect_equal(result, 2)
})
```

## Input Validation

```{r function-validate_fields}
#' Validate input structure and required fields
#'
#' @description
#' An internal helper to ensure that a list (typically an input or result 
#' object) contains all the necessary fields required for a specific 
#' analysis stage.
#'
#' @param inputs The list or named vector to validate.
#' @param required_fields Character vector of the field names that must 
#'   be present in `inputs`.
#' @param input_name Character scalar. A descriptive name for the 
#'   object being validated, used in error messages.
#'
#' @return Invisibly TRUE if all required fields are present.
#' @throws An error if `inputs` is NULL, if any required fields are 
#'   missing, or if `input_name` is not a character scalar.
#'
#' @keywords internal
#' @importFrom rlang is_scalar_character
#' @importFrom cli cli_abort

validate_fields <- function(inputs, required_fields, input_name = 'analysis') {
  if (!rlang::is_scalar_character(input_name)) {
    cli::cli_abort("{.arg input_name} must be a character scalar.")
  }
  
  if (is.null(inputs)) {
    abort("Input for {input_name} is empty")
  }
  
  missing_fields <- setdiff(required_fields, names(inputs))
  if (length(missing_fields) > 0) {
    abort(
      "Input for {input_name} misses required fields: {.var {missing_fields}}"
    )
  }
  
  invisible(TRUE)
}
```

```{r tests-validate_fields}
test_that("validate_fields returns invisible TRUE when all fields are present", {
  inputs <- list(texts = "some text", language = "en", targets = "target")
  required <- c("texts", "language")
  
  expect_true(validate_fields(inputs, required, "test_input"))
  expect_invisible(validate_fields(inputs, required, "test_input"))
})

test_that("validate_fields throws error when input is NULL", {
  expect_error(
    validate_fields(NULL, c("field1"), "expert data"),
    regexp = "Input for expert data is empty",
    fixed = TRUE
  )
})

test_that("validate_fields identifies missing fields", {
  inputs <- list(a = 1, b = 2)
  required <- c("a", "b", "c", "d")
  
  # The error message should list the missing fields: c, d
  expect_error(
    validate_fields(inputs, required, "content analysis results"),
    regexp = "Input for content analysis results misses required fields: `c` and `d`",
    fixed = TRUE
  )
})

test_that("validate_fields handles empty required_fields list", {
  # If nothing is required, any list (or even an empty one) should pass
  expect_true(validate_fields(list(a = 1), character(0)))
  expect_true(validate_fields(list(), character(0)))
})

test_that("validate_fields validates its own arguments", {
  expect_error(
    validate_fields(list(a = 1), "a", input_name = 123),
    regexp = "must be a character scalar.",
    fixed = TRUE
  )
})

test_that("validate_fields works with named vectors", {
  inputs <- c(field1 = "val1", field2 = "val2")
  expect_true(validate_fields(inputs, "field1"))
  
  expect_error(
    validate_fields(inputs, "field3", "input"),
    regexp = "Input for input misses required fields: `field3`",
    fixed = TRUE
  )
})
```

```{r function-validate_inputs}
#' Validate input length
#'
#' @param inputs Input vector
#' @param expected_length Expected length
#' @param input_name Informative name for error messages
#'
#' @return Invisibly TRUE or throws error
#' @keywords internal
#' @importFrom rlang is_scalar_character
#' @importFrom cli cli_abort

validate_inputs <- function(inputs, expected_length, input_name = 'Inputs') {
  if (!rlang::is_scalar_character(input_name)) {
    cli::cli_abort("{.arg input_name} must be a character scalar.")
  }
  
  if (is.null(inputs) || length(inputs) != expected_length) {
    abort("{input_name} returned unexpected results")
  }
  
  invisible(TRUE)
}
```

```{r tests-validate_inputs}
# Test 1: validate_inputs returns TRUE invisibly for valid inputs
test_that('validate_inputs returns TRUE invisibly for valid inputs', {
  result <- validate_inputs(c(1, 2, 3), expected_length = 3)
  expect_true(result)
  expect_invisible(validate_inputs(c(1, 2, 3), expected_length = 3))
})

# Test 2: validate_inputs with matching length
test_that('validate_inputs accepts inputs with matching length', {
  expect_invisible(validate_inputs(list(a = 1, b = 2), expected_length = 2))
})

# Test 3: validate_inputs rejects NULL inputs
test_that('validate_inputs rejects NULL inputs', {
  expect_error(
    validate_inputs(NULL, expected_length = 1),
    "returned unexpected results",
    fixed = TRUE
  )
})

# Test 4: validate_inputs rejects mismatched length
test_that('validate_inputs rejects mismatched length', {
  expect_error(
    validate_inputs(c(1, 2, 3), expected_length = 2),
    "returned unexpected results",
    fixed = TRUE
  )
})

# Test 5: validate_inputs with custom input_name in error message
test_that('validate_inputs includes custom input_name in error', {
  expect_error(
    validate_inputs(c(1, 2), expected_length = 3, input_name = "Results"),
    "Results returned unexpected results",
    fixed = TRUE
  )
})

# Test 6: validate_inputs rejects non-character input_name
test_that('validate_inputs rejects non-character input_name', {
  expect_error(
    validate_inputs(c(1, 2, 3), expected_length = 3, input_name = 123),
    "must be a character scalar",
    fixed = TRUE
  )
})

# Test 7: validate_inputs rejects non-scalar character input_name
test_that('validate_inputs rejects non-scalar character input_name', {
  expect_error(
    validate_inputs(c(1, 2, 3), expected_length = 3, input_name = c("a", "b")),
    "must be a character scalar",
    fixed = TRUE
  )
})

# Test 8: validate_inputs with empty vector
test_that('validate_inputs rejects empty vector when expecting non-zero length', {
  expect_error(
    validate_inputs(c(), expected_length = 1),
    "returned unexpected results",
    fixed = TRUE
  )
})
```

```{r function-validate_character}
#' Validate a character argument
#'
#' @param x Input to validate
#'
#' @return Invisibly TRUE or throws error
#' @keywords internal
#' @importFrom rlang is_character

validate_character <- function(x) {
  arg_name <- deparse(substitute(x))
  
  if (!rlang::is_character(x)) {
    abort(
      "{.arg {arg_name}} must be a character vector, got {.cls {class(x)}}"
    )
  }
  if (length(x) == 0) {
    abort("{.arg {arg_name}} cannot be empty")
  }
  if (any(is.na(x)) || any(ellmer:::trim(x) == '')) {
    abort(
      c(
        "{.arg {arg_name}} contains empty strings.",
        "i" = "consider removing them before analysis."
      )
    )
  }
  
  invisible(TRUE)
}
```

```{r tests-validate_character}
# Test 1: validate_character accepts valid character vector
test_that('validate_character accepts valid character vector', {
  result <- validate_character(c("text1", "text2"))
  expect_true(result)
  expect_invisible(validate_character(c("text1", "text2")))
})

# Test 2: validate_character accepts single character string
test_that('validate_character accepts single character string', {
  expect_invisible(validate_character("single text"))
})

# Test 3: validate_character rejects non-character input
test_that('validate_character rejects numeric input', {
  expect_error(
    validate_character(123),
    "must be a character vector",
    fixed = TRUE
  )
})

# Test 4: validate_character rejects logical input
test_that('validate_character rejects logical input', {
  expect_error(
    validate_character(TRUE),
    "must be a character vector",
    fixed = TRUE
  )
})

# Test 5: validate_character rejects list input
test_that('validate_character rejects list input', {
  expect_error(
    validate_character(list("a", "b")),
    "must be a character vector",
    fixed = TRUE
  )
})

# Test 6: validate_character rejects empty character vector
test_that('validate_character rejects empty vector', {
  expect_error(
    validate_character(character()),
    "cannot be empty",
    fixed = TRUE
  )
})

# Test 7: validate_character rejects vector with NA values
test_that('validate_character rejects vector with NA', {
  expect_error(
    validate_character(c("text", NA, "more")),
    "contains empty strings",
    fixed = TRUE
  )
})

# Test 8: validate_character rejects vector with empty strings
test_that('validate_character rejects vector with empty strings', {
  expect_error(
    validate_character(c("text", "", "more")),
    "contains empty strings",
    fixed = TRUE
  )
})

# Test 9: validate_character rejects vector with whitespace-only strings
test_that('validate_character rejects whitespace-only strings', {
  expect_error(
    validate_character(c("text", "   ", "more")),
    "contains empty strings",
    fixed = TRUE
  )
})

# Test 10: validate_character includes argument name in error message
test_that('validate_character includes arg name in error', {
  my_text <- 123
  expect_error(
    validate_character(my_text),
    "my_text",
    fixed = TRUE
  )
})

# Test 11: validate_character with mixed whitespace and empty
test_that('validate_character rejects mixed empty and whitespace', {
  expect_error(
    validate_character(c("valid", "\t", "\n")),
    "contains empty strings",
    fixed = TRUE
  )
})

# Test 12: validate_character accepts vector with special characters
test_that('validate_character accepts special characters', {
  expect_invisible(
    validate_character(c("text@123", "привіт", "\U1F680 emoji"))
  )
})

# Test 13: validate_character error message includes info about removal
test_that('validate_character error includes removal suggestion', {
  expect_error(
    validate_character(c("text", NA)),
    "consider removing them",
    fixed = TRUE
  )
})
```

```{r function-validate_stage}
#' Validate stage configuration
#'
#' @param x Stage configuration
#' @param stage_name Name of stage for error messages
#'
#' @return Unmodified `x` or throws an error
#' @keywords internal

validate_stage <- function(x, stage_name) {
  if (is.null(x)) {
    abort("{stage_name} returned NULL")
  }
  x
}
```

```{r tests-validate_stage}
# Test 1: validate_stage returns input unchanged when not NULL
test_that('validate_stage returns input unchanged', {
  input <- list(a = 1, b = 2)
  result <- validate_stage(input, stage_name = "Processing")
  expect_identical(result, input)
})

# Test 2: validate_stage rejects NULL input
test_that('validate_stage rejects NULL input', {
  expect_error(
    validate_stage(NULL, stage_name = "MyStage"),
    "MyStage returned NULL",
    fixed = TRUE
  )
})
```

## Argument Recycling

```{r function-recycle_arg}
#' Recycle argument to specified length
#'
#' @param arg Values to recycle
#' @param n Target length
#'
#' @return Recycled argument values
#' @keywords internal

recycle_arg <- function(arg, n) {
  arg_name <- deparse(substitute(arg))
  
  if (length(arg) == 1) {
    return(rep(arg, n))
  } else if (length(arg) == n) {
    return(arg)
  } else {
    abort(
      "{.arg {arg_name}} must have length 1 or {n} (same as {.arg text}). \\
            Got {length(arg)}."
    )
  }
}
```

```{r tests-recycle_arg}
# Test 1: recycle_arg replicates single value to length n
test_that('recycle_arg replicates single value', {
  result <- recycle_arg("x", n = 3)
  expect_equal(result, c("x", "x", "x"))
})

# Test 2: recycle_arg returns unchanged vector of length n
test_that('recycle_arg returns vector of matching length', {
  input <- c("a", "b", "c")
  result <- recycle_arg(input, n = 3)
  expect_identical(result, input)
})

# Test 3: recycle_arg rejects vector with mismatched length
test_that('recycle_arg rejects mismatched length', {
  expect_error(
    recycle_arg(c("a", "b"), n = 3),
    "must have length 1 or 3",
    fixed = TRUE
  )
})

# Test 4: recycle_arg includes argument name in error message
test_that('recycle_arg includes arg name in error', {
  my_vector <- c(1, 2, 3, 4)
  expect_error(
    recycle_arg(my_vector, n = 2),
    "my_vector",
    fixed = TRUE
  )
})
```

## Timing

```{r function-stage_complete}
#' Format timing message
#'
#' @param tic Start time
#' @param toc End time
#' @param msg Message prefix
#'
#' @return Formatted timing message
#' @keywords internal
#' @importFrom glue glue

stage_complete <- function(tic, toc, msg) {
  tocmsg <- paste0(round(toc - tic, 3), " seconds elapsed")
  
  if (is.null(msg) || is.na(msg) || length(msg) == 0) {
    outmsg <- tocmsg
  } else {
    outmsg <- glue::glue('\u2705 {msg} complete – {tocmsg}\n\n')
  }
  outmsg
}
```

```{r tests-stage_complete}
# Test 1: stage_complete returns elapsed time when msg is NULL
test_that('stage_complete returns elapsed time for NULL msg', {
  result <- stage_complete(tic = 0, toc = 1.5, msg = NULL)
  expect_equal(result, "1.5 seconds elapsed")
})

# Test 2: stage_complete returns elapsed time when msg is NA
test_that('stage_complete returns elapsed time for NA msg', {
  result <- stage_complete(tic = 0, toc = 2.345, msg = NA)
  expect_equal(result, "2.345 seconds elapsed")
})

# Test 3: stage_complete formats message with checkmark and time
test_that('stage_complete formats message with checkmark', {
  result <- stage_complete(tic = 0, toc = 1, msg = "Processing")
  expect_match(result, "\u2705 Processing complete – 1 seconds elapsed")
})

# Test 4: stage_complete rounds elapsed time to 3 decimal places
test_that('stage_complete rounds time to 3 decimals', {
  result <- stage_complete(tic = 0, toc = 1.23456, msg = "Test")
  expect_match(result, "1.235 seconds elapsed")
})
```

---

# Translations and Localization

## Load Translations

```{r function-load_translations}
#' Load translation dictionary
#'
#' @return List containing translations for all supported languages
#' @keywords internal

load_translations <- function() {
  yaml_file <- system.file("i18n/translations.yaml", package = "rcola")
  if (!file.exists(yaml_file)) yaml_file <- here::here("inst/i18n/translations.yaml")
  rlang::try_fetch(
    yaml::read_yaml(yaml_file),
    error = function(e) cli::cli_abort("Tranlsations not found.")
  )
}
```

```{r tests-load_translations}
test_that('load_translations is a function', {
  expect_true(inherits(load_translations, 'function'))
})

test_that('load_translations works', {
  expect_no_error(dict <- load_translations())
  expect_is(dict, 'list')
  expect_named(dict, c('en', 'uk', 'ru'))
})
```

## Available Languages

```{r function-rcola_available_languages}
#' Get available languages for rcola
#'
#' @return Character vector of supported language codes
#' @export

rcola_available_languages <- function() {
  names(load_translations())
}
```

```{r examples-rcola_available_languages}
rcola_available_languages()
```

```{r tests-rcola_available_languages}
test_that('rcola_available_languages is a function', {
  expect_true(inherits(rcola_available_languages, 'function'))
})

test_that('load_translations works', {
  expect_no_error(rcola_available_languages())
  expect_is(rcola_available_languages(), 'character')
  expect_equal(rcola_available_languages(), c('en', 'uk', 'ru'))
})
```

## Localization Helper

```{r function-l}
#' Get localised string
#'
#' @param language Target language code
#' @param caption_id Translation key
#' @param case Optional second level key (e.g., 'gen', 'dat', 'likert')
#'
#' @return Character string with translation
#' @keywords internal

l <- function(language, caption_id, case = NULL) {
  # Check lanuage
  if (!language %in% rcola_available_languages()) {
    abort(
      "Language '{language}' not available. \\
            Use: {.val {rcola_available_languages()}}"
    )
  }
  
  dict <- load_translations()
  
  # Проверка ключа
  if (!caption_id %in% names(dict[[language]])) {
    cli::cli_abort(
      "Caption ID '{caption_id}' not found ",
      "in language '{language}'"
    )
  }
  
  translation <- dict[[language]][[caption_id]]
  
  if (!is.null(case)) {
    case <- rlang::arg_match(
      case,
      c('gen', 'dat', 'categorical', 'numeric', 'likert'),
      multiple = FALSE
    )
    
    if (!is.list(translation)) {
      cli::cli_abort(
        "Case '{case}' not supported for '{caption_id}' ",
        "(expected list with cases)"
      )
    }
    
    if (!case %in% names(translation)) {
      cli::cli_abort(
        "Case '{case}' not found for '{caption_id}'"
      )
    }
    
    translation[[case]]
  } else {
    if (is.list(translation)) {
      cli::cli_abort(
        "Case required for '{caption_id}'. ",
        "Available: {.val {names(translation)}}"
      )
    }
    
    translation
  }
}
```

```{r tests-l}
# Test 1: l() returns simple string translation without case
test_that('l() returns simple string translation', {
  result <- l(language = "en", caption_id = "analysis")
  expect_equal(result, "Text analysis for")
  expect_type(result, "character")
})

# Test 2: l() returns translation with case parameter
test_that('l() returns translation with case', {
  result <- l(language = "en", caption_id = "object", case = "gen")
  expect_equal(result, "the object")
})

# Test 3: l() rejects invalid language
test_that('l() rejects invalid language', {
  expect_error(
    l(language = "fr", caption_id = "analysis"),
    "Language 'fr' not available",
    fixed = TRUE
  )
})

# Test 4: l() rejects invalid caption_id
test_that('l() rejects invalid caption_id', {
  expect_error(
    l(language = "en", caption_id = "nonexistent"),
    "Caption ID 'nonexistent' not found",
    fixed = TRUE
  )
})

# Test 5: l() rejects invalid case value
test_that('l() rejects invalid case value', {
  expect_error(
    l(language = "en", caption_id = "object", case = "invalid"),
    "must be one of",
    fixed = TRUE
  )
})

# Test 6: l() requires case for dictionary entries with cases
test_that('l() requires case for multi-case entries', {
  expect_error(
    l(language = "en", caption_id = "object"),
    "Case required for 'object'",
    fixed = TRUE
  )
})

# Test 7: l() rejects case for simple string entries
test_that('l() rejects case for simple string entries', {
  expect_error(
    l(language = "en", caption_id = "analysis", case = "gen"),
    "Case 'gen' not supported for 'analysis'",
    fixed = TRUE
  )
})

# Test 8: l() rejects missing case in multi-case entry
test_that('l() rejects missing case in entry', {
  expect_error(
    l(language = "en", caption_id = "stance_positive", case = "ablative"),
    "must be one of",
    fixed = TRUE
  )
})

# Test 9: l() works with different languages
test_that('l() works with Ukrainian translation', {
  result <- l(language = "uk", caption_id = "analysis")
  expect_equal(result, "Аналіз тексту для")
})

# Test 10: l() works with Russian translation
test_that('l() works with Russian translation', {
  result <- l(language = "ru", caption_id = "stage_1")
  expect_equal(result, "Проведение многомерного анализа")
})

# Test 11: l() returns correct case variant
test_that('l() returns correct case variant for different cases', {
  gen <- l(language = "en", caption_id = "statement", case = "gen")
  dat <- l(language = "en", caption_id = "statement", case = "dat")
  expect_equal(gen, "the statement")
  expect_equal(dat, "the following statement:")
  expect_true(gen != dat)
})

# Test 12: l() works with numeric case type
test_that('l() works with numeric case type', {
  result <- l(language = "en", caption_id = "stance_positive", case = "numeric")
  expect_equal(result, "positive, near 1")
})

# Test 13: l() works with likert case type
test_that('l() works with likert case type', {
  result <- l(language = "ru", caption_id = "stance_neutral", case = "likert")
  expect_equal(result, "нейтральная")
})
```

## Check Translations

```{r function-rcola_check_translations}
#' Check translation dictionary structure
#'
#' @description
#' Validates and displays information about the translation dictionary,
#' including available languages, number of keys, and keys with case variants.
#' Useful for verifying translation completeness and structure.
#'
#' @details
#' This function performs the following checks:
#' - Lists all available languages in the dictionary
#' - Counts total translation keys per language
#' - Identifies keys with case variants (e.g., 'gen', 'dat', 'likert')
#'
#' The function prints a formatted report to the console and returns
#' invisibly TRUE.
#'
#' @return
#' Invisibly `TRUE`. Called for side effects (printing to console).
#'
#' @examples
#' \dontrun{
#'   rcola_check_translations()
#'   # Available languages: en, uk, ru
#'   # 
#'   # Language: en
#'   #   Keys: 13
#'   #   Keys with cases: 5
#'   # 
#'   # Language: uk
#'   #   Keys: 13
#'   #   Keys with cases: 5
#'   # 
#'   # Language: ru
#'   #   Keys: 13
#'   #   Keys with cases: 5
#' }
#'
#' @keywords internal
#' @export

rcola_check_translations <- function() {
  dict <- load_translations()
  
  cat("Available languages:", paste(names(dict), collapse = ", "), "\n\n")
  
  for (language in names(dict)) {
    cat("Language:", language, "\n")
    cat("  Keys:", length(dict[[language]]), "\n")
    
    # Проверка структуры
    cases_keys <- sapply(dict[[language]], function(x) is.list(x))
    if (any(cases_keys)) {
      cat("  Keys with cases:", sum(cases_keys), "\n")
    }
    cat("\n")
  }
  
  invisible(TRUE)
}
```

```{r examples-rcola_check_translations}
rcola_check_translations()
```

```{r tests-rcola_check_translations}
test_that('rcola_check_translations returns invisibly TRUE', {
  result <- rcola_check_translations()
  expect_true(result)
  expect_invisible(rcola_check_translations())
})

test_that('rcola_check_translations prints available languages', {
  expect_output(
    rcola_check_translations(),
    "Available languages:"
  )
})

test_that('rcola_check_translations lists all languages', {
  expect_output(
    rcola_check_translations(),
    "en"
  )
  expect_output(
    rcola_check_translations(),
    "uk"
  )
  expect_output(
    rcola_check_translations(),
    "ru"
  )
})

test_that('rcola_check_translations shows key count per language', {
  expect_output(
    rcola_check_translations(),
    "Language: en"
  )
  expect_output(
    rcola_check_translations(),
    "Keys: \\d+"
  )
})

test_that('rcola_check_translations identifies keys with cases', {
  expect_output(
    rcola_check_translations(),
    "Keys with cases:"
  )
})

test_that('rcola_check_translations output is consistent', {
  output1 <- capture.output(rcola_check_translations())
  output2 <- capture.output(rcola_check_translations())
  expect_equal(output1, output2)
})

test_that('rcola_check_translations handles all languages equally', {
  output <- capture.output(rcola_check_translations())
  
  # Проверяем, что каждый язык имеет свою секцию
  en_count <- sum(grepl("Language: en", output))
  uk_count <- sum(grepl("Language: uk", output))
  ru_count <- sum(grepl("Language: ru", output))
  
  expect_equal(en_count, 1)
  expect_equal(uk_count, 1)
  expect_equal(ru_count, 1)
})
```

## Type to Term Conversion

```{r function-type_to_term}
#' Convert analysis type to human-readable term
#'
#' @param type Analysis type: "object" or "statement"
#' @param language Language code
#'
#' @return Character string with localised term
#' @keywords internal

type_to_term <- function(
    type = c('object', 'statement'),
    language = rcola_available_languages()
) {
  type <- rlang::arg_match(type, c('object', 'statement'), multiple = FALSE)
  language <- rlang::arg_match(language, rcola_available_languages(), multiple = FALSE)
  
  switch(
    type,
    object = l(language, 'object', 'dat'),
    statement = l(language, 'statement', 'dat')
  )
}
```

```{r tests-type_to_term}
# Test 1: type_to_term returns correct term for object type
test_that('type_to_term returns object term in English', {
  result <- type_to_term(type = "object", language = "en")
  expect_equal(result, "the object")
})

# Test 2: type_to_term returns correct term for statement type
test_that('type_to_term returns statement term in English', {
  result <- type_to_term(type = "statement", language = "en")
  expect_equal(result, "the following statement:")
})

# Test 3: type_to_term works with Ukrainian language
test_that('type_to_term works with Ukrainian', {
  result <- type_to_term(type = "object", language = "uk")
  expect_equal(result, "об'єкту")
})

# Test 4: type_to_term works with Russian language
test_that('type_to_term works with Russian', {
  result <- type_to_term(type = "statement", language = "ru")
  expect_equal(result, "суждению")
})

# Test 5: type_to_term rejects invalid type
test_that('type_to_term rejects invalid type', {
  expect_error(
    type_to_term(type = "invalid", language = "en"),
    "must be one of",
    fixed = TRUE
  )
})

# Test 6: type_to_term rejects invalid language
test_that('type_to_term rejects invalid language', {
  expect_error(
    type_to_term(type = "object", language = "fr"),
    "must be one of",
    fixed = TRUE
  )
})

# Test 7: type_to_term uses dative case (dat) for all terms
test_that('type_to_term uses dative case', {
  # Проверяем, что используется именно дательный падеж (dat), а не именительный (gen)
  object_dat <- type_to_term(type = "object", language = "en")
  # Для английского "the object" одинаков, но проверяем логику
  expect_type(object_dat, "character")
  expect_length(object_dat, 1)
})

# Test 8: type_to_term returns character string
test_that('type_to_term returns character string', {
  result <- type_to_term(type = "object", language = "en")
  expect_type(result, "character")
  expect_length(result, 1)
})

# Test 9: type_to_term works with all available languages
test_that('type_to_term works with all available languages', {
  available_langs <- rcola_available_languages()
  
  for (lang in available_langs) {
    result_obj <- type_to_term(type = "object", language = lang)
    result_stmt <- type_to_term(type = "statement", language = lang)
    
    expect_type(result_obj, "character")
    expect_type(result_stmt, "character")
    expect_length(result_obj, 1)
    expect_length(result_stmt, 1)
  }
})

# Test 10: type_to_term returns different terms for different types
test_that('type_to_term returns different terms for different types', {
  object_term <- type_to_term(type = "object", language = "en")
  statement_term <- type_to_term(type = "statement", language = "en")
  
  expect_false(object_term == statement_term)
})
```

---

# Scale System

## Scale Definitions

```{r function-type_stance}
#' Define stance scales for structured data extraction
#'
#' @description
#' These functions define scale types for extracting stance information
#' using `ellmer::parallel_chat_structured()`. They serve as scale definitions
#' for the ellmer structured data extraction tool.
#'
#' @details
#' These functions create type definitions compatible with ellmer's
#' structured extraction framework:
#'
#' - `type_stance_categorical()`: Defines a three-point categorical scale
#'   (Positive, Negative, Neutral)
#' - `type_stance_numeric()`: Defines a continuous numeric scale
#'   (typically -1 to 1)
#' - `type_stance_likert()`: Defines a five-point Likert scale
#'   (Strongly Disagree to Strongly Agree)
#'
#' Each function reads a description from a file using
#' `ellmer::interpolate_file()`, allowing for dynamic content insertion
#' and template rendering.
#'
#' @param description_file Path to a file containing the scale description.
#'   The file may include template variables that will be interpolated
#'   using `ellmer::interpolate_file()`.
#'
#' @return
#' A type definition object compatible with `ellmer::parallel_chat_structured()`.
#' - `type_stance_categorical()` returns an enum type with three values
#' - `type_stance_numeric()` returns a numeric type
#' - `type_stance_likert()` returns an enum type with five values
#'
#' @seealso
#' [ellmer::parallel_chat_structured()] for using these type definitions
#' in structured data extraction tasks.
#'
#' @examples
#' \dontrun{
#'   # Initialise a `Chat` object
#'   my_chat <- ellmer::chat_anthropic()
#' 
#'   # Define scale types for stance extraction
#'   categorical_scale <- type_stance_categorical("scales/categorical.txt")
#'   numeric_scale <- type_stance_numeric("scales/numeric.txt")
#'   likert_scale <- type_stance_likert("scales/likert.txt")
#'
#'   # Use in structured extraction
#'   result <- my_chat$chat_structured(
#'     "Some text",
#'     type = likert_scale
#'   )
#' }
#'
#' @keywords internal
#' @name type_stance

#' @rdname type_stance
type_stance_categorical <- function(description_file) {
  ellmer::type_enum(
    values = c('Positive', 'Negative', 'Neutral'),
    description = ellmer::interpolate_file(description_file)
  )
}

#' @rdname type_stance
type_stance_numeric <- function(description_file) {
  ellmer::type_number(
    description = ellmer::interpolate_file(description_file)
  )
}

#' @rdname type_stance
type_stance_likert <- function(description_file) {
  ellmer::type_enum(
    values = c(
      'Strongly Disagree',
      'Disagree',
      'Neutral',
      'Agree',
      'Strongly Agree'
    ),
    description = ellmer::interpolate_file(description_file)
  )
}
```

```{r tests-type_stance}
# Set-up common paths for tests
prompts_dir <- system.file('prompts', 'en', package = 'rcola')
if (!dir.exists(prompts_dir)) prompts_dir <- here::here('inst', 'prompts', 'en')

test_that('type_stance_categorical returns enum type', {
  template_file <- file.path(prompts_dir, 'description-categorical.md')
  skip_if_not(file.exists(template_file))
  result <- type_stance_categorical(template_file)
  expect_is(result, "ellmer::TypeEnum")
})

test_that('type_stance_categorical includes three values', {
  template_file <- file.path(prompts_dir, 'description-categorical.md')
  skip_if_not(file.exists(template_file))
  result <- type_stance_categorical(template_file)
  expect_equal(result@values, c('Positive', 'Negative', 'Neutral'))
})

test_that('type_stance_categorical loads description from file', {
  template_file <- file.path(prompts_dir, 'description-categorical.md')
  skip_if_not(file.exists(template_file))
  result <- type_stance_categorical(template_file)
  expect_type(result@description, "character")
  expect_length(result@description, 1)
  expect_true(nchar(result@description) > 0)
})

test_that('type_stance_numeric returns basic type', {
  template_file <- file.path(prompts_dir, 'description-numeric.md')
  skip_if_not(file.exists(template_file))
  result <- type_stance_numeric(template_file)
  expect_is(result, "ellmer::TypeBasic")
})

test_that('type_stance_numeric loads description from file', {
  template_file <- file.path(prompts_dir, 'description-numeric.md')
  skip_if_not(file.exists(template_file))
  result <- type_stance_numeric(template_file)
  expect_type(result@description, "character")
  expect_length(result@description, 1)
  expect_true(nchar(result@description) > 0)
})

test_that('type_stance_likert returns enum type', {
  template_file <- file.path(prompts_dir, 'description-likert.md')
  skip_if_not(file.exists(template_file))
  result <- type_stance_likert(template_file)
  expect_is(result, "ellmer::TypeEnum")
})

test_that('type_stance_likert includes five values', {
  template_file <- file.path(prompts_dir, 'description-likert.md')
  skip_if_not(file.exists(template_file))
  result <- type_stance_likert(template_file)
  expect_equal(
    result@values,
    c('Strongly Disagree', 'Disagree', 'Neutral', 'Agree', 'Strongly Agree')
  )
})

test_that('type_stance_likert loads description from file', {
  template_file <- file.path(prompts_dir, 'description-likert.md')
  skip_if_not(file.exists(template_file))
  result <- type_stance_likert(template_file)
  expect_type(result@description, "character")
  expect_length(result@description, 1)
  expect_true(nchar(result@description) > 0)
})

test_that('type_stance_categorical and likert have different values', {
  cat_file <- file.path(prompts_dir, 'description-categorical.md')
  likert_file <- file.path(prompts_dir, 'description-likert.md')
  skip_if_not(file.exists(cat_file) && file.exists(likert_file))
  
  cat_result <- type_stance_categorical(cat_file)
  likert_result <- type_stance_likert(likert_file)
  
  expect_length(cat_result@values, 3)
  expect_length(likert_result@values, 5)
  expect_false(cat_result@description == likert_result@description)
})

test_that('type_stance_categorical handles missing file', {
  template_file <- file.path(prompts_dir, 'nonexistent.md')
  skip_if(file.exists(template_file))
  expect_error(
    type_stance_categorical(template_file),
    class = "error"
  )
})

test_that('type_stance_numeric handles missing file', {
  template_file <- file.path(prompts_dir, 'nonexistent.md')
  skip_if(file.exists(template_file))
  expect_error(
    type_stance_numeric(template_file),
    class = "error"
  )
})

test_that('type_stance_likert handles missing file', {
  template_file <- file.path(prompts_dir, 'nonexistent.md')
  skip_if(file.exists(template_file))
  expect_error(
    type_stance_likert(template_file),
    class = "error"
  )
})
```

## Data Schema

```{r function-type_stance_analysis}
#' Define stance analysis schema for structured extraction
#'
#' @description
#' Creates a complete schema for stance analysis that combines a stance scale
#' with an explanation field. This schema is designed for use with
#' `ellmer::parallel_chat_structured()` to extract structured stance analysis results.
#'
#' @details
#' This function creates a composite type object containing:
#' - **stance:** A scale type (categorical, numeric, or Likert) for the stance value
#' - **explanation:** A string field for the stance classification explanation
#'
#' The function supports three scale types:
#' - `"numeric"`: Continuous scale (typically -1 to 1)
#' - `"likert"`: Five-point Likert scale (Strongly Disagree to Strongly Agree)
#' - `"categorical"`: Three-point categorical scale (Positive, Negative, Neutral)
#'
#' The explanation field description is automatically localised based on the
#' provided language code using the `'explanation_description'` translation key.
#'
#' @param language Language code for localisation (e.g., "en", "uk", "ru").
#'   Must be one of the available languages in the rcola translation dictionary.
#'
#' @param description_file Path to a file containing the stance scale description.
#'   The file may include template variables that will be interpolated
#'   using `ellmer::interpolate_file()`.
#'
#' @param scale Type of stance scale to use. One of:
#'   - `"numeric"`: Continuous numeric scale
#'   - `"likert"`: Five-point Likert scale
#'   - `"categorical"`: Three-point categorical scale (default)
#'
#' @return
#' A type object compatible with `ellmer::parallel_chat_structured()`.
#' The returned object is a composite type containing:
#' - `stance`: A scale type definition
#' - `explanation`: A string type definition with localised description
#'
#' @seealso
#' [type_stance_categorical()], [type_stance_numeric()], [type_stance_likert()]
#' for individual scale definitions.
#'
#' [ellmer::parallel_chat_structured()] for using this schema in
#' structured data extraction tasks.
#'
#' [l()] for translation lookup.
#'
#' @examples
#' \dontrun{
#'   # Create schema for categorical stance analysis in English
#'   schema <- type_stance_analysis(
#'     language = "en",
#'     description_file = "inst/prompts/en/description-categorical.md",
#'     scale = "categorical"
#'   )
#'
#'   # Create schema for numeric stance analysis in Russian
#'   schema <- type_stance_analysis(
#'     language = "ru",
#'     description_file = "inst/prompts/ru/description-numeric.md",
#'     scale = "numeric"
#'   )
#'
#'   # Use in structured extraction
#'   result <- ellmer::parallel_chat_structured(
#'     "Some text",          
#'     chat = my_chat,
#'     schema = schema
#'   )
#' }
#'
#' @keywords internal
type_stance_analysis <- function(language, description_file, scale = "categorical") {
  # Validate scale parameter
  scale <- rlang::arg_match(scale, c("numeric", "likert", "categorical"))
  
  # A schema for the structured output
  type_stance <- switch(
    scale,
    numeric = type_stance_numeric(description_file),
    likert = type_stance_likert(description_file),
    type_stance_categorical(description_file)
  )
  
  type_object(
    stance = type_stance,
    explanation = type_string(
      description = l(language, 'explanation_description')
    )
  )
}
```

```{r tests-type_stance_analysis}
# Set-up common paths for tests
prompts_dir <- system.file('prompts', 'en', package = 'rcola')
if (!dir.exists(prompts_dir)) prompts_dir <- here::here('inst', 'prompts', 'en')

test_that('type_stance_analysis returns object type', {
  template_file <- file.path(prompts_dir, 'description-categorical.md')
  skip_if_not(file.exists(template_file))
  result <- type_stance_analysis(
    language = "en",
    description_file = template_file,
    scale = "categorical"
  )
  expect_is(result, "ellmer::TypeObject")
})

test_that('type_stance_analysis includes stance and explanation fields', {
  template_file <- file.path(prompts_dir, 'description-categorical.md')
  skip_if_not(file.exists(template_file))
  result <- type_stance_analysis(
    language = "en",
    description_file = template_file,
    scale = "categorical"
  )
  expect_true("stance" %in% names(result@properties))
  expect_true("explanation" %in% names(result@properties))
})

test_that('type_stance_analysis creates categorical scale by default', {
  template_file <- file.path(prompts_dir, 'description-categorical.md')
  skip_if_not(file.exists(template_file))
  result <- type_stance_analysis(
    language = "en",
    description_file = template_file
  )
  expect_is(result@properties$stance, "ellmer::TypeEnum")
  expect_equal(result@properties$stance@values, c('Positive', 'Negative', 'Neutral'))
})

test_that('type_stance_analysis creates numeric scale when specified', {
  template_file <- file.path(prompts_dir, 'description-numeric.md')
  skip_if_not(file.exists(template_file))
  result <- type_stance_analysis(
    language = "en",
    description_file = template_file,
    scale = "numeric"
  )
  expect_is(result@properties$stance, "ellmer::TypeBasic")
  expect_equal(result@properties$stance@type, "number")
})

test_that('type_stance_analysis creates Likert scale when specified', {
  template_file <- file.path(prompts_dir, 'description-likert.md')
  skip_if_not(file.exists(template_file))
  result <- type_stance_analysis(
    language = "en",
    description_file = template_file,
    scale = "likert"
  )
  expect_is(result@properties$stance, "ellmer::TypeEnum")
  expect_equal(
    result@properties$stance@values,
    c('Strongly Disagree', 'Disagree', 'Neutral', 'Agree', 'Strongly Agree')
  )
})

test_that('type_stance_analysis includes localised explanation description', {
  template_file <- file.path(prompts_dir, 'description-categorical.md')
  skip_if_not(file.exists(template_file))
  result <- type_stance_analysis(
    language = "en",
    description_file = template_file
  )
  expect_is(result@properties$explanation, "ellmer::TypeBasic")
  expect_equal(result@properties$explanation@type, "string")
  expect_type(result@properties$explanation@description, "character")
  expect_true(nchar(result@properties$explanation@description) > 0)
})

test_that('type_stance_analysis uses correct language for explanation', {
  template_file <- file.path(prompts_dir, 'description-categorical.md')
  skip_if_not(file.exists(template_file))
  
  result_en <- type_stance_analysis(
    language = "en",
    description_file = template_file
  )
  result_uk <- type_stance_analysis(
    language = "uk",
    description_file = template_file
  )
  
  expect_false(
    result_en@properties$explanation@description ==
      result_uk@properties$explanation@description
  )
})

test_that('type_stance_analysis rejects invalid scale', {
  template_file <- file.path(prompts_dir, 'description-categorical.md')
  skip_if_not(file.exists(template_file))
  expect_error(
    type_stance_analysis(
      language = "en",
      description_file = template_file,
      scale = "invalid"
    ),
    "must be one of",
    fixed = TRUE
  )
})

test_that('type_stance_analysis rejects invalid language', {
  template_file <- file.path(prompts_dir, 'description-categorical.md')
  skip_if_not(file.exists(template_file))
  expect_error(
    type_stance_analysis(
      language = "fr",
      description_file = template_file
    ),
    "Language 'fr' not available",
    fixed = TRUE
  )
})

test_that('type_stance_analysis handles missing description file', {
  template_file <- file.path(prompts_dir, 'nonexistent.md')
  skip_if(file.exists(template_file))
  expect_error(
    type_stance_analysis(
      language = "en",
      description_file = template_file
    ),
    class = "error"
  )
})

test_that('type_stance_analysis works with all available languages', {
  template_file <- file.path(prompts_dir, 'description-categorical.md')
  skip_if_not(file.exists(template_file))
  
  available_langs <- rcola_available_languages()
  
  for (lang in available_langs) {
    result <- type_stance_analysis(
      language = lang,
      description_file = template_file
    )
    expect_is(result, "ellmer::TypeObject")
    expect_true("stance" %in% names(result@properties))
    expect_true("explanation" %in% names(result@properties))
  }
})

test_that('type_stance_analysis works with all scale types', {
  cat_file <- file.path(prompts_dir, 'description-categorical.md')
  num_file <- file.path(prompts_dir, 'description-numeric.md')
  likert_file <- file.path(prompts_dir, 'description-likert.md')
  
  skip_if_not(
    file.exists(cat_file) &&
      file.exists(num_file) &&
      file.exists(likert_file)
  )
  
  result_cat <- type_stance_analysis(
    language = "en",
    description_file = cat_file,
    scale = "categorical"
  )
  result_num <- type_stance_analysis(
    language = "en",
    description_file = num_file,
    scale = "numeric"
  )
  result_likert <- type_stance_analysis(
    language = "en",
    description_file = likert_file,
    scale = "likert"
  )
  
  expect_is(result_cat@properties$stance, "ellmer::TypeEnum")
  expect_equal(result_num@properties$stance@type, "number")
  expect_is(result_likert@properties$stance, "ellmer::TypeEnum")
})
```

---

# Prompt and Task Preparation

## Prompts

```{r function-prompts_prepare}
#' Prepare system and user prompts for a given role
#'
#' @description
#' Loads and interpolates system and user prompt templates for a specific role
#' in the stance analysis workflow. This function prepares the prompts that
#' guide the language model's behaviour for different analysis stages.
#'
#' @details
#' The function supports five distinct roles in the stance analysis pipeline:
#'
#' - **linguist**: Performs linguistic analysis of statements
#' - **domain**: Provides domain-specific expertise and context
#' - **interpreter**: Interprets and clarifies statements
#' - **debater**: Presents arguments and counter-arguments
#' - **judger**: Makes final stance judgements (supports additional options)
#'
#' For the `judger` role, additional instructions can be included:
#' - `add_reference_resolution`: Include instructions for resolving references
#' - `add_statement_resolution`: Include instructions for resolving statements
#'
#' The function performs the following operations:
#' 1. Validates the role and language parameters
#' 2. Loads template files for the specified role
#' 3. Interpolates templates with provided variables (via `...`)
#' 4. Validates that prompts are non-empty after interpolation
#'
#' Templates are loaded using `ellmer::interpolate_file()`, which allows
#' for dynamic content insertion and template variable rendering.
#'
#' @param role Role for which to prepare prompts. One of:
#'   - `"linguist"`: Linguistic analysis
#'   - `"domain"`: Domain expertise
#'   - `"interpreter"`: Statement interpretation
#'   - `"debater"`: Argument presentation
#'   - `"judger"`: Final judgement
#'
#' @param templates A list containing template file paths. Expected keys:
#'   - `system-{role}`: Path to system prompt template
#'   - `user-{role}`: Path to user prompt template
#'   - `reference_resolution`: Path to reference resolution instructions (for judger)
#'   - `statement_resolution`: Path to statement resolution instructions (for judger)
#'   - `scale_guide`: Path to scale guide (for judger)
#'
#' @param language Language code for localisation (e.g., "en", "uk", "ru").
#'   Defaults to the first available language. Must be one of the available
#'   languages in the rcola translation dictionary.
#'
#' @param add_reference_resolution Logical. For the `judger` role only.
#'   If `TRUE`, includes instructions for resolving references in statements.
#'   Defaults to `FALSE`.
#'
#' @param add_statement_resolution Logical. For the `judger` role only.
#'   If `TRUE`, includes instructions for resolving statements.
#'   Defaults to `FALSE`.
#'
#' @param ... Additional named arguments passed to `ellmer::interpolate_file()`.
#'   These allow for template variable substitution (e.g., `target = "climate change"`).
#'
#' @return
#' A list with two elements:
#' - `system`: Character string containing the system prompt
#' - `task`: Character string containing the user prompt
#'
#' Both strings are non-empty after interpolation.
#'
#' @seealso
#' [ellmer::interpolate_file()] for template interpolation details.
#'
#' @examples
#' \dontrun{
#'   # Prepare prompts for linguistic analysis
#'   templates <- list(
#'     `system-linguist` = "inst/prompts/en/system-linguist.md",
#'     `user-linguist` = "inst/prompts/en/user-linguist.md"
#'   )
#'
#'   prompts <- prompts_prepare(
#'     role = "linguist",
#'     templates = templates,
#'     language = "en",
#'     target = "climate change",
#'     text = c("some text", "and more"),
#'     target_type = "the object"
#'   )
#'
#'   # Prepare prompts for domain expert
#'   templates <- list(
#'     `system-domain` = "inst/prompts/en/system-domain.md",
#'     `user-domain` = "inst/prompts/en/user-domain.md"
#'   )
#'
#'   prompts <- prompts_prepare(
#'     role = "domain",
#'     templates = templates,
#'     language = "en",
#'     target = "artificial intelligence",
#'     text = "AI will replace all human jobs",
#'     target_type = "the object",
#'     domain = "technology and employment expert"
#'   )
#'
#'   # Prepare prompts for judger with additional options
#'   templates <- list(
#'     `system-judger` = "inst/prompts/en/system-judger.md",
#'     `user-judger` = "inst/prompts/en/user-judger.md",
#'     reference_resolution = "inst/prompts/en/reference-resolution.md",
#'     statement_resolution = "inst/prompts/en/statement-resolution.md",
#'     scale_guide = "inst/prompts/en/guide-likert.md"
#'   )
#'
#'   prompts <- prompts_prepare(
#'     role = "judger",
#'     templates = templates,
#'     language = "en",
#'     add_reference_resolution = TRUE,
#'     add_statement_resolution = TRUE,
#'     target = "Climate change is a real concern",
#'     text = "Global warming is a serious threat",
#'     target_type = "the statement",
#'     FavourResponse = "This text is almost a paraphrase of the target statement",
#'     AgainstResponse = "This is most probably a parody of climate speeches.",
#'     NeutralResponse = "Context is missing and we should be sceptical" 
#'   )
#' }
#'
#' @keywords internal
prompts_prepare <- function(
    role,
    templates,
    language = rcola_available_languages(),
    add_reference_resolution = FALSE,
    add_statement_resolution = FALSE,
    ...
) {
  role <- rlang::arg_match(
    role,
    values = c('linguist', 'domain', 'interpreter', 'debater', 'judger'),
    multiple = FALSE
  )
  language <- rlang::arg_match(
    language,
    rcola_available_languages(),
    multiple = FALSE
  )
  
  # A lot of work for a judger
  if (role == 'judger') {
    if (add_reference_resolution) {
      reference_instruction <- ellmer::interpolate_file(
        templates$reference_resolution
      )
    } else {
      reference_instruction <- ''
    }
    
    if (add_statement_resolution) {
      statement_instruction <- ellmer::interpolate_file(
        templates$statement_resolution
      )
    } else {
      # Default settings
      statement_instruction <- ''
    }
    
    scale_guide <- ellmer::interpolate_file(templates$scale_guide)
  }
  
  template_system <- templates[[glue::glue('system-{role}')]]
  template_user <- templates[[glue::glue('user-{role}')]]
  
  # Check for file existence
  if (!file.exists(template_system)) {
    abort(
      c(
        "System template not found: {.file {template_system}}",
        "i" = "Available roles: {.val linguist}, {.val domain}, {.val interpreter}, {.val debater}, {.val judger}"
      )
    )
  }
  if (!file.exists(template_user)) {
    abort(
      c(
        "User template not found: {.file {template_user}}",
        "i" = "Available roles: {.val linguist}, {.val domain}, {.val interpreter}, {.val debater}, {.val judger}"
      )
    )
  }
  
  prompts <- list(
    system = ellmer::interpolate_file(template_system, ...),
    task = ellmer::interpolate_file(template_user, ...)
  )
  
  # Check for user prompt existence
  if (is.null(prompts$task)) {
    abort("User prompt is NULL for role {.val {role}}")
  }
  
  # Check for empty strings
  if (any(nchar(prompts$system) == 0)) {
    abort("System prompt is empty after interpolation for role {.val {role}}")
  }
  if (any(nchar(prompts$task) == 0)) {
    abort("User prompt is empty after interpolation for role {.val {role}}")
  }
  
  prompts
}
```

```{r tests-prompts_prepare}
# Set-up common paths for tests
prompts_dir <- system.file('prompts', 'en', package = 'rcola')
if (!dir.exists(prompts_dir)) prompts_dir <- here::here('inst', 'prompts', 'en')

test_that('prompts_prepare returns list with system and task', {
  templates <- list(
    `system-linguist` = file.path(prompts_dir, 'system-linguist.md'),
    `user-linguist` = file.path(prompts_dir, 'user-linguist.md')
  )
  skip_if_not(all(file.exists(unlist(templates))))
  
  result <- prompts_prepare(
    role = "linguist",
    templates = templates,
    language = "en",
    target = "climate change",
    text = c("some text", "and more"),
    target_type = "the object"
  )
  
  expect_type(result, "list")
  expect_named(result, c("system", "task"))
})

test_that('prompts_prepare returns non-empty strings', {
  templates <- list(
    `system-linguist` = file.path(prompts_dir, 'system-linguist.md'),
    `user-linguist` = file.path(prompts_dir, 'user-linguist.md')
  )
  skip_if_not(all(file.exists(unlist(templates))))
  
  result <- prompts_prepare(
    role = "linguist",
    templates = templates,
    language = "en",
    target = "climate change",
    text = c("some text", "and more"),
    target_type = "the object"
  )
  
  expect_type(result$system, "character")
  expect_type(result$task, "character")
  expect_true(nchar(result$system) > 0)
  expect_true(all(nchar(result$task) > 0))
})

test_that('prompts_prepare works with character vector for text', {
  templates <- list(
    `system-linguist` = file.path(prompts_dir, 'system-linguist.md'),
    `user-linguist` = file.path(prompts_dir, 'user-linguist.md')
  )
  skip_if_not(all(file.exists(unlist(templates))))
  
  texts <- c("some text", "and more")
  
  result <- prompts_prepare(
    role = "linguist",
    templates = templates,
    language = "en",
    target = "climate change",
    text = texts,
    target_type = "the object"
  )
  
  expect_type(result, "list")
  expect_length(result$system, 1)
  expect_length(result$task, length(texts))
})

test_that('prompts_prepare works with all valid roles', {
  roles <- c('linguist', 'domain', 'interpreter', 'debater', 'judger')
  
  templates <- list(
    `system-linguist` = file.path(prompts_dir, 'system-linguist.md'),
    `user-linguist` = file.path(prompts_dir, 'user-linguist.md'),
    `system-domain` = file.path(prompts_dir, 'system-domain.md'),
    `user-domain` = file.path(prompts_dir, 'user-domain.md'),
    `system-interpreter` = file.path(prompts_dir, 'system-interpreter.md'),
    `user-interpreter` = file.path(prompts_dir, 'user-interpreter.md'),
    `system-debater` = file.path(prompts_dir, 'system-debater.md'),
    `user-debater` = file.path(prompts_dir, 'user-debater.md'),
    `system-judger` = file.path(prompts_dir, 'system-judger.md'),
    `user-judger` = file.path(prompts_dir, 'user-judger.md'),
    reference_resolution = file.path(prompts_dir, 'reference-resolution.md'),
    statement_resolution = file.path(prompts_dir, 'statement-resolution.md'),
    scale_guide = file.path(prompts_dir, 'guide-likert.md')
  )
  skip_if_not(all(file.exists(unlist(templates))))
  
  texts <- c("some text", "and more")
  
  for (r in roles) {
    result <- prompts_prepare(
      role = r,
      templates = templates,
      language = "en",
      target = "climate change",
      text = texts,
      target_type = "the object",
      domain = "content analysis expert",
      stance = "neutral",
      LingResponse = "This is just a placeholder phrase",
      ExpertResponse = "Just a dummy text document with no much sense",
      UserResponse = "Every second real tweet be like",
      FavourResponse = "I have a feeling it is positive",
      AgainstResponse = "Most likely, this is some kind of malicious mockery",
      NeutralResponse = "Context is missing and we should be sceptical"
    )
    
    expect_type(result, "list")
    expect_named(result, c("system", "task"))
    expect_type(result$system, "character")
    expect_type(result$task, "character")
    expect_length(result$system, 1)
    expect_length(result$task, length(texts))
    expect_true(nchar(result$system) > 0)
    expect_true(all(nchar(result$task) > 0))
  }
})

test_that('prompts_prepare rejects invalid role', {
  templates <- list(
    `system-linguist` = file.path(prompts_dir, 'system-linguist.md'),
    `user-linguist` = file.path(prompts_dir, 'user-linguist.md')
  )
  skip_if_not(all(file.exists(unlist(templates))))
  
  expect_error(
    prompts_prepare(
      role = "invalid_role",
      templates = templates,
      language = "en",
      target = "climate change",
      text = c("some text", "and more"),
      target_type = "the object"
    ),
    "must be one of",
    fixed = TRUE
  )
})

test_that('prompts_prepare rejects invalid language', {
  templates <- list(
    `system-linguist` = file.path(prompts_dir, 'system-linguist.md'),
    `user-linguist` = file.path(prompts_dir, 'user-linguist.md')
  )
  skip_if_not(all(file.exists(unlist(templates))))
  
  expect_error(
    prompts_prepare(
      role = "linguist",
      templates = templates,
      language = "fr",
      target = "climate change",
      text = c("some text", "and more"),
      target_type = "the object"
    ),
    "must be one of",
    fixed = TRUE
  )
})

test_that('prompts_prepare handles missing system template', {
  templates <- list(
    `system-linguist` = file.path(prompts_dir, 'nonexistent-system.md'),
    `user-linguist` = file.path(prompts_dir, 'user-linguist.md')
  )
  skip_if(file.exists(templates$`system-linguist`))
  
  expect_error(
    prompts_prepare(
      role = "linguist",
      templates = templates,
      language = "en",
      target = "climate change",
      text = c("some text", "and more"),
      target_type = "the object"
    ),
    "System template not found",
    fixed = TRUE
  )
})

test_that('prompts_prepare handles missing user template', {
  templates <- list(
    `system-linguist` = file.path(prompts_dir, 'system-linguist.md'),
    `user-linguist` = file.path(prompts_dir, 'nonexistent-user.md')
  )
  skip_if(file.exists(templates$`user-linguist`))
  
  expect_error(
    prompts_prepare(
      role = "linguist",
      templates = templates,
      language = "en",
      target = "climate change",
      text = c("some text", "and more"),
      target_type = "the object"
    ),
    "User template not found",
    fixed = TRUE
  )
})

test_that('prompts_prepare works with domain role and domain parameter', {
  templates <- list(
    `system-domain` = file.path(prompts_dir, 'system-domain.md'),
    `user-domain` = file.path(prompts_dir, 'user-domain.md')
  )
  skip_if_not(all(file.exists(unlist(templates))))
  
  result <- prompts_prepare(
    role = "domain",
    templates = templates,
    language = "en",
    target = "artificial intelligence",
    text = "AI will replace all human jobs",
    target_type = "the object",
    domain = "technology and employment expert"
  )
  
  expect_type(result, "list")
  expect_true(nchar(result$system) > 0)
  expect_true(nchar(result$task) > 0)
})

test_that('prompts_prepare works with judger role and no additional options', {
  templates <- list(
    `system-judger` = file.path(prompts_dir, 'system-judger.md'),
    `user-judger` = file.path(prompts_dir, 'user-judger.md'),
    reference_resolution = file.path(prompts_dir, 'reference-resolution.md'),
    statement_resolution = file.path(prompts_dir, 'statement-resolution.md'),
    scale_guide = file.path(prompts_dir, 'guide-likert.md')
  )
  skip_if_not(all(file.exists(unlist(templates))))
  
  texts <- c("some text", "and more")
  
  result <- prompts_prepare(
    role = "judger",
    templates = templates,
    language = "en",
    add_reference_resolution = FALSE,
    add_statement_resolution = FALSE,
    target = "content analysis",
    text = texts,
    target_type = "the object",
    FavourResponse = "I have a feeling it is positive",
    AgainstResponse = "Most likely, this is some kind of malicious mockery",
    NeutralResponse = "Context is missing and we should be sceptical"
  )
  
  expect_type(result, "list")
  expect_true(nchar(result$system) > 0)
  expect_true(all(nchar(result$task) > 0))
  expect_length(result$system, 1)
  expect_length(result$task, length(texts))
  expect_no_match(result$system, "reference_instruction")
  expect_no_match(result$system, "statement_instruction")
  expect_no_match(result$system, "Resolve a reference if")
  expect_no_match(result$system, "Rules for Statements")
})

test_that('prompts_prepare works with judger role and reference resolution', {
  templates <- list(
    `system-judger` = file.path(prompts_dir, 'system-judger.md'),
    `user-judger` = file.path(prompts_dir, 'user-judger.md'),
    reference_resolution = file.path(prompts_dir, 'reference-resolution.md'),
    statement_resolution = file.path(prompts_dir, 'statement-resolution.md'),
    scale_guide = file.path(prompts_dir, 'guide-likert.md')
  )
  skip_if_not(all(file.exists(unlist(templates))))
  
  texts <- c("some text", "and more")
  
  result <- prompts_prepare(
    role = "judger",
    templates = templates,
    language = "en",
    add_reference_resolution = TRUE,
    add_statement_resolution = FALSE,
    target = "content analysis",
    text = texts,
    target_type = "the object",
    FavourResponse = "I have a feeling it is positive",
    AgainstResponse = "Most likely, this is some kind of malicious mockery",
    NeutralResponse = "Context is missing and we should be sceptical"
  )
  
  expect_type(result, "list")
  expect_true(nchar(result$system) > 0)
  expect_true(all(nchar(result$task) > 0))
  expect_length(result$system, 1)
  expect_length(result$task, length(texts))
  expect_no_match(result$system, "reference_instruction")
  expect_no_match(result$system, "statement_instruction")
  expect_match(result$system, "Resolve a reference if")
  expect_no_match(result$system, "Rules for Statements")
})

test_that('prompts_prepare works with judger role and statement resolution', {
  templates <- list(
    `system-judger` = file.path(prompts_dir, 'system-judger.md'),
    `user-judger` = file.path(prompts_dir, 'user-judger.md'),
    reference_resolution = file.path(prompts_dir, 'reference-resolution.md'),
    statement_resolution = file.path(prompts_dir, 'statement-resolution.md'),
    scale_guide = file.path(prompts_dir, 'guide-likert.md')
  )
  skip_if_not(all(file.exists(unlist(templates))))
  
  texts <- c("some text", "and more")
  
  result <- prompts_prepare(
    role = "judger",
    templates = templates,
    language = "en",
    add_reference_resolution = FALSE,
    add_statement_resolution = TRUE,
    target = "content analysis",
    text = texts,
    target_type = "the object",
    FavourResponse = "I have a feeling it is positive",
    AgainstResponse = "Most likely, this is some kind of malicious mockery",
    NeutralResponse = "Context is missing and we should be sceptical"
  )
  
  expect_type(result, "list")
  expect_true(nchar(result$system) > 0)
  expect_true(all(nchar(result$task) > 0))
  expect_length(result$system, 1)
  expect_length(result$task, length(texts))
  expect_no_match(result$system, "reference_instruction")
  expect_no_match(result$system, "statement_instruction")
  expect_no_match(result$system, "Resolve a reference if")
  expect_match(result$system, "Rules for Statements")
})

test_that('prompts_prepare works with judger role and both options', {
  templates <- list(
    `system-judger` = file.path(prompts_dir, 'system-judger.md'),
    `user-judger` = file.path(prompts_dir, 'user-judger.md'),
    reference_resolution = file.path(prompts_dir, 'reference-resolution.md'),
    statement_resolution = file.path(prompts_dir, 'statement-resolution.md'),
    scale_guide = file.path(prompts_dir, 'guide-likert.md')
  )
  skip_if_not(all(file.exists(unlist(templates))))
  
  texts <- c("some text", "and more")
  
  result <- prompts_prepare(
    role = "judger",
    templates = templates,
    language = "en",
    add_reference_resolution = TRUE,
    add_statement_resolution = TRUE,
    target = "content analysis",
    text = texts,
    target_type = "the object",
    FavourResponse = "I have a feeling it is positive",
    AgainstResponse = "Most likely, this is some kind of malicious mockery",
    NeutralResponse = "Context is missing and we should be sceptical"
  )
  
  expect_type(result, "list")
  expect_true(nchar(result$system) > 0)
  expect_true(all(nchar(result$task) > 0))
  expect_length(result$system, 1)
  expect_length(result$task, length(texts))
  expect_no_match(result$system, "reference_instruction")
  expect_no_match(result$system, "statement_instruction")
  expect_match(result$system, "Resolve a reference if")
  expect_match(result$system, "Rules for Statements")
})

test_that('prompts_prepare returns consistent results for same inputs', {
  templates <- list(
    `system-linguist` = file.path(prompts_dir, 'system-linguist.md'),
    `user-linguist` = file.path(prompts_dir, 'user-linguist.md')
  )
  skip_if_not(all(file.exists(unlist(templates))))
  
  result1 <- prompts_prepare(
    role = "linguist",
    templates = templates,
    language = "en",
    target = "content analysis",
    text = c("some text", "and more"),
    target_type = "the object"
  )
  result2 <- prompts_prepare(
    role = "linguist",
    templates = templates,
    language = "en",
    target = "content analysis",
    text = c("some text", "and more"),
    target_type = "the object"
  )
  
  expect_identical(result1$system, result2$system)
  expect_identical(result1$task, result2$task)
})

test_that('prompts_prepare produces different results for different target', {
  templates <- list(
    `system-linguist` = file.path(prompts_dir, 'system-linguist.md'),
    `user-linguist` = file.path(prompts_dir, 'user-linguist.md')
  )
  skip_if_not(all(file.exists(unlist(templates))))
  
  result1 <- prompts_prepare(
    role = "linguist",
    templates = templates,
    language = "en",
    target = "climate change",
    text = c("some text", "and more"),
    target_type = "the object"
  )
  result2 <- prompts_prepare(
    role = "linguist",
    templates = templates,
    language = "en",
    target = "artificial intelligence",
    text = c("some text", "and more"),
    target_type = "the object"
  )
  
  expect_identical(result1$system, result2$system)
  expect_disjoint(result1$task, result2$task)
})
```

## Tasks

```{r function-tasks_validate}
#' Validate task configuration
#'
#' @description
#' Internal helper to ensure that the generated tasks (user prompts) meet
#' the expected criteria after interpolation. It checks for correct length,
#' data type, and ensures no empty strings were produced.
#'
#' @param tasks A character vector of interpolated tasks.
#' @param expected_length The number of tasks expected (usually matching the number of input texts).
#'
#' @return Invisibly TRUE if validation passes.
#' @throws An error if tasks are not a character vector, have incorrect length, or contain empty strings.
#'
#' @keywords internal

tasks_validate <- function(tasks, expected_length) {
  rlang::try_fetch(
    ellmer:::check_number_whole(expected_length, min = 1),
    error = function(e) abort("Wrong {.arg expected_length} argument.")
  )
  
  if (any(is.na(tasks))) {
    abort("User prompt list contains missing values")
  }
  
  if (length(tasks) != expected_length || !is.character(tasks)) {
    abort(
      c(
        "User prompt interpolation returned unexpected results",
        "i" = "Consider checking variable placeholders in user templates"
      )
    )
  }
  
  if (any(nchar(tasks) == 0)) {
    abort(
      c(
        "User prompt is empty after interpolation",
        "i" = "Check that all required variables are provided"
      )
    )
  }
  
  invisible(TRUE)
}
```

```{r tests-tasks_validate}
test_that("tasks_validate returns invisible TRUE on valid input", {
  tasks <- c("Content analysis task 1", "Content analysis task 2")
  expect_true(tasks_validate(tasks, 2))
  expect_invisible(tasks_validate(tasks, 2))
})

test_that("tasks_validate throws error on incorrect length", {
  tasks <- c("Content analysis task 1", "Content analysis task 2")
  expect_error(
    tasks_validate(tasks, 3),
    "User prompt interpolation returned unexpected results",
    fixed = TRUE
  )
  
  expect_error(
    tasks_validate(tasks, 1),
    "User prompt interpolation returned unexpected results",
    fixed = TRUE
  )
})

test_that("tasks_validate throws error if tasks is not character", {
  expect_error(
    tasks_validate(list("Content analysis task 1", "Content analysis task 2"), 2),
    "User prompt interpolation returned unexpected results",
    fixed = TRUE
  )
  
  expect_error(
    tasks_validate(c(1, 2), 2),
    "User prompt interpolation returned unexpected results",
    fixed = TRUE
  )
})

test_that("tasks_validate throws error on empty strings", {
  tasks_with_empty <- c("Valid task", "")
  expect_error(
    tasks_validate(tasks_with_empty, 2),
    "User prompt is empty after interpolation",
    fixed = TRUE
  )
})

test_that("tasks_validate throws error on NA values", {
  # nchar(NA) is NA, which fails the if(any(nchar...)) check or behaves unexpectedly
  tasks_with_na <- c("Valid task", NA_character_)
  expect_error(
    tasks_validate(tasks_with_na, 2),
    "contains missing",
    fixed = TRUE
  )
})

test_that("tasks_validate works with single task", {
  expect_true(tasks_validate("Single task", 1))
})

test_that("tasks_validate rejects zero length expectation", {
  expect_error(tasks_validate(character(0), 0), "Wrong", fixed = TRUE)
})

test_that("tasks_validate rejects fractional `expected_length`", {
  expect_error(tasks_validate("A task", 1.5), "Wrong", fixed = TRUE)
})
```

```{r function-tasks_prepare}
#' Prepare chat objects and tasks for execution
#'
#' @description
#' Organises the final structure for parallel execution by cloning the base
#' chat object and pairing it with the corresponding system and user prompts.
#'
#' @details
#' This function handles two main scenarios for system prompts:
#' 1. A single system prompt applied to all tasks.
#' 2. A vector of system prompts matching the number of texts (e.g., when 
#'    different domain experts are assigned to different texts).
#'
#' The function clones the `chat_base` for each system prompt to ensure 
#' environment isolation during parallel processing. It also performs 
#' validation on the user tasks using [tasks_validate()].
#'
#' @param chat_base An `ellmer::Chat` object used as a template.
#' @param prompts A list containing `system` (character vector) and 
#'   `task` (character vector) prompts, typically from [prompts_prepare()].
#' @param n_texts Integer. The expected number of texts/tasks to process.
#'
#' @return A list with two elements:
#' - `chats`: A list of cloned and configured `ellmer::Chat` objects.
#' - `tasks`: A character vector of user prompts (tasks).
#'
#' @keywords internal

tasks_prepare <- function(chat_base, prompts, n_texts) {
  # General case when there may be > 1 system prompts
  ## (e.g. multiple domain roles)
  chats <- lapply(
    prompts$system,
    function(system_prompt) {
      # Each time clone the base chat
      chat <- chat_base$clone(deep = TRUE)
      # Set a system prompt
      chat$set_system_prompt(system_prompt)
      return(chat)
    }
  )
  
  if (length(chats) != 1 & length(chats) != n_texts) {
    abort(
      "Length of system and user prompts must match {.val {n_texts}}, \\
            got {.val {length(chats)}}"
    )
  }
  
  tasks_validate(prompts$task, n_texts)
  
  list(
    chats = chats,
    tasks = prompts$task
  )
}
```

```{r tests-tasks_prepare}
# Mocking a Chat object for testing
MockChat <- R6::R6Class(
  "MockChat",
  public = list(
    system_prompt = NULL,
    set_system_prompt = function(prompt) {
      self$system_prompt <- prompt
    }
  )
)

test_that("tasks_prepare returns correct structure for single system prompt", {
  chat_base <- MockChat$new()
  prompts <- list(
    system = "System Instructions",
    task = c("Task 1", "Task 2", "Task 3")
  )
  n_texts <- 3
  
  result <- tasks_prepare(chat_base, prompts, n_texts)
  
  expect_type(result, "list")
  expect_named(result, c("chats", "tasks"))
  expect_length(result$chats, 1)
  expect_length(result$tasks, 3)
  expect_equal(result$chats[[1]]$system_prompt, "System Instructions")
})

test_that("tasks_prepare returns correct structure for multiple system prompts", {
  chat_base <- MockChat$new()
  prompts <- list(
    system = c("System 1", "System 2"),
    task = c("Task 1", "Task 2")
  )
  n_texts <- 2
  
  result <- tasks_prepare(chat_base, prompts, n_texts)
  
  expect_named(result, c("chats", "tasks"))
  expect_length(result$chats, 2)
  expect_equal(result$chats[[1]]$system_prompt, "System 1")
  expect_equal(result$chats[[2]]$system_prompt, "System 2")
})

test_that("tasks_prepare throws error on mismatched system prompt length", {
  chat_base <- MockChat$new()
  # 2 system prompts for 3 texts is invalid (must be 1 or n_texts)
  prompts <- list(
    system = c("System 1", "System 2"),
    task = c("Task 1", "Task 2", "Task 3")
  )
  n_texts <- 3
  
  expect_error(
    tasks_prepare(chat_base, prompts, n_texts),
    regexp = "Length of system and user prompts must match.+?3.+?\\,\\sgot.+?2",
    fixed = FALSE
  )
})

test_that("tasks_prepare validates user tasks using tasks_validate", {
  chat_base <- MockChat$new()
  # Empty task should trigger tasks_validate error
  prompts <- list(
    system = "System",
    task = c("Task 1", "")
  )
  n_texts <- 2
  
  expect_error(
    tasks_prepare(chat_base, prompts, n_texts),
    regexp = "User prompt is empty after interpolation",
    fixed = TRUE
  )
})

test_that("tasks_prepare clones the base chat deeply", {
  chat_base <- MockChat$new()
  chat_base$system_prompt <- "Original"
  
  prompts <- list(
    system = "New System",
    task = "Task 1"
  )
  
  result <- tasks_prepare(chat_base, prompts, 1)
  
  # Check that the clone has the new prompt but the base remains unchanged
  expect_equal(result$chats[[1]]$system_prompt, "New System")
  expect_equal(chat_base$system_prompt, "Original")
  # Verify they are different objects
  expect_false(identical(chat_base, result$chats[[1]]))
})
```

---

# Expert Analysis

## Prepare Expert Chats

```{r function-prepare_expert_chats}
#' Prepare chat configurations for expert agents
#'
#' @description
#' A high-level internal helper that orchestrates the preparation of expert 
#' sessions for specific analysis roles (linguist, domain expert, or interpreter).
#'
#' @details
#' This function acts as a bridge between the input data and the execution 
#' engine. It extracts necessary variables from the `inputs` list, 
#' interpolates the corresponding templates using [prompts_prepare()], 
#' and then initialises the chat objects via [tasks_prepare()].
#'
#' @param inputs A list containing all necessary data for prompt interpolation:
#'   - `texts`: Character vector of statements to analyse.
#'   - `targets`: Character vector of analysis targets.
#'   - `target_types`: Character vector of localised target types.
#'   - `prompt_templates`: List of paths to markdown template files.
#'   - `language`: Character string (e.g., "en").
#'   - `domain_roles`: (Optional) Character vector for the domain expert role.
#' @param chat_base An `ellmer::Chat` object used as a template for cloning.
#' @param expert_role Character. One of `"linguist"`, `"domain"`, or `"interpreter"`.
#'
#' @return A list containing:
#' - `chats`: A list of configured `ellmer::Chat` objects.
#' - `tasks`: A character vector of interpolated user prompts.
#'
#' @keywords internal
#' @importFrom rlang arg_match

prepare_expert_chats <- function(
    inputs,
    chat_base,
    expert_role = c('linguist', 'domain', 'interpreter')
) {
  expert_role <- rlang::arg_match(
    expert_role,
    c('linguist', 'domain', 'interpreter'),
    multiple = FALSE
  )
  
  # Load instructions for the expert
  prompts <- with(
    inputs,
    prompts_prepare(
      expert_role,
      templates = prompt_templates,
      language = language,
      text = texts,
      target = targets,
      target_type = target_types,
      domain = domain_roles
    )
  )
  
  tasks_prepare(chat_base, prompts, length(inputs$texts))
}
```

```{r tests-prepare_expert_chats}
# Reusing and slightly enhancing the MockChat
MockChat <- R6::R6Class(
  "MockChat",
  public = list(
    system_prompt = NULL,
    set_system_prompt = function(prompt) {
      self$system_prompt <- prompt
    }
  )
)

# Set-up common paths for tests
prompts_dir <- system.file('prompts', 'en', package = 'rcola')
if (!dir.exists(prompts_dir)) prompts_dir <- here::here('inst', 'prompts', 'en')

test_that("prepare_expert_chats successfully prepares linguist chats", {
  # 1. Prepare inputs list
  inputs <- list(
    texts = c("Statement 1", "Statement 2"),
    targets = c("Target A", "Target B"),
    target_types = c("statement", "statement"),
    language = "en",
    domain_roles = "linguistic expert",
    prompt_templates = list(
      `system-linguist` = file.path(prompts_dir, 'system-linguist.md'),
      `user-linguist` = file.path(prompts_dir, 'user-linguist.md')
    )
  )
  
  # Skip if files don't exist in the environment
  skip_if_not(all(file.exists(unlist(inputs$prompt_templates))))
  
  chat_base <- MockChat$new()
  
  # 2. Execute
  result <- prepare_expert_chats(inputs, chat_base, "linguist")
  
  # 3. Verify
  expect_type(result, "list")
  expect_named(result, c("chats", "tasks"))
  expect_length(result$tasks, 2)
  expect_length(result$chats, 1) # Single system prompt case
  expect_match(result$tasks[1], "Statement 1", fixed = TRUE)
  expect_is(result$chats[[1]]$system_prompt, 'character')
})

test_that("prepare_expert_chats successfully prepares domain chats", {
  inputs <- list(
    texts = "Statement about AI",
    targets = "AI",
    target_types = "object",
    language = "en",
    domain_roles = "AI ethics expert",
    prompt_templates = list(
      `system-domain` = file.path(prompts_dir, 'system-domain.md'),
      `user-domain` = file.path(prompts_dir, 'user-domain.md')
    )
  )
  
  skip_if_not(all(file.exists(unlist(inputs$prompt_templates))))
  
  chat_base <- MockChat$new()
  result <- prepare_expert_chats(inputs, chat_base, "domain")
  
  expect_length(result$tasks, 1)
  # Check if domain variable was correctly interpolated in system prompt
  # (Assuming the template uses {domain} placeholder)
  expect_match(result$chats[[1]]$system_prompt, "AI ethics expert", fixed = TRUE)
})

test_that("prepare_expert_chats throws error for invalid role", {
  inputs <- list(texts = "test")
  chat_base <- MockChat$new()
  
  expect_error(
    prepare_expert_chats(inputs, chat_base, "invalid_role"),
    regexp = "must be one of",
    fixed = TRUE
  )
})

test_that("prepare_expert_chats fails if inputs list is missing required elements", {
  # Missing 'texts' which is used in length(inputs$texts)
  inputs <- list(
    language = "en",
    prompt_templates = list()
  )
  chat_base <- MockChat$new()
  
  expect_error(prepare_expert_chats(inputs, chat_base, "linguist"))
})

test_that("prepare_expert_chats handles multiple system prompts (if prompts_prepare returns them)", {
  # This tests the logic flow through to tasks_prepare
  inputs <- list(
    texts = c("T1", "T2"),
    targets = c("A", "B"),
    target_types = c("object", "statement"),
    language = "en",
    domain_roles = c("Content analysis expert", "Statistician"), # Multiple roles
    prompt_templates = list(
      `system-domain` = file.path(prompts_dir, 'system-domain.md'),
      `user-domain` = file.path(prompts_dir, 'user-domain.md')
    )
  )
  
  skip_if_not(all(file.exists(unlist(inputs$prompt_templates))))
  
  chat_base <- MockChat$new()
  result <- prepare_expert_chats(inputs, chat_base, "domain")
  
  # If prompts_prepare returns 2 system prompts, tasks_prepare should return 2 chats
  # Note: This depends on how prompts_prepare handles vectorised 'domain'
  expect_named(result, c("chats", "tasks"))
  expect_type(result$chats, "list")
  expect_length(result$tasks, 2)
  expect_length(result$chats, 2)
  expect_is(result$chats[[1]]$system_prompt, 'character')
  expect_is(result$chats[[2]]$system_prompt, 'character')
  expect_match(
    result$chats[[1]]$system_prompt,
    "Content analysis expert",
    fixed = TRUE
  )
  expect_match(
    result$chats[[2]]$system_prompt,
    "Statistician",
    fixed = TRUE
  )
})
```

## Execute Role

```{r function-execute_role}
#' Execute expert role analysis
#'
#' @description
#' An internal worker that handles the execution of LLM requests for a 
#' specific expert role. It switches between parallel execution 
#' (when a single system prompt is shared) and sequential execution (when 
#' each task requires a unique system prompt).
#'
#' @param expert_role Character. The role to execute (e.g., "linguist").
#' @param inputs A list containing texts, targets, and prompt templates.
#' @param chat_base An `ellmer::Chat` object used as a template.
#' @param info Character. A descriptive label for logging and error messages.
#' @param rpm Integer. Requests per minute limit for parallel execution.
#' @param verbose Logical. Whether to print progress messages to the console.
#'
#' @return A character vector of LLM responses, one for each input text.
#' @throws An error if the number of responses does not match the number of 
#'   input texts, or if the LLM returns non-character data.
#'
#' @keywords internal
#' @importFrom ellmer parallel_chat_text

execute_role <- function(
    expert_role,
    inputs,
    chat_base,
    info,
    rpm = 500,
    verbose = TRUE
) {
  tasks <- prepare_expert_chats(
    inputs,
    chat_base = chat_base,
    expert_role = expert_role
  )
  
  if (verbose) cat('    \U1F4CA', glue::glue('{info}...'), '\n')
  
  # If multiple system_prompts, e.g
  if (length(tasks$chats) > 1) {
    # Use sequential analysis
    results <- vapply(
      seq_along(tasks$tasks),
      function(task_i) {
        chat <- tasks$chats[[task_i]]
        result <- chat$chat(tasks$tasks[[task_i]], echo = 'none')
        if (!is.character(result)) {
          abort("Unexpected results in {info} {task_i}")
        }
        result
      },
      character(1)
    ) |> catch(tolower(info))
  } else {
    # Use parallel analysis
    results <- ellmer::parallel_chat_text(
      # Unpack the Chat object
      chat = tasks$chats[[1]],
      prompts = tasks$tasks,
      rpm = rpm
    ) |> catch(tolower(info))
  }
  
  if (!is.character(results) || (length(results) != length(inputs$texts))) {
    abort(
      c(
        "Unexpected results in {info}",
        "x" = "Input and output lengths differ: {length(inputs$texts)} vs {length(results)}"
      )
    )
  }
  
  results
}
```

```{r tests-execute_role}
# Reusing the MockChat
MockChat <- R6::R6Class(
  "MockChat",
  public = list(
    system_prompt = NULL,
    set_system_prompt = function(prompt) {
      self$system_prompt <- prompt
    },
    chat = function(..., echo = NULL) paste("Response", seq_along(list(...)))
  )
)

test_that("execute_role uses parallel execution when there is one chat object", {
  # 1. Set-up Mocks
  mockery::stub(
    execute_role,
    "prepare_expert_chats",
    list(
      chats = list(MockChat$new()), # Only 1 chat -> parallel
      tasks = c("Content Analysis Prompt 1", "Content Analysis Prompt 2")
    )
  )
  
  mock_parallel <- mockery::mock(c("Response A", "Response B"))
  mockery::stub(execute_role, "ellmer::parallel_chat_text", mock_parallel)
  
  inputs <- list(texts = c("T1", "T2"))
  
  # 2. Execute
  res <- execute_role(
    "linguist", inputs, MockChat$new(), "Testing Parallel", verbose = FALSE
  )
  
  # 3. Verify
  expect_equal(res, c("Response A", "Response B"))
  mockery::expect_called(mock_parallel, 1)
  # Check arguments passed to parallel_chat_text
  args <- mockery::mock_args(mock_parallel)[[1]]
  expect_equal(
    args$prompts,
    c("Content Analysis Prompt 1", "Content Analysis Prompt 2")
  )
})

test_that("execute_role uses sequential execution when there are multiple chats", {
  # 1. Set-up Mocks
  chat1 <- MockChat$new()
  chat2 <- MockChat$new()
  
  mockery::stub(
    execute_role,
    "prepare_expert_chats",
    list(
      chats = list(chat1, chat2), # 2 chats -> sequential
      tasks = c("Task 1", "Task 2")
    )
  )
  
  inputs <- list(texts = c("T1", "T2"))
  
  # 2. Execute
  res <- execute_role(
    "domain", inputs, MockChat$new(), "Testing Sequential", verbose = FALSE
  )
  
  # 3. Verify
  expect_equal(res, c("Response 1", "Response 1"))
  # For every Chat, there is only one response
})

test_that("execute_role throws error if output length is mismatched", {
  mockery::stub(
    execute_role,
    "prepare_expert_chats",
    list(chats = list(MockChat$new()), tasks = c("P1", "P2"))
  )
  
  # Parallel returns only 1 response for 2 inputs
  mockery::stub(execute_role, "ellmer::parallel_chat_text", "Single Response")
  
  inputs <- list(texts = c("T1", "T2"))
  
  expect_error(
    execute_role("linguist", inputs, MockChat$new(), "Error Test", verbose = FALSE),
    regexp = "Input and output lengths differ",
    fixed = TRUE
  )
})

test_that("execute_role prints progress message when verbose is TRUE", {
  mockery::stub(
    execute_role,
    "prepare_expert_chats",
    list(chats = list(MockChat$new()), tasks = "P1")
  )
  mockery::stub(execute_role, "ellmer::parallel_chat_text", "Result A")
  
  expect_output(
    execute_role(
      "linguist", list(texts = "T1"), MockChat$new(), "MyInfo", verbose = TRUE
    ),
    regexp = "MyInfo..."
  )
})
```

---

# Debate Stage

## Prepare Debater Chats

```{r function-prepare_debater_chats}
#' Prepare chat configurations for debate agents
#'
#' @description
#' Prepares the prompts and chat objects for Stage 2 (Debate). It takes the 
#' linguistic, domain, and social media analysis results from Stage 1 and 
#' injects them into the 'debater' templates.
#'
#' @param stance Character. The stance label to be debated for (used in prompts).
#' @param inputs A list containing:
#'   - `texts`, `targets`, `target_types`, `language`, `prompt_templates`
#'   - `analysis_results`: A list with `linguistic`, `domain`, and `social_media` 
#'     vectors from Stage 1.
#' @param chat_base An `ellmer::Chat` object.
#'
#' @return A list with two elements:
#'   - `chats`: A list containing a single `ellmer::Chat` object.
#'   - `tasks`: A character vector of user prompts for each input text.
#'
#' @throws An error via `abort()` if the templates result in multiple 
#'   distinct system prompts.
#'
#' @keywords internal

prepare_debater_chats <- function(
    stance,
    inputs,
    chat_base
) {
  # Prepare prompts using all context from Stage 1
  prompts <- with(
    inputs,
    prompts_prepare(
      'debater',
      language = language,
      templates = prompt_templates,
      text = texts,
      target_type = target_types,
      target = targets,
      stance = stance,
      LingResponse = analysis_results$linguistic,
      ExpertResponse = analysis_results$domain,
      UserResponse = analysis_results$social_media
    )
  )
  
  # Stage 2 requires a consistent system prompt across all items
  if (length(prompts$system) > 1) {
    abort(
      c(
        "Multiple system prompts are not allowed at stage 2",
        "i" = "Consider checking variable placeholders in system templates"
      )
    )
  }
  
  tasks_prepare(chat_base, prompts, length(inputs$texts))
}
```

```{r tests-prepare_debater_chats}
# Reusing the MockChat
MockChat <- R6::R6Class(
  "MockChat",
  public = list(
    system_prompt = NULL,
    set_system_prompt = function(prompt) {
      self$system_prompt <- prompt
    },
    chat = function(..., echo = NULL) paste("Response", seq_along(list(...)))
  )
)

test_that("prepare_debater_chats correctly passes Stage 1 results to prompts_prepare", {
  # 1. Set-up Data
  inputs <- list(
    texts = "Text 1",
    targets = "Target 1",
    target_types = "statement",
    language = "en",
    prompt_templates = list(t = "inst/prompts"),
    analysis_results = list(
      linguistic = "Linguistic Analysis",
      domain = "Domain Expert Analysis",
      social_media = "Social Media Analysis"
    )
  )
  
  # 2. Mock prompts_prepare to verify arguments
  mock_prompts <- mockery::mock(list(
    system = "System Prompt",
    user = "User Prompt"
  ))
  mockery::stub(prepare_debater_chats, "prompts_prepare", mock_prompts)
  
  # Mock tasks_prepare to return a standard structure
  mockery::stub(prepare_debater_chats, "tasks_prepare", list(chats = list(1), tasks = "T"))
  
  # 3. Execute
  prepare_debater_chats("negative", inputs, MockChat$new())
  
  # 4. Verify arguments passed to prompts_prepare
  args <- mockery::mock_args(mock_prompts)[[1]]
  expect_equal(args[[1]], "debater")
  expect_equal(args$stance, "negative")
  expect_equal(args$LingResponse, "Linguistic Analysis")
  expect_equal(args$ExpertResponse, "Domain Expert Analysis")
  expect_equal(args$UserResponse, "Social Media Analysis")
})

test_that("prepare_debater_chats aborts if multiple system prompts are generated", {
  # Mock prompts_prepare to return 2 system prompts
  mockery::stub(prepare_debater_chats, "prompts_prepare", list(
    system = c("System 1", "System 2"),
    user = c("User 1", "User 2")
  ))
  
  inputs <- list(texts = c("T1", "T2"), analysis_results = list())
  
  # Check for our custom abort message
  expect_error(
    prepare_debater_chats("con", inputs, MockChat$new()),
    regexp = "Multiple system prompts are not allowed at stage 2"
  )
})

test_that("prepare_debater_chats returns tasks_prepare output on success", {
  mockery::stub(prepare_debater_chats, "prompts_prepare", list(
    system = "Single System",
    user = "User Prompt"
  ))
  
  expected_output <- list(chats = list("MockChatObject"), tasks = "Final Prompt")
  mockery::stub(prepare_debater_chats, "tasks_prepare", expected_output)
  
  res <- prepare_debater_chats("neutral", list(texts = "T1"), MockChat$new())
  expect_equal(res, expected_output)
})

test_that("prepare_debater_chats fails with multiple system prompts", {
  mockery::stub(prepare_debater_chats, "prompts_prepare", list(
    system = c("System Prompt 1", "System Prompt 2"),
    user = "User Prompt"
  ))
  
  expect_error(
    prepare_debater_chats("positive", list(texts = "T1"), MockChat$new()),
    "Multiple system prompts are not allowed at stage 2",
    fixed = TRUE
  )
})
```

---

# Judgment Stage

## Prepare Judger Chats

```{r function-prepare_judger_chats}
#' Prepare chat configurations for judge agent
#'
#' @param inputs Input texts
#' @param chat_base Base chat configuration
#'
#' @return List of chat configurations
#' @keywords internal

prepare_judger_chats <- function(inputs, chat_base) {
  # Implementation
}
```

```{r tests-prepare_judger_chats}
test_that('prepare_judger_chats is a function', {
  expect_true(inherits(prepare_judger_chats, 'function'))
})
```

---

# Pipeline Stages

## Stage 1: Parallel Analysis

```{r function-stage_1_parallel_analysis}
#' Execute Stage 1: Parallel expert analysis
#'
#' @description
#' Orchestrates the first stage of the analysis pipeline. It validates the 
#' input data, then sequentially triggers three distinct expert analyses: 
#' linguistic, domain-specific, and social media interpretation.
#'
#' @details
#' Although the function name contains "parallel", it refers to the 
#' independent nature of the expert roles (they don't depend on each 
#' other's output). Each role is executed via [execute_role()], which 
#' may itself use parallel API requests.
#'
#' @param inputs A list containing all necessary data:
#'   - `texts`: Character vector to analyse.
#'   - `targets`: Character vector of targets.
#'   - `target_types`: Character vector of target types.
#'   - `prompt_templates`: List of template paths.
#'   - `language`: Character string.
#'   - `domain_roles`: Character vector for the domain expert.
#' @param chat_base An `ellmer::Chat` object used as a template.
#' @param verbose Logical. Whether to print stage progress and timing.
#' @param rpm Integer. Requests per minute limit for API calls.
#'
#' @return The `inputs` list with an additional element `analysis_results`, 
#'   which is a named list containing character vectors:
#'   - `linguistic`: Results from the linguist role.
#'   - `domain`: Results from the domain expert role.
#'   - `social_media`: Results from the interpreter role.
#'
#' @keywords internal
#' @importFrom tictoc tic toc
#' @importFrom glue glue

stage_1_parallel_analysis <- function(
    inputs,
    chat_base,
    verbose = TRUE,
    rpm = 500
) {
  # Validate input structure
  required_fields <- c(
    'prompt_templates',
    'language',
    'texts',
    'targets',
    'target_types',
    'domain_roles'
  )
  validate_fields(inputs, required_fields, 'expert analysis')
  
  n <- length(inputs$texts)
  
  tictoc::tic('Stage 1')
  if (verbose) {
    cat(
      glue::glue("\U23F3 Stage 1: Expert analysis ({n} items)..."), 
      "\n"
    )
  }
  
  expert_roles <- c('linguist', 'domain', 'interpreter')
  expert_info <- c(
    linguist = 'Linguistic analysis',
    domain = 'Domain expert analysis',
    interpreter = 'Social media interpretation'
  )
  analysis_results <- lapply(
    expert_roles,
    function(role) {
      execute_role(
        role,
        inputs = inputs,
        chat_base = chat_base,
        info = expert_info[[role]],
        rpm = rpm,
        verbose = verbose
      )
    }
  )
  names(analysis_results) <- c('linguistic', 'domain', 'social_media')
  
  # The result is a list of character vectors of length length(texts)
  ## One vector for each role
  inputs[['analysis_results']] <- analysis_results
  
  tictoc::toc(func.toc = stage_complete, quiet = !verbose)
  
  inputs
}
```

```{r tests-stage_1_parallel_analysis}
test_that("stage_1_parallel_analysis validates inputs and returns results", {
  # 1. Set-up Data
  inputs <- list(
    texts = c("T1", "T2"),
    targets = c("A", "B"),
    target_types = c("statement", "statement"),
    language = "en",
    domain_roles = "AI Expert",
    prompt_templates = list(a = 1)
  )
  
  # 2. Set-up Mocks
  # Mock validate_fields to just pass
  mockery::stub(stage_1_parallel_analysis, "validate_fields", TRUE)
  
  # Mock execute_role to return role-specific strings
  mock_execute <- mockery::mock(
    c("L1", "L2"), # First call: linguist
    c("D1", "D2"), # Second call: domain
    c("S1", "S2")  # Third call: interpreter
  )
  mockery::stub(stage_1_parallel_analysis, "execute_role", mock_execute)
  
  # Mock tictoc to avoid console noise/errors
  mockery::stub(stage_1_parallel_analysis, "tictoc::tic", NULL)
  mockery::stub(stage_1_parallel_analysis, "tictoc::toc", NULL)
  
  # 3. Execute
  result <- stage_1_parallel_analysis(
    inputs, 
    chat_base = MockChat$new(), 
    verbose = FALSE
  )
  
  # 4. Verify
  expect_true("analysis_results" %in% names(result))
  expect_named(result$analysis_results, c("linguistic", "domain", "social_media"))
  expect_equal(result$analysis_results$linguistic, c("L1", "L2"))
  expect_equal(result$analysis_results$domain, c("D1", "D2"))
  expect_equal(result$analysis_results$social_media, c("S1", "S2"))
  
  # Verify execute_role was called exactly 3 times
  mockery::expect_called(mock_execute, 3)
})

test_that("stage_1_parallel_analysis fails if required fields are missing", {
  # Missing domain_roles
  incomplete_inputs <- list(
    texts = "T1",
    targets = "X",
    target_types = "statement",
    language = "en",
    prompt_templates = list()
  )
  
  # We don't stub validate_fields here because we want to test that it's actually called
  # and catches the error as expected.
  mockery::stub(stage_1_parallel_analysis, "validate_fields", validate_fields)
  expect_error(
    stage_1_parallel_analysis(incomplete_inputs, MockChat$new(), verbose = FALSE),
    regexp = "misses required fields: `domain_roles`"
  )
})

test_that("stage_1_parallel_analysis respects verbose flag", {
  inputs <- list(
    texts = "T1", targets = "A", target_types = "statement", 
    language = "en", domain_roles = "AI Expert", prompt_templates = list()
  )
  
  mockery::stub(stage_1_parallel_analysis, "validate_fields", TRUE)
  mockery::stub(stage_1_parallel_analysis, "execute_role", list("Result"))
  mockery::stub(stage_1_parallel_analysis, "tictoc::tic", NULL)
  mockery::stub(stage_1_parallel_analysis, "tictoc::toc", NULL)
  
  expect_output(
    stage_1_parallel_analysis(inputs, MockChat$new(), verbose = TRUE),
    regexp = "Stage 1: Expert analysis"
  )
  
  expect_silent(
    stage_1_parallel_analysis(inputs, MockChat$new(), verbose = FALSE)
  )
})
```

## Stage 2: Parallel Debates

```{r function-stage_2_parallel_debates}
#' Execute Stage 2: Parallel agent debates
#'
#' @param inputs Input texts
#' @param chat_base Chat configuration
#' @param verbose Show progress
#' @param rpm Requests per minute limit
#'
#' @return Debate results
#' @keywords internal

stage_2_parallel_debates <- function(
    inputs,
    chat_base,
    verbose = TRUE,
    rpm = 500
) {
  # Validate input structure
  required_fields <- c(
    'prompt_templates',
    'language',
    'texts',
    'targets',
    'target_types',
    'analysis_results',
    'scale'
  )
  validate_fields(inputs, required_fields, 'stance debates')
  
  n <- length(inputs$texts)
  
  # Ensure Stage 1 results are present and valid
  validate_inputs(inputs$analysis_results$linguistic, n, 'Linguistic analysis')
  validate_inputs(inputs$analysis_results$domain, n, 'Domain expert analysis')
  validate_inputs(inputs$analysis_results$social_media, n, 'Social media analysis')
  
  tictoc::tic('Stage 2')
  if (verbose) {
    cat(
      glue::glue(
        "\U23F3 Stage 2: Parallel debates ({n} items × 3 stances)..."), 
      "\n"
    )
  }
  
  stance_labels <- c('positive', 'negative', 'neutral')
  
  # 1. Prepare tasks for each stance
  debater_tasks <- lapply(
    stance_labels,
    function(stance_label) {
      prepare_debater_chats(
        stance = l(
          inputs$language,
          glue::glue('stance_{stance_label}'),
          inputs$scale
        ),
        inputs = inputs,
        chat_base = chat_base
      )
    }
  )
  names(debater_tasks) <- stance_labels
  
  # 2. Execute parallel debates for each stance group
  inputs[['debate_results']] <- lapply(
    debater_tasks,
    function(debater_task) {
      ellmer::parallel_chat_text(
        chat = debater_task$chats[[1]],
        prompts = debater_task$tasks,
        rpm = rpm
      ) |> catch('stance debates')
    }
  )
  
  tictoc::toc(func.toc = stage_complete, quiet = !verbose)
  
  inputs
}
```

```{r tests-stage_2_parallel_debates}
# Reusing the MockChat
MockChat <- R6::R6Class(
  "MockChat",
  public = list(
    system_prompt = NULL,
    set_system_prompt = function(prompt) {
      self$system_prompt <- prompt
    },
    chat = function(..., echo = NULL) paste("Response", seq_along(list(...)))
  )
)

test_that("stage_2_parallel_debates executes three debate branches", {
  # 1. Set-up Data
  inputs <- list(
    texts = c("T1", "T2"),
    targets = c("A", "B"),
    target_types = c("statement", "statement"),
    language = "en",
    scale = 'likert',
    prompt_templates = list(t = 1),
    analysis_results = list(
      linguistic = c("L1", "L2"),
      domain = c("D1", "D2"),
      social_media = c("S1", "S2")
    )
  )
  
  # 2. Set-up Mocks
  # mockery::stub(stage_2_parallel_debates, "validate_fields", TRUE)
  # mockery::stub(stage_2_parallel_debates, "validate_inputs", TRUE)
  mockery::stub(stage_2_parallel_debates, "tictoc::tic", NULL)
  mockery::stub(stage_2_parallel_debates, "tictoc::toc", NULL)
  
  # Mock l() localization helper
  mockery::stub(stage_2_parallel_debates, "l", function(lang, key, scale) key)
  
  # Mock prepare_debater_chats
  # We return a dummy chat and tasks
  mockery::stub(
    stage_2_parallel_debates,
    "prepare_debater_chats",
    function(stance, ...) {
      list(chats = list(MockChat$new()), tasks = paste(stance, "task"))
    }
  )
  
  # Mock parallel_chat_text to return stance-specific results
  mock_parallel <- mockery::mock(
    "Pos Result", # first call (positive)
    "Neg Result", # second call (negative)
    "Neu Result"  # third call (neutral)
  )
  mockery::stub(stage_2_parallel_debates, "ellmer::parallel_chat_text", mock_parallel)
  
  # 3. Execute
  result <- stage_2_parallel_debates(inputs, MockChat$new(), verbose = FALSE)
  
  # 4. Verify
  expect_named(result$debate_results, c("positive", "negative", "neutral"))
  expect_equal(result$debate_results$positive, "Pos Result")
  expect_equal(result$debate_results$negative, "Neg Result")
  expect_equal(result$debate_results$neutral, "Neu Result")
  
  # Verify it was called 3 times (once for each stance)
  mockery::expect_called(mock_parallel, 3)
})

test_that("stage_2_parallel_debates validates Stage 1 results before proceeding", {
  inputs <- list(
    texts = c("T1", "T2"),
    analysis_results = list(
      linguistic = "Too Short", # Should be length 2
      domain = c("D1", "D2"),
      social_media = c("S1", "S2")
    )
  )
  
  # Mock validate_fields but let validate_inputs run
  mockery::stub(stage_2_parallel_debates, "validate_fields", TRUE)
  
  # validate_inputs should throw error because linguistic length (1) != n (2)
  expect_error(
    stage_2_parallel_debates(inputs, MockChat$new(), verbose = FALSE),
    "Linguistic analysis returned unexpected results",
    fixed = TRUE
  )
})

test_that("stage_2_parallel_debates passes the correct RPM to parallel_chat_text", {
  # Set-up minimal valid inputs
  inputs <- list(
    texts = "T1", targets = "A", target_types = "statement", 
    language = "en", scale = "likert",
    prompt_templates = list(t = 1),
    analysis_results = list(
      linguistic = "L1", domain = "D1", social_media = "S1"
    )
  )
  
  # Mock dependencies
  # mockery::stub(stage_2_parallel_debates, "validate_fields", TRUE)
  # mockery::stub(stage_2_parallel_debates, "validate_inputs", TRUE)
  mockery::stub(stage_2_parallel_debates, "l", "mock stance")
  mockery::stub(
    stage_2_parallel_debates,
    "prepare_debater_chats", 
    list(chats = list(MockChat$new()), tasks = "task")
  )
  
  mock_parallel <- mockery::mock("Pos Result", "Neg Result", "Neu Result")
  mockery::stub(stage_2_parallel_debates, "ellmer::parallel_chat_text", mock_parallel)
  
  # Execute with custom RPM
  stage_2_parallel_debates(inputs, MockChat$new(), rpm = 42, verbose = FALSE)
  
  # Verify RPM in the first call to parallel_chat_text
  args <- mockery::mock_args(mock_parallel)[[1]]
  expect_equal(args$rpm, 42)
  expect_equal(args$prompts, "task")
})

test_that("stage_2_parallel_debates correctly localises stances for each branch", {
  # 1. Set-up data
  inputs <- list(
    texts = "T1", targets = "A", target_types = "statement", 
    language = "en", scale = "likert",
    prompt_templates = list(t = "file"),
    analysis_results = list(
      linguistic = "L1", domain = "D1", social_media = "S1"
    )
  )
  
  # 2. Set-up Mocks
  # mockery::stub(stage_2_parallel_debates, "validate_fields", TRUE)
  # mockery::stub(stage_2_parallel_debates, "validate_inputs", TRUE)
  mockery::stub(stage_2_parallel_debates, "tictoc::tic", NULL)
  mockery::stub(stage_2_parallel_debates, "tictoc::toc", NULL)
  
  # Мокаем l(), чтобы она возвращала узнаваемый маркер для каждой позиции
  mockery::stub(stage_2_parallel_debates, "l", function(lang, key, scale) {
    paste0("loc_", key)
  })
  
  # Мокаем prepare_debater_chats, чтобы отследить, какой stance в неё пришел
  mock_prepare <- mockery::mock(
    list(chats = list(MockChat$new()), tasks = "task"), # для positive
    list(chats = list(MockChat$new()), tasks = "task"), # для negative
    list(chats = list(MockChat$new()), tasks = "task")  # для neutral
  )
  mockery::stub(stage_2_parallel_debates, "prepare_debater_chats", mock_prepare)
  
  # Mock parallel_chat_text to return stance-specific results
  mock_parallel <- mockery::mock(
    "Pos Result", # first call (positive)
    "Neg Result", # second call (negative)
    "Neu Result"  # third call (neutral)
  )
  mockery::stub(stage_2_parallel_debates, "ellmer::parallel_chat_text", mock_parallel)
  
  # 3. Execute
  stage_2_parallel_debates(inputs, MockChat$new(), verbose = FALSE)
  
  # 4. Verify
  # Достаем аргументы всех трех вызовов prepare_debater_chats
  mockery::expect_called(mock_prepare, 3)
  
  prep_args <- mockery::mock_args(mock_prepare)
  
  # Проверяем, что в каждом вызове аргумент stance был локализован правильно
  expect_equal(prep_args[[1]]$stance, "loc_stance_positive")
  expect_equal(prep_args[[2]]$stance, "loc_stance_negative")
  expect_equal(prep_args[[3]]$stance, "loc_stance_neutral")
  
  # Также можно проверить, что inputs и chat_base передаются без изменений
  expect_equal(prep_args[[1]]$inputs$language, "en")
  
  expect_equal(prep_args[[1]]$inputs$scale, "likert")
})

test_that("stage_2_parallel_debates handles empty input texts gracefully", {
  # If n=0, the loops should still work but return empty lists
  inputs <- list(
    texts = character(0), targets = character(0), target_types = character(0),
    language = "en", scale = "likert",
    prompt_templates = list(t = "file"),
    analysis_results = list(
      linguistic = character(0), domain = character(0), social_media = character(0)
    )
  )
  
  # mockery::stub(stage_2_parallel_debates, "validate_fields", TRUE)
  # mockery::stub(stage_2_parallel_debates, "validate_inputs", TRUE)
  mockery::stub(
    stage_2_parallel_debates,
    "prepare_debater_chats", 
    list(chats = list(MockChat$new()), tasks = character(0))
  )
  # Mock parallel_chat_text to return stance-specific results
  mock_parallel <- mockery::mock(character(0), character(0), character(0))
  mockery::stub(stage_2_parallel_debates, "ellmer::parallel_chat_text", mock_parallel)
  
  result <- stage_2_parallel_debates(inputs, MockChat$new(), verbose = FALSE)
  
  expect_length(result$debate_results$positive, 0)
  expect_named(result$debate_results, c("positive", "negative", "neutral"))
})
```

## Stage 3: Parallel Judgment

```{r function-stage_3_parallel_judgment}
#' Execute Stage 3: Judge synthesis
#'
#' @param inputs Input texts
#' @param chat_base Chat configuration
#' @param verbose Show progress
#' @param rpm Requests per minute limit
#' @param ... Additional arguments
#'
#' @return Final stance judgments
#' @keywords internal

stage_3_parallel_judgment <- function(
    inputs,
    chat_base,
    verbose = TRUE,
    rpm = 500,
    ...
) {
  # Implementation
}
```

```{r tests-stage_3_parallel_judgment}
test_that('stage_3_parallel_judgment is a function', {
  expect_true(inherits(stage_3_parallel_judgment, 'function'))
})
```

---

# Main Analysis Function

## LLM Stance Detection

```{r function-llm_stance}
#' Detect stance using COLA framework
#'
#' @param text Input text(s)
#' @param ... Additional arguments passed to methods
#'
#' @return stance_result object
#' @export

llm_stance <- function(text, ...) {
  UseMethod("llm_stance")
}
```

```{r function-llm_stance.character}
#' Stance detection for character vectors
#'
#' @param text Character vector of texts
#' @param target Target entity or statement
#' @param chat_base Chat model configuration
#' @param type Analysis type: "object" or "statement"
#' @param language Language code
#' @param scale Stance scale: "categorical", "numeric", or "likert"
#' @param domain_role Optional domain-specific role
#' @param prompts_dir Optional custom prompts directory
#' @param verbose Show progress messages
#' @param rpm Requests per minute limit
#' @param ... Additional arguments
#'
#' @return stance_result object
#' @export

llm_stance.character <- function(
    text,
    target,
    chat_base,
    type = c("object", "statement"),
    language = rcola_available_languages(),
    scale = c("categorical", "numeric", "likert"),
    domain_role = NULL,
    prompts_dir = NULL,
    verbose = TRUE,
    rpm = 20,
    ...
) {
  # Implementation
}
```

```{r examples-llm_stance.character}
# Example usage would go here
# result <- llm_stance(
#   text = "I support renewable energy",
#   target = "renewable energy",
#   chat_base = chat_model,
#   type = "object",
#   scale = "categorical"
# )
```

```{r tests-llm_stance.character}
test_that('llm_stance.character is a function', {
  expect_true(inherits(llm_stance.character, 'function'))
})
```

```{r function-llm_stance.data.frame}
#' Stance detection for data frames
#'
#' @param data Input data frame
#' @param text Column name containing texts
#' @param target Column name or vector of targets
#' @param type Analysis type
#' @param domain_role Optional domain role
#' @param language Language code
#' @param chat_base Chat model configuration
#' @param scale Stance scale
#' @param prompts_dir Optional custom prompts directory
#' @param verbose Show progress
#' @param rpm Requests per minute limit
#' @param .output_col Name of output column
#' @param ... Additional arguments
#'
#' @return Data frame with stance results
#' @export

llm_stance.data.frame <- function(
    data,
    text,
    target,
    type = NULL,
    domain_role = NULL,
    language = NULL,
    chat_base,
    scale = "categorical",
    prompts_dir = NULL,
    verbose = TRUE,
    rpm = 20,
    .output_col = ".stance",
    ...
) {
  # Implementation
}
```

```{r examples-llm_stance.data.frame}
# Example usage would go here
# results <- llm_stance(
#   data = tweets_df,
#   text = "text",
#   target = "climate_change",
#   chat_base = chat_model
# )
```

```{r tests-llm_stance.data.frame}
test_that('llm_stance.data.frame is a function', {
  expect_true(inherits(llm_stance.data.frame, 'function'))
})
```

---

# Result Inspection and Display

## Print Method

```{r function-print.stance_result}
#' Print stance_result object
#'
#' @param x stance_result object
#' @param ... Additional arguments
#'
#' @return Invisibly x
#' @export

print.stance_result <- function(x, ...) {
  # Implementation
}
```

```{r examples-print.stance_result}
# result <- llm_stance(...)
# print(result)
```

```{r tests-print.stance_result}
test_that('print.stance_result is a function', {
  expect_true(inherits(print.stance_result, 'function'))
})
```

## Summary Method

```{r function-summary.stance_result}
#' Summarize stance_result object
#'
#' @param object stance_result object
#' @param ... Additional arguments
#'
#' @return Invisibly object
#' @export

summary.stance_result <- function(object, ...) {
  # Implementation
}
```

```{r examples-summary.stance_result}
# result <- llm_stance(...)
# summary(result)
```

```{r tests-summary.stance_result}
test_that('summary.stance_result is a function', {
  expect_true(inherits(summary.stance_result, 'function'))
})
```

## Data Frame Conversion

```{r function-as.data.frame.stance_result}
#' Convert stance_result to data frame
#'
#' @param x stance_result object
#' @param row.names Optional row names
#' @param optional Logical
#' @param ... Additional arguments
#'
#' @return Data frame
#' @export

as.data.frame.stance_result <- function(
    x,
    row.names = NULL,
    optional = FALSE,
    ...
) {
  # Implementation
}
```

```{r examples-as.data.frame.stance_result}
# result <- llm_stance(...)
# df <- as.data.frame(result)
```

```{r tests-as.data.frame.stance_result}
test_that('as.data.frame.stance_result is a function', {
  expect_true(inherits(as.data.frame.stance_result, 'function'))
})
```

## Inspect Method

```{r function-inspect}
#' Inspect stance_result contents
#'
#' @param x stance_result object
#' @param ... Additional arguments passed to methods
#'
#' @return Invisibly x
#' @export

inspect <- function(x, ...) {
  UseMethod("inspect")
}
```

```{r function-inspect.stance_result}
#' Inspect stance_result object details
#'
#' @param x stance_result object
#' @param what Primary inspection level: 'metadata', 'analysis', or 'debates'
#' @param within Secondary inspection level
#' @param index Row index to inspect
#'
#' @return Invisibly x
#' @export

inspect.stance_result <- function(
    x,
    what = c('metadata', 'analysis', 'debates'),
    within = NULL,
    index = 1
) {
  # Implementation
}
```

```{r examples-inspect.stance_result}
# result <- llm_stance(...)
# inspect(result, what = 'metadata')
# inspect(result, what = 'analysis', within = 'linguistic_analysis')
# inspect(result, what = 'debates', within = 'counterarguments', index = 1)
```

```{r tests-inspect.stance_result}
test_that('inspect.stance_result is a function', {
  expect_true(inherits(inspect.stance_result, 'function'))
})
```

---

# Inflate

```{r development-inflate, eval=FALSE}
fusen::inflate(
  flat_file = 'dev/flat_rcola.Rmd',
  codecov = TRUE,
  clean = TRUE
)
```
